### 2023-11-28 text/proposta/prop.tex
r

% RUN in terminal (without bibliography):
% pdflatex -output-directory=/Users/salvatorpes/Desktop/LFEUI/text/proposta /Users/salvatorpes/Desktop/LFEUI/text/proposta/prop.tex

\documentclass{article}

\author{
    \begin{tabular}{rl}
        Estêvão Gomes (ist1102650) & Sofia Nunes (ist1102633) \\
        Pedro Curvo (ist1102716) & Salvador Torpes (ist1102474)
    \end{tabular}
}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[letterpaper,top=10mm,bottom=15mm,left=10mm,right=10mm,marginparwidth=1.75cm]{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{url}
\usepackage{csquotes}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{textcomp}
\usepackage{pdflscape}
\usepackage{makeidx}
% \usepackage{tocbibind}
\providecommand{\tightlist}{\relax}
\usepackage{tocloft}
\renewcommand{\cftsecindent}{0em}
\renewcommand{\cftsubsecindent}{1em}
\renewcommand{\cftsecfont}{\bfseries}
\renewcommand{\cftsubsecfont}{\itshape}
\setlength{\cftsubsecnumwidth}{0em}

\usepackage[version=4]{mhchem}
\usepackage{hyperref} % Remove "pdftex" option here
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{ragged2e}
\usepackage{xkeyval}
%\usepackage{minted}
%\usemintedstyle{manni}
\usepackage{listings}
\usepackage{amssymb}

\usepackage{xcolor}
\usepackage{tikz}

\usetikzlibrary{positioning}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{adjustbox}
\usepackage{sidecap}
\usepackage{graphicx}

\usepackage{tikz-3dplot}
\usepackage{pgfplots}
\usetikzlibrary{calc, 3d, arrows}

\usetikzlibrary{shapes.geometric, arrows}

\lstset{
    language=Python,
    basicstyle=\ttfamily,
    keywordstyle=\color{blue},
    commentstyle=\color{gray},
    stringstyle=\color{orange},
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    breaklines=true,
    frame=tb,
    framexleftmargin=2em,
    xleftmargin=2em,
}

%\usepackage{fontspec}

%\setmonofont{Fira Code}

\fancyhf{}
\cfoot{\thepage}
\fancyhf{} % Clear all header and footer fields
\renewcommand{\headrulewidth}{0pt} % Remove the header rule line
\cfoot{\thepage} % Set the page number in the center of the footer

\pagestyle{fancy} % Apply the fancy page style

\setlength\columnsep{20pt}

\renewcommand{\familydefault}{\sfdefault}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\makeatletter
\newenvironment{figurehere}
{\def\@captype{figure}}
{}
\makeatother

\hypersetup{
  colorlinks,
  linkcolor=blue,
  anchorcolor=black,
  citecolor=cyan,
  filecolor=cyan,
  menucolor=cyan,
  urlcolor=cyan,
  bookmarksopen=true,
  bookmarksnumbered=true
}

\makeindex

\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
LFEUI - Investigation Proposal\vspace{-5mm}}
\date{November 2023}

\usepackage{sansmathfonts}
\usepackage[T1]{fontenc}
\usepackage[OT1]{fontenc}

\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}

\usepackage[style=numeric]{biblatex} % Choose your desired citation style
\addbibresource{../references/prop.bib} % Specify your .bib file

\begin{document}

\renewcommand{\arraystretch}{1.5}
\setlength{\columnseprule}{0.4pt}
\tdplotsetmaincoords{70}{110} % Set the viewing angle
\newcolumntype{M}[1]{>{\centering\arraybackslash\vspace{#1}}m{0.5\linewidth}<{\vspace{#1}}}
\newcolumntype{C}[2]{>{\centering\arraybackslash\vspace{#1}\rule{0pt}{#1}\hspace{0pt}}m{#2}}
\newcolumntype{w}[1]{>{\centering\arraybackslash}m{#1}}

\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif

\maketitle

\vspace{-5mm}

\hrulefill

\begin{multicols}{2}

% \tableofcontents

\section{Proposal Summary}

State clearly the aim of the proposed experiment together with the general and specific scientific background.

\section{Experimental technique}

Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of an experiment should be clarified with hosts before the proposal is submitted.

\section{Safety considerations}

Experimental conditions requiring special safety precautions such as the use of lasers, high pressure cells, bacteria, dangerous substances, toxic substances and radioactive materials, must be identified in the proposal.

\section{Beamtime requested}

Specify the dates of the beamtime requested and the number of shifts. The beamtime requested should be in accordance with the estimated duration of the experiment. Specify the place and the investigators who will guide the experiment.

\section{Results expected}

Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to existing work on the topic.

\section{Concluding remarks}

Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment. Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential details.

\printbibliography
\nocite{*}

\end{multicols}

\end{document}

### 2023-11-30 text/LogBooks/logbook1.tex
logbook 1

--- 2023-11-28 text/proposta/prop.tex
@@
 \title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal\vspace{-5mm}}
-\date{November 2023}
+LFEUI - Logbook 1\vspace{-5mm}}
+\date{30 November 2023}
 
@@
 
-\begin{multicols}{2}
+% \begin{multicols}{2}
 
-% \tableofcontents
+\vspace*{1cm}
 
-\section{Proposal Summary}
+On the 30\textsuperscript{th} of November, the four members of our group went to the second meeting with professors Rodrigo Mateus, Rui \_\_ and Norberto \_\_ at CTN Loures.
+Our goal was to discuss the work we will be doing in the following session at the same laboratory.
+We arrived at 9:30 AM and started discussing the questions that we had been requested to answer the previous session.
 
-State clearly the aim of the proposed experiment together with the general and specific scientific background.
+\paragraph*{}
 
-\section{Experimental technique}
+The professors turned on the accelerator and went through the different components of its layout and experiment setup.
+We focused on the detector part and we were able to see both the samples and the silicon detectors next to it.
 
-Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of an experiment should be clarified with hosts before the proposal is submitted.
+\paragraph*{}
 
-\section{Safety considerations}
+The rest of the session was followed by a discussion about multiple theoretical and experimental aspects of the project. We discussed the following topics:
 
-Experimental conditions requiring special safety precautions such as the use of lasers, high pressure cells, bacteria, dangerous substances, toxic substances and radioactive materials, must be identified in the proposal.
+\begin{enumerate}
+  \item The structure of the tandem 2.5 MeV accelerator and both it's acceleration tubes.
+They allow the proton beam to reach the target with a kinetic energy of 5 MeV instead of 2.5 MeV (only one tube);
+  \item The radiation at the laboratory - we were told that the radiation levels are very low unless we were very close to the accelerator.
+  \item The vacuum inside the tubes - we estimated that, at a pressure of $10^{-7}$ mbar (the pressure inside the tubes), there are approximately $10^9$ particles per cubic centimeter.
+This is a good vacuum meaning that the protons won't collide with other particles and will be able to reach the target without losing energy;
+  \item The equation that we aim to study: a proton collides with $^7$Li and produces $^8$Be, which then decays into two alpha particles.
+The decay process is not detectable, so we will only be able to detect the alpha particles:
 
-\section{Beamtime requested}
+  \begin{equation}
+    \ce{^7Li + ^1H -> ^8Be -> 2\alpha}
+  \end{equation}
 
-Specify the dates of the beamtime requested and the number of shifts.
-The beamtime requested should be in accordance with the estimated duration of the experiment.
-Specify the place and the investigators who will guide the experiment.
+  \item The importance of the electromagnets used to deviate the trajectory of the proton beam.
+We understood that deflection angle is related to the mass and charge of the particle, therefore different particles will have different trajectories, meaning we can isolate the protons on the incident beam;
+  \item Samples used: in the next session, we are going to work with need to collect the radiation spectrum of two different Lithium-7 samples: one of them is glass sample composed of silicon, boron and lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate;
+  \item Finally, we discussed the expected results for both samples - phenomena such as Rutherford Backscattering, Elastic Scattering and profile of the energy spectrum of the alpha particles were mentioned.
+\end{enumerate}
 
-\section{Results expected}
+We left the laboratory at 12:00 PM.
 
-Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to existing work on the topic.
-
-\section{Concluding remarks}
-
-Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment.
-Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential details.
-
-\printbibliography
-\nocite{*}
-
-\end{multicols}
+% \end{multicols}
 

### 2023-11-30 text/notas/notas.tex
logbook 1

% RUN in terminal (without bibliography):
% pdflatex -output-directory=/Users/salvatorpes/Desktop/LFEUI/text/proposta /Users/salvatorpes/Desktop/LFEUI/text/proposta/prop.tex

\documentclass{article}

\author{
    \begin{tabular}{rl}
        Estêvão Gomes (ist1102650) & Sofia Nunes (ist1102633) \\
        Pedro Curvo (ist1102716) & Salvador Torpes (ist1102474)
    \end{tabular}
}

\usepackage[utf8]{inputenc}
\usepackage[portuguese]{babel}
\usepackage[letterpaper,top=10mm,bottom=15mm,left=10mm,right=10mm,marginparwidth=1.75cm]{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{url}
\usepackage{csquotes}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{textcomp}
\usepackage{pdflscape}
\usepackage{makeidx}
% \usepackage{tocbibind}
\providecommand{\tightlist}{\relax}
\usepackage{tocloft}
\renewcommand{\cftsecindent}{0em}
\renewcommand{\cftsubsecindent}{1em}
\renewcommand{\cftsecfont}{\bfseries}
\renewcommand{\cftsubsecfont}{\itshape}
\setlength{\cftsubsecnumwidth}{0em}

\usepackage[version=4]{mhchem}
\usepackage{hyperref} % Remove "pdftex" option here
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{ragged2e}
\usepackage{xkeyval}
%\usepackage{minted}
%\usemintedstyle{manni}
\usepackage{listings}
\usepackage{amssymb}

\usepackage{xcolor}
\usepackage{tikz}

\usetikzlibrary{positioning}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{adjustbox}
\usepackage{sidecap}
\usepackage{graphicx}

\usepackage{tikz-3dplot}
\usepackage{pgfplots}
\usetikzlibrary{calc, 3d, arrows}

\usetikzlibrary{shapes.geometric, arrows}

\lstset{
    language=Python,
    basicstyle=\ttfamily,
    keywordstyle=\color{blue},
    commentstyle=\color{gray},
    stringstyle=\color{orange},
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    breaklines=true,
    frame=tb,
    framexleftmargin=2em,
    xleftmargin=2em,
}

%\usepackage{fontspec}

%\setmonofont{Fira Code}

\fancyhf{}
\cfoot{\thepage}
\fancyhf{} % Clear all header and footer fields
\renewcommand{\headrulewidth}{0pt} % Remove the header rule line
\cfoot{\thepage} % Set the page number in the center of the footer

\pagestyle{fancy} % Apply the fancy page style

\setlength\columnsep{20pt}

\renewcommand{\familydefault}{\sfdefault}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\makeatletter
\newenvironment{figurehere}
{\def\@captype{figure}}
{}
\makeatother

\hypersetup{
  colorlinks,
  linkcolor=blue,
  anchorcolor=black,
  citecolor=cyan,
  filecolor=cyan,
  menucolor=cyan,
  urlcolor=cyan,
  bookmarksopen=true,
  bookmarksnumbered=true
}

\makeindex

\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
LFEUI - Notas Teóricas \vspace{-5mm}}
\date{30 Novembro 2023}

\usepackage{sansmathfonts}
\usepackage[T1]{fontenc}
\usepackage[OT1]{fontenc}

\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}

\usepackage[style=numeric]{biblatex} % Choose your desired citation style
\addbibresource{../references/prop.bib} % Specify your .bib file

\begin{document}

\renewcommand{\arraystretch}{1.5}
\setlength{\columnseprule}{0.4pt}
\tdplotsetmaincoords{70}{110} % Set the viewing angle
\newcolumntype{M}[1]{>{\centering\arraybackslash\vspace{#1}}m{0.5\linewidth}<{\vspace{#1}}}
\newcolumntype{C}[2]{>{\centering\arraybackslash\vspace{#1}\rule{0pt}{#1}\hspace{0pt}}m{#2}}
\newcolumntype{w}[1]{>{\centering\arraybackslash}m{#1}}

\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif

\maketitle

\vspace{-5mm}

\hrulefill

\begin{multicols}{2}

\section{Acelerador de Partículas}

O acelerador utilizado possui no seu interior um gás.

\section{Lítio}

Vamos trabalhar com o lítio, um elemento altamente reativo, maioritariamente através de reações de oxidação com o vapor de água. 
A solução para trabalhar com este elemento é usar fontes inertes - uma fonte inerte de lítio consiste em lítio no interior de duas camadas de vidro.
O vidro é composto por silicatos de boro.

\section{Objetivo}

O objetivo é obter o espetro de emissão de duas fontes diferentes, ambas contendo lítio.

\section{Reação}

A reação que pretendemos estudar envolve o choque de um átomo de lítio 7 com um protão, que resulta na formação de um átomo de berílio 8 que decai para duas partículas alfa (núcleo de hélio 4) - o decaimento é tão rápido que não é detetado.

\begin{equation}
    \ce{^7Li + ^1H -> ^8Be -> 2\alpha = 2^4He}
\end{equation}

Esta reação é idêntica para as duas amostras que vamos estudar.

\subsection*{Valor $Q$ da reação}

O valor $Q$ da reação é a energia libertada na reação - é a energia que as partículas alfa possuem quando são emitidas.
Calcula-se a partir da diferença de massa entre os reagentes e os produtos da reação:

\begin{equation}
  \begin{split}
      Q &= \Delta m c^2 \\
      &= c^2 \left( m_{\text{reagents}} - m_{\text{products}} \right) \\
      &= c^2 \left( m_{\ce{^7Li}} + m_{\ce{^1H}} - m_{\ce{^8Be}} \right)
  \end{split}
\end{equation}

\section{Processo Experimental}

Queremos obter o espetro energético para perceber qual a energia com que as partículas são emitidas

\section{Montagem Experimental}

Temos um acelerador Tandem de 2.5 MeV, que acelera protões - temos 2 tubos de aceleração em paralelo o que faz com que a tensão à qual os partículas protões são aceleradas possa chegar aos 5 MV. Os dois tubos estão alinhados o que faz com que o protão seja 'duplamente' acelerado.
Acelerar um protão a 5 MeV significa que a variação de energia cinética do eletrão desde o íncio do tubo (em repouso) até ao seu fim é de 5 MeV - este é o valor da diferença de potencial elétrico entre as duas extremidades/terminais do tubo de aceleração. 
Este acelerador possui um elevador de tensão - associação de condensadores e resistências que permitem aumentar a tensão.

\paragraph*{Vácuo} O feixe propaga-se ao longo dos dois tubos de aceleração que se encontra sobre um vácuo de $10^{-7}$ mbar. O vácuo é necessário para que as partículas não colidam com moléculas de ar e percam energia.
Este vácuo é bastante bom sendo possível que existam colisão no caminho apesar de em média o número destas colisões é nulo.
A este nível de pressão ($10^{-7}$ mbar), existem cerca de $10^9$ moléculas de ar por cm$^3$. Na atmosfera normal (pressão de 1 bar) existem $10^{19}$ moléculas de ar por cm$^3$ - no vácuo intergalático temos 5 moléculas por cm$^3$ - apesar de o vácuo deste vácuo não ser comparável ao do espaço, é um vácuo bastante bom. 
Para estimar estas quantidades de partículas por cm$^3$ consideramos que a diferença de pressão entre a atmosfera normal e o vácuo e consideramos que o volume de um mol de gás é $22.4$ L/mol na atmosfera normal - isto implica que há uma mol de moléculas por litro - $6.02 \times 10^{23}$ moléculas por litro.

\paragraph*{Eletroímane} No nosso eletroímane, partículas com diferentes racios carga/massa têm ângulos de deflexão diferentes - o ângulo de deflexão é inversamente proporcional ao racio carga/massa. Este elemento é responsável por fazer com que todas as partículas que não os protões que se encontrem no feixe sejam desviadas para fora do feixe e por isso não atingem o sensor. É por isto que existe um grande benefício em ter um detetor não linear - fitração do feixe de partículas.

\paragraph*{Detetores} Vamos usar detetores de silício - a energia de uma partícula incidente no detetor é calculada a partir do número de pares electrão-buraco que são criados no detetor devido à sua passagem.

\paragraph*{Eletrónica} Existe um sistema de eletrónica digital responsável por fazer a leitura dos detetores e por enviar os sinais para o computador onde podemos observar os espetros obtidos.

\section{Amostras}

A atividade envolve o estudo dos espetros provenientes de duas amostras diferentes, ambas contendo lítio - uma amostra que é apenas um vidro composto por silicatos e óxidos e outra que é uma amostra com lítio implantado em silício.

\paragraph*{Implantação} A implantação é o processo de formacação de amostras implantadas - para obter uma amostra de lítio 7 implantado em silício usou-se um feixe de lítio 7 com potência na ordem dos kilovolts - quanto maior a potência do feixe, maior é a profundidade a que o lítio é implantado no silício. 
Existe um fórmula que nos permite saber a profundidade a que o lítio é implantado no silício em função da potência do feixe de lítio 7 e da densidade do silício.

\paragraph*{Diferenças} O objetivo de estudar estas duas amostras é avaliar a diferença de espetros de emissão de partículas alfa provenientes de lítio 7 em diferentes profundidades.

\begin{table}[H]
\centering
\begin{tabular}{|w{0.2\linewidth}|w{0.7\linewidth}|}
\hline
Amostra & Descrição \\ \hline
Vidro Normal & Amostra formada por vidro normal com silício, lítio, boro e talvez outros resíduos - o lítio aparece na forma de óxido de lítio de modo a que seja pouco reativo. Esta amaostra é homogénea na sua constituição ao longo de toda a sua espessura uma vez que o vidro é formado diretamente com os elementos. \\ \hline
Implantada & Amostra obtida através da implatação de uma camada de lítio numa amostra de sílicio (pura) - o que obtemos é uma amostra que nos primeiros nanómetrod e espessura possui lítio e depois apenas possui silício. \\ \hline
\end{tabular}
\end{table}

\section{Resultados Esperados}

\paragraph*{Perfis} Um perfil é uma parte contínua do espetro na zona energética anterior aos picos de partículas alfa.
Esperamos ver isto na amostra de vidro devido às partículas alfa que provém de diferentes profundidades da amostra de vidro.
Quanto maior for a profundidade, como estamos a falar de partículas carregadas e não de fotões, maior é a energia dissipada para o material durante o percurso. 
Para partículas emitidas com a mesma energia, quanto maior for a profundidade, menor é a energia com que a partícula é detetada - cria-se um perfil no espetro devido a isto. 
Como na amostra implantada o lítio está apenas presente numa camada superior muito fina, não esperamos obter um perfil no espetro.

\paragraph*{RBS - Rutherford Backscattering} é um fenómeno no qual as partículas alfa são retrodispaersas para outras direções e por isso possuem energias diferentes. 
A maior parte das partículas alfa são detetadas com o valor de energia esperada mas podemos ter outras zonas do espetro onde tenhamos partículas alfa com energias diferentes e dispersas devido a este fenómeno.

\paragraph*{Elastic Backscattering} Isto acontece quando um dos protões do feixe incidente é refletido elasticamente na superfície da amostra - o protão é refletido com a mesma energia com que incidiu na amostra e continua a trajetória na mesma direção mas com sentido oposto.

\section{Utilidade}

\begin{enumerate}
  \item Fusão Nuclear - lítio 6 para produção de neutrões -  usam-se materiais muito enriquecidos em lítio 6;
  \item Baterias - lítio 7;
  \item Indústria Farmaceûtica - lítio é utilizado em vários medicamentos;
\end{enumerate}

\printbibliography
\nocite{*}

\end{multicols}

\end{document}

### 2023-12-02 text/LogBooks/logbook1.tex
logbook 1

@@
 
-On the 30\textsuperscript{th} of November, the four members of our group went to the second meeting with professors Rodrigo Mateus, Rui \_\_ and Norberto \_\_ at CTN Loures.
+On the 30\textsuperscript{th} of November, the four members of our group went to the second meeting with professors Rodrigo Mateus, Rui Silva and Norberto Catarino at CTN Loures.
 Our goal was to discuss the work we will be doing in the following session at the same laboratory.
@@
   \item The vacuum inside the tubes - we estimated that, at a pressure of $10^{-7}$ mbar (the pressure inside the tubes), there are approximately $10^9$ particles per cubic centimeter.
-This is a good vacuum meaning that the protons won't collide with other particles and will be able to reach the target without losing energy;
+This is a good vacuum meaning that the protons won't collide with other particles and will be able to reach the target without losing energy because the mean free path is longer than the accelerator;
   \item The equation that we aim to study: a proton collides with $^7$Li and produces $^8$Be, which then decays into two alpha particles.
@@
 We understood that deflection angle is related to the mass and charge of the particle, therefore different particles will have different trajectories, meaning we can isolate the protons on the incident beam;
-  \item Samples used: in the next session, we are going to work with need to collect the radiation spectrum of two different Lithium-7 samples: one of them is glass sample composed of silicon, boron and lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate;
+  \item Samples used: in the next session, we need to collect the radiation spectrum of two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate;
   \item Finally, we discussed the expected results for both samples - phenomena such as Rutherford Backscattering, Elastic Scattering and profile of the energy spectrum of the alpha particles were mentioned.

### 2023-12-02 text/proposta/prop.tex
logbook 1

@@
 \title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal\vspace{-5mm}}
+LFEUI - Investigation Proposal \\ \vspace{4mm} {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-1mm}}
 \date{November 2023}
@@
 
+\paragraph*{}
+
+Our main goal is to develop a scientific research project alongside three senior investigators from CTN (Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
+
+Our project will be focused on the detection of Li in technological materials through nuclear reactions.
+The main motivation for this project is to study and develop a detection technique for elements such as Li and Br with high sensitivity.
+
+Through the year there have been multiple techniques with the same purpose such as atomic spectroscopy and ?. Particularly, the one we aim to use was introduced in ?.
+
 \section{Experimental technique}
@@
 Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of an experiment should be clarified with hosts before the proposal is submitted.
+
+\paragraph{Set-up}
+
+The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with the samples and the detectors.
+The accelerator produces a beam of protons with a kinetic energy of 5 MeV which is deviated by the electromagnet and hits the samples. 
+The electromagnets deflect particles with different angles according to their charge-to-mass ratio.
+This allows us to remove any unwanted particles from the beam and only let the protons pass through.
+The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.
+
+In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we can see the results.
+We aim to analyze the energy spectrum resulting from the reaction between the protons and the samples.
+These spectra will unveil the presence of Li in the samples.
+
+\paragraph{Samples}
+
+We are going to work with two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate.
+
+\paragraph{Requirements}
 
@@
 
+\paragraph*{}
+
+In order to complete our project we will need one full day of beamtime at the CTN laboratory.
+Alongside this day, we will also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the results after the experiment:
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|w{0.45\linewidth}|w{0.3\linewidth}|w{0.15\linewidth}|}
+\hline
+Phase & Date & Duration \\ \hline
+Theoretical Background, Planning and Lab Overview & 30/11/2023 9h30 & 4h \\ \hline
+Experiment & 7/12/2023 8h00 & 8h \\ \hline
+Results Discussion & ?? 9h30 & 4h \\ \hline
+\end{tabular}
+\end{table}
+
 \section{Results expected}
@@
 Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to existing work on the topic.
+
+\paragraph{}
 

### 2023-12-02 text/proposta/prop.tex
logbook 1

@@
 
+We expect to obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the samples. 
+For the glass samples, we expect to see a profile due to the particles that come from different depths of the sample and therefore have lost different amounts of energy.
+In the implant samples, we expect to see a peak at the energy corresponding to the energy of the alpha particles produced by the reaction.
+In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged particles.
+
+\paragraph{}
+
+Other phenomena such as Rutherford Backscattering and Elastic Scattering are also expected to occur.
+
 \section{Concluding remarks}

### 2023-12-02 text/proposta/prop.tex
logbook 1

@@
 
-\vspace{-5mm}
+\vspace{-8mm}
+
+\hrulefill
+
+\begin{center}
+  \textbf{\Large Abstract}
+\end{center}
+
+% \begin{abstract}
+%   \par textvtgerthrt
+% \end{abstract}
+
+\par In the following document we present a proposal for a scientific research being developed
+
+\vspace{4.0mm}
 
@@
 
-\paragraph*{}
+\paragraph{Goal}
 
-Our main goal is to develop a scientific research project alongside three senior investigators from CTN (Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
+Our main goal is to develop a scientific research project alongside three senior investigators from CTN
+(Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
 
 Our project will be focused on the detection of Li in technological materials through nuclear reactions.
-The main motivation for this project is to study and develop a detection technique for elements such as Li and Br with high sensitivity.
+The main motivation for this project is to study and develop a detection technique for elements such as Li with
+high sensitivity.
 
-Through the year there have been multiple techniques with the same purpose such as atomic spectroscopy and ?. Particularly, the one we aim to use was introduced in ?.
+\paragraph{Scientific Background}
+
+Through the year there have been multiple techniques with the same purpose such as atomic spectroscopy and ?.
+Particularly, the one we aim to use was introduced in ?.
 
@@
 
-Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of an experiment should be clarified with hosts before the proposal is submitted.
+Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of
+an experiment should be clarified with hosts before the proposal is submitted.
 
@@
 
-The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with the samples and the detectors.
-The accelerator produces a beam of protons with a kinetic energy of 5 MeV which is deviated by the electromagnet and hits the samples. 
+The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with the samples
+and the detectors.
+The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet
+and hits the samples. 
 The electromagnets deflect particles with different angles according to their charge-to-mass ratio.
 This allows us to remove any unwanted particles from the beam and only let the protons pass through.
-The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.
+The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the
+silicon.
 
-In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we can see the results.
+In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we
+can see the results.
 We aim to analyze the energy spectrum resulting from the reaction between the protons and the samples.
@@
 
-We are going to work with two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate.
+We are going to work with two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and
+lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate.
 
@@
 
+\paragraph{WorkPlan}
+
+Obtaining the energy spectrum of the alpha particles resulting from the reaction between the protons and the samples will
+evolve the following steps:
+
+\begin{enumerate}
+  \item Accelerator Startup;
+  \item Calibration of the detectors with a sample of known composition;
+  \item Obtaining the energy spectrums of both samples;
+  \item Analysis of the results.
+\end{enumerate}
+
+\paragraph*{Beam Energy}
+
+In order to choose the energy of the beam we need to take into account the cross section of the reaction between the the protons and the Li-7. We
+
+\begin{center}
+  \includegraphics[width=0.95\linewidth]{../images/Li_crosssection_energy.jpeg}
+  \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy.}
+\end{center}
+
 \section{Safety considerations}
 
-Experimental conditions requiring special safety precautions such as the use of lasers, high pressure cells, bacteria, dangerous substances, toxic substances and radioactive materials, must be identified in the proposal.
+During the experiments, the accelerator will be producing a beam of protons with a kinetic energy of up to 5 MeV. Because of this there will be a high radiation level near the accelerator.
+There are radiation maps that show the radiation levels in different areas of the laboratory.
 
@@
 
-Specify the dates of the beamtime requested and the number of shifts.
-The beamtime requested should be in accordance with the estimated duration of the experiment.
-Specify the place and the investigators who will guide the experiment.
+Specify the dates of the beamtime requested and the number of shifts. The beamtime requested should be in accordance with
+the estimated duration of the experiment. Specify the place and the investigators who will guide the experiment.
 
@@
 
-In order to complete our project we will need one full day of beamtime at the CTN laboratory.
-Alongside this day, we will also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the results after the experiment:
+In order to complete our project we will need one full day of beamtime at the CTN laboratory. Alongside this day, we will
+also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the
+results after the experiment:
 
@@
 \centering
-\begin{tabular}{|w{0.45\linewidth}|w{0.3\linewidth}|w{0.15\linewidth}|}
+\begin{tabular}{|w{0.4\linewidth}|w{0.3\linewidth}|w{0.15\linewidth}|}
 \hline
@@
 
-Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to existing work on the topic.
+Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to
+existing work on the topic.
 
@@
 
-We expect to obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the samples. 
-For the glass samples, we expect to see a profile due to the particles that come from different depths of the sample and therefore have lost different amounts of energy.
-In the implant samples, we expect to see a peak at the energy corresponding to the energy of the alpha particles produced by the reaction.
-In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged particles.
+We expect to obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the
+samples. 
+For the glass samples, we expect to see a profile due to the particles that come from different depths of the sample and
+therefore have lost different amounts of energy. In the implant samples, we expect to see a peak at the energy corresponding
+to the energy of the alpha particles produced by the reaction.
+In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged
+particles.
 
@@
 Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment.
-Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential details.
+Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential
+details.
 

### 2023-12-02 text/proposta/prop.tex
beam energy

@@
   \item Calibration of the detectors with a sample of known composition;
-  \item Obtaining the energy spectrums of both samples;
+  \item Obtaining the energy spectrums of both samples with the appropriate beam energy;
   \item Analysis of the results.
@@
 
-In order to choose the energy of the beam we need to take into account the cross section of the reaction between the the protons and the Li-7. We
+In order to choose the energy of the beam we need to take into account the cross section of the reaction between the the protons and the Li-7.
+We collected data from the IBANDL \cite{IAEA_EXFOR} database for this reaction in a range of $[0.5,7]$ MeV and plotted it in the following graph:
 
@@
   \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy.}
+  \label{fig:crosssection}
 \end{center}
+
+We are only interested in the energy range of $[0.5,5]$ MeV because the energy of the beam is limited by the accelerator.
+As we can see in figure \ref{fig:crosssection}, the cross section of the reaction is maximum in the $[2.5,3]$ MeV range. 
+Therefore, since we want to detect the Li-7 in the samples, we should choose the energy of the beam to be in this range.
 

### 2023-12-02 text/proposta/prop.tex
beamtime requested

@@
 Experiment & 7/12/2023 8h00 & 8h \\ \hline
-Results Discussion & ?? 9h30 & 4h \\ \hline
+Results Discussion & To be arranged & 4h \\ \hline
 \end{tabular}

### 2023-12-02 text/proposta/prop.tex
vários

@@
 
+\setlist[enumerate]{itemsep=3pt}
+
 \begin{document}
@@
 
-\par In the following document we present a proposal for a scientific research being developed
+\par In the following document we present a proposal for a scientific research being developed at CTN Lisbon. We will...
 
@@
 \section{Proposal Summary}
-
-State clearly the aim of the proposed experiment together with the general and specific scientific background.
 
@@
 
-Through the year there have been multiple techniques with the same purpose such as atomic spectroscopy and ?.
-Particularly, the one we aim to use was introduced in ?.
+This detection method is based on the nuclear reaction between the element we want to detect ($\ce{^7Li}$) and a proton, as follows:
+
+\begin{equation}
+  \ce{^7Li + ^1H -> ^8Be -> 2\alpha}
+  \label{eq:reaction}
+\end{equation}
+
+The reaction results in the production of a $\ce{^8Be}$ nucleus which then rapidly decays into two alpha particles. These are the particles we detect.
+
+According to the incident proton's energy and the scattering angle of the detectors ($\theta = 165 $ degrees) we were able to use the NRA Calculator \cite{NRAEnergyCalc} to calculate the energy of the alpha particles resulting from the reaction as well as it's $Q$ value:
+
+\begin{equation}
+\begin{split}
+  E_{\alpha_1} &= 7.65320 \text{ MeV} \\
+  E_{\alpha_2} &= 10.89305 \text{ MeV} \\
+  Q &= 17.34625 \text{ MeV}
+\end{split}
+\label{eq:energies}
+\end{equation}
+
+The formula this website uses to compute the energies is the one on the book \cite{krane1988}.
 
@@
 
-Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of
-an experiment should be clarified with hosts before the proposal is submitted.
-
 \paragraph{Set-up}
 
-The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with the samples
-and the detectors.
-The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet
-and hits the samples. 
-The electromagnets deflect particles with different angles according to their charge-to-mass ratio.
-This allows us to remove any unwanted particles from the beam and only let the protons pass through.
-The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the
-silicon.
+The experimental set-up consists of a Tandem accelerator, a beam line with an electromagnet and a chamber with both the samples and the detectors.
+The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet and hits the sample. 
+The electromagnets deflect particles with different angles according to their charge-to-mass ratio and by that allow us to remove any unwanted particles from the beam and only let the protons pass through.
+The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.
 
@@
 can see the results.
-We aim to analyze the energy spectrum resulting from the reaction between the protons and the samples.
-These spectra will unveil the presence of Li in the samples.
+
+The beam line is under a vacuum of around $10^{-7}$ mbar in order to avoid the protons colliding with other particles and losing energy. 
+They reach the sample with the specified kinetic energy and cause a nuclear reaction whose products are the charged particles detected.
+We aim to analyze the energy spectrum of these particles.
+
+The following diagram shows the experimental set-up:
+
+\begin{center}
+  \includegraphics[width=0.95\linewidth]{../images/scheme.jpeg}
+  \captionof{figure}{Experimental set-up.}
+  \label{fig:setup}
+\end{center}
 
@@
 lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate.
-
-\paragraph{Requirements}
 
@@
 
-Specify the dates of the beamtime requested and the number of shifts. The beamtime requested should be in accordance with
-the estimated duration of the experiment. Specify the place and the investigators who will guide the experiment.
-
-\paragraph*{}
-
-In order to complete our project we will need one full day of beamtime at the CTN laboratory. Alongside this day, we will
+In order to complete our project we will need one full day of beamtime at the CTN laboratory \cite{IST}. Alongside this day, we will
 also meet with the investigators at CTN for another two mornings to discuss the project, before the experiment and the
@@
 
-Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to
-existing work on the topic.
-
-\paragraph{}
-
-We expect to obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the
+We will obtain the energy spectrum of the alpha particles resulting from the reaction between the protons and the
 samples. 
@@
 
+\paragraph*{Energy Values} The alpha particles are supposed to have the energies calculated in equation \ref{eq:energies} ($E_{\alpha_1}$ and $E_{\alpha_2}$). 
+In the glass sample, we expect to see an energy profile before each peak and in the implant sample we only expect to see the peaks.
+
 \paragraph{}
@@
 Other phenomena such as Rutherford Backscattering and Elastic Scattering are also expected to occur.
+These are responsible for energy peaks in both sample's spectra at lower than the expected energy values.
 
@@
 
-Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment.
-Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential
-details.
+By identifying the presence of light isotopes at low energies ($<5$MeV), we achieve a practical method to evaluate the presence and the amount of these light isotopes in technological materials.
+Until now, the only way to do this was through atomic spectroscopy, which, besides not being able to detect the presence of light isotopes, is a much more complex and expensive method.
 

### 2023-12-02 text/proposta/prop.tex
Proposal Update

@@
 
-The formula this website uses to compute the energies is the one on the book \cite{krane1988}.
+The formula this website uses to compute the energies is on the page 382 of the book \cite{krane1988}.
 
@@
 The accelerator produces a beam of protons with a kinetic energy of up to 5 MeV which is deviated by the electromagnet and hits the sample. 
-The electromagnets deflect particles with different angles according to their charge-to-mass ratio and by that allow us to remove any unwanted particles from the beam and only let the protons pass through.
+The electromagnets deflect particles with different angles according to their charge-to-mass ratio and, with that, allows us to remove any unwanted particles from the beam and only let the protons pass through.
 The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.
 
-In addition, their are connected to a electronic system that amplifies the signal and sends it to a computer where we
+In addition, they are connected to a electronic system that amplifies the signal and sends it to a computer where we
 can see the results.
@@
 The following diagram shows the experimental set-up:
+Note: We use a tandem accelerator, so the image should have 2 tubes inside the accelerator.
 
@@
 During the experiments, the accelerator will be producing a beam of protons with a kinetic energy of up to 5 MeV. Because of this there will be a high radiation level near the accelerator.
-There are radiation maps that show the radiation levels in different areas of the laboratory.
+There are radiation maps that show the radiation levels in different areas of the laboratory and need to be strictly followed.
 
@@
 to the energy of the alpha particles produced by the reaction.
-In both cases, the samples is composed by other types of atoms such as silicon and boron, which will also emit charged
-particles.
+In both cases, the samples are composed by other types of atoms such as silicon and boron, which will also emit charged
+particles. By obtaining the expected energy spectrums we can infer the presence of Li in the samples. If possible we will also try
+to quantify the amount of Li present by analysing the amout of energy deposited in the spectrums.
 

### 2023-12-02 text/proposta/prop.tex
final

@@
 \usepackage[english]{babel}
-\usepackage[letterpaper,top=10mm,bottom=15mm,left=10mm,right=10mm,marginparwidth=1.75cm]{geometry}
+\usepackage[letterpaper,top=6mm,bottom=15mm,left=6mm,right=6mm,marginparwidth=1.55cm]{geometry}
 \usepackage{multicol}
@@
 \title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal \\ \vspace{4mm} {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-1mm}}
+LFEUI - Investigation Proposal \\ \vspace{0mm} {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-0mm}}
 \date{November 2023}
@@
 
-\begin{center}
-  \textbf{\Large Abstract}
-\end{center}
+% \begin{center}
+%   \textbf{\Large Abstract}
+% \end{center}
 
-% \begin{abstract}
-%   \par textvtgerthrt
-% \end{abstract}
-
-\par In the following document we present a proposal for a scientific research being developed at CTN Lisbon. We will...
-
-\vspace{4.0mm}
+\vspace{-0.1cm}
+\begin{abstract}
+  \par In the following document we present a proposal for a scientific research being developed at CTN Lisbon in December 2023.
+  Our work consists of the detection of $\ce{^7Li}$ in two different samples through nuclear reactions.
+  This document describes the time required for the project as well as it's goals, theoretical background, experimental technique, safety considerations and expected results.
+  In addition, we also present a brief description of the experimental set-up.
+\end{abstract}
+\vspace{-0.1cm}
 
@@
 particles. By obtaining the expected energy spectrums we can infer the presence of Li in the samples. If possible we will also try
-to quantify the amount of Li present by analysing the amout of energy deposited in the spectrums.
+to quantify the amount of Li present by analyzing the amount of energy deposited in the spectrums.
 

### 2023-12-02 text/proposta/prop.tex
Proposal
final

@@
 \title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal \\ \vspace{0mm} {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-0mm}}
+LFEUI - Investigation Proposal \\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-0mm}}
 \date{November 2023}
@@
   \captionof{figure}{Experimental set-up.}
-\label{fig:setup}
+  \label{setup}
 \end{center}
@@
   \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy.}
-\label{fig:crosssection}
+  \label{crosssection}
 \end{center}
@@
 We are only interested in the energy range of $[0.5,5]$ MeV because the energy of the beam is limited by the accelerator.
-As we can see in figure \ref{fig:crosssection}, the cross section of the reaction is maximum in the $[2.5,3]$ MeV range. 
+As we can see in figure \ref{crosssection}, the cross section of the reaction is maximum in the $[2.5,3]$ MeV range. 
 Therefore, since we want to detect the Li-7 in the samples, we should choose the energy of the beam to be in this range.

### 2023-12-04 text/proposta/prop.tex
final
gsreg
Resolved Conflicts and GitIgnore

@@
 \documentclass{article}
+\input{./preamble}
 
@@
 }
-
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage[letterpaper,top=6mm,bottom=15mm,left=6mm,right=6mm,marginparwidth=1.55cm]{geometry}
-\usepackage{multicol}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{makecell}
-\usepackage{titlesec}
-\usepackage{multirow}
-\usepackage{amsmath}
-\usepackage{makecell}
-\usepackage{url}
-\usepackage{csquotes}
-\usepackage{caption}
-\usepackage{enumitem}
-\usepackage{textcomp}
-\usepackage{pdflscape}
-\usepackage{makeidx}
-% \usepackage{tocbibind}
-\providecommand{\tightlist}{\relax}
-\usepackage{tocloft}
-\renewcommand{\cftsecindent}{0em}
-\renewcommand{\cftsubsecindent}{1em}
-\renewcommand{\cftsecfont}{\bfseries}
-\renewcommand{\cftsubsecfont}{\itshape}
-\setlength{\cftsubsecnumwidth}{0em}
-
-\usepackage[version=4]{mhchem}
-\usepackage{hyperref} % Remove "pdftex" option here
-\usepackage{float}
-\usepackage{fancyhdr}
-\usepackage{ragged2e}
-\usepackage{xkeyval}
-%\usepackage{minted}
-%\usemintedstyle{manni}
-\usepackage{listings}
-\usepackage{amssymb}
-
-\usepackage{xcolor}
-\usepackage{tikz}
-
-\usetikzlibrary{positioning}
-\usetikzlibrary{positioning, arrows.meta}
-\usepackage{adjustbox}
-\usepackage{sidecap}
-\usepackage{graphicx}
-
-\usepackage{tikz-3dplot}
-\usepackage{pgfplots}
-\usetikzlibrary{calc, 3d, arrows}
-
-\usetikzlibrary{shapes.geometric, arrows}
-
-\lstset{
-    language=Python,
-    basicstyle=\ttfamily,
-    keywordstyle=\color{blue},
-    commentstyle=\color{gray},
-    stringstyle=\color{orange},
-    numbers=left,
-    numberstyle=\tiny,
-    numbersep=5pt,
-    showspaces=false,
-    showstringspaces=false,
-    breaklines=true,
-    frame=tb,
-    framexleftmargin=2em,
-    xleftmargin=2em,
-}
-
-%\usepackage{fontspec}
-
-%\setmonofont{Fira Code}
-
-\fancyhf{}
-\cfoot{\thepage}
-\fancyhf{} % Clear all header and footer fields
-\renewcommand{\headrulewidth}{0pt} % Remove the header rule line
-\cfoot{\thepage} % Set the page number in the center of the footer
-
-\pagestyle{fancy} % Apply the fancy page style
-
-\setlength\columnsep{20pt}
-
-\renewcommand{\familydefault}{\sfdefault}
-
-\newenvironment{Figure}
-  {\par\medskip\noindent\minipage{\linewidth}}
-  {\endminipage\par\medskip}
-
-\makeatletter
-\newenvironment{figurehere}
-{\def\@captype{figure}}
-{}
-\makeatother
-
-\hypersetup{
-  colorlinks,
-  linkcolor=blue,
-  anchorcolor=black,
-  citecolor=cyan,
-  filecolor=cyan,
-  menucolor=cyan,
-  urlcolor=cyan,
-  bookmarksopen=true,
-  bookmarksnumbered=true
-}
-
-\makeindex
-
-\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal \\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-0mm}}
-\date{November 2023}
-
-\usepackage{sansmathfonts}
-\usepackage[T1]{fontenc}
-\usepackage[OT1]{fontenc}
-
-\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
-
-\usepackage[style=numeric]{biblatex} % Choose your desired citation style
-\addbibresource{../references/prop.bib} % Specify your .bib file
-
-\setlist[enumerate]{itemsep=3pt}
 

### 2023-12-07 text/LogBooks/logbook2.tex
gtrg

--- 2023-11-28 text/proposta/prop.tex
@@
 
-\begin{multicols}{2}
-
-% \tableofcontents
-
-\section{Proposal Summary}
-
-State clearly the aim of the proposed experiment together with the general and specific scientific background.
-
-\section{Experimental technique}
-
-Describe the experimental set-up and requirements. Queries concerning the feasibility (technical or safety aspects) of an experiment should be clarified with hosts before the proposal is submitted.
-
-\section{Safety considerations}
-
-Experimental conditions requiring special safety precautions such as the use of lasers, high pressure cells, bacteria, dangerous substances, toxic substances and radioactive materials, must be identified in the proposal.
-
-\section{Beamtime requested}
-
-Specify the dates of the beamtime requested and the number of shifts.
-The beamtime requested should be in accordance with the estimated duration of the experiment.
-Specify the place and the investigators who will guide the experiment.
-
-\section{Results expected}
-
-Describe the results expected from the measurements, their scientific (or technical) relevance, and how they relate to existing work on the topic.
-
-\section{Concluding remarks}
-
-Justify the beam time you request and give technical reasons which make the apparatus necessary for your experiment.
-Include REFERENCES where these are relevant, but do not assume that they will be systematically consulted for essential details.
-
-\printbibliography
-\nocite{*}
-
-\end{multicols}
-
 \end{document}

### 2023-12-07 text/LogBooks/logbook2.tex
rgegre

@@
 % pdflatex -output-directory=/Users/salvatorpes/Desktop/LFEUI/text/proposta /Users/salvatorpes/Desktop/LFEUI/text/proposta/prop.tex
+
+% Limpar:
+% latexmk -c
 
@@
 
+We started by discussing the samples that we wil be using during the day.
+In the previous session we had already concluded that there would be two different types of samples: glass samples with lithium in their composition and also implanted samples.
+The later consists of a thin layer of lithium implanted in a silicon substrate.
+As for the glass samples, they are not available and we will simply use three different samples, each one with a different substance/material whose composition is known and varies from sample to sample.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|w{2cm}|w{2cm}|w{2cm}|}
+\hline
+Samples & Description & Composition \\ \hline
+1 & Lithium battery & 2.5\% Li \\ \hline
+2 & Lithium fluoride & $\ce{LiF}$ - 50\% Li \\ \hline
+3 & Lithium aluminate & $\ce{LiAlO2}$ - 25\% Li \\ \hline
+4 & Implanted sample & $\ce{Li}$ implanted in $\ce{Si}$ \\ \hline
+\end{tabular}
+\caption{Samples used in the experiment}
+\label{tab:samples}
+\end{table}
+
 \end{document}

### 2023-12-07 text/notas/notas.tex
gsreg
rgegre

@@
 % pdflatex -output-directory=/Users/salvatorpes/Desktop/LFEUI/text/proposta /Users/salvatorpes/Desktop/LFEUI/text/proposta/prop.tex
+
+% limpar:
+% latexmk -c
 
@@
 
+\section{Fenómenos}
+
+Existem 3 fenómenos diferentes que acontecem nas nossas amostras quando os seus atómos colidem com os protões do feixe: reação nuclear, RBS e elastic backscattering.
+Cada um destes 3 fenómenos possui uma secção eficaz diferente - a secção eficaz é uma medida da probabilidade de um dado fenómeno acontecer.
+A secção eficaz do RBS e do elastic backscattering são maiores do que a secção eficaz da reação nuclear - isto significa que a probabilidade de um protão colidir com um átomo da amostra e ser refletido é maior do que a probabilidade de um protão colidir com um átomo da amostra e reagir com ele.
+Nos espetros detetados podemos ver uma 'bossa' inicial que corresponde ao RBS e ao elastic backscattering e depois vemos um pico que corresponde à reação nuclear do lítio 7 com o protão.
+A seguinte fórmula é válida para a geometria da nossa experiência, sendo apenas necessário aplicá-la para cada uma das 3 secções eficazes:
+
+\[ A = \left( \frac{q_{\text{beam}}}{e} \right) \Omega \overline{\sigma} (N \Delta x) \]
+
+\section{Pico de Reação Nuclear}
+
+Existem 2 picos teóricos esperados no espetro devido à reação nuclear entre o lítio e o protão.
+O feixe que utilizamos na atividade experimental possui uma energia de 1.3 MeV - a zona de energia dos protões que maximiza a intensidade deste pico é de $[2.5,3]$ MeV, no entanto, não conseguimos alcançar estas energias com o nosso acelerador (apesar de teoricamente serem possíveis).
+Para além disto estamos também interessados em detetar uma possível presença de boro nas amostras e o boro não reage com protões de tão elevada energia - assim escohemos uma energia onde tanto o lítio como o boro reagem com os protões: 1.3 MeV.
+
+\paragraph*{Cálculo do Valor}
+
+Recorremos ao NRA calculator para calcular o valor dos picos de reação nuclear para o lítio 7 e para o boro 11.
+Usamos 1.3 MeV para os protões incidentes, 165 graus para o ângulo de scattering do detetor e obtivemos os seguintes valores:
+
+\begin{equation}
+\begin{split}
+  E_{\alpha_{0_1}} &= 7.6298 \text{ MeV} \\
+  E_{\alpha_{0_2}} &= 11.0314 \text{ MeV} \\
+  \theta_{\alpha_{0_1}} &= 165 \\
+  \theta_{\alpha_{0_2}} &= 12 \\
+  Q (\ce{^7Li}) &= 17.346 \text{ MeV} \\
+\end{split}
+\label{eq:energiesli}
+\end{equation}
+
+Para o lítio não temos partículas $\alpha_1$ uma vez que: $\alpha_0$ são partículas alfa no estado fundamental e $\alpha_1$ são partículas alfa no estado excitado - a energia que temos no feixe não é suficiente para excitar as partículas alfa e por isso elas apenas são emitidas no estado fundamental.
+Os átomos de hélio 4 são ultra estáveis e o primeiro estado excitado está muito acima do estado fundamental - a energia necessária para excitar as partículas alfa é muito elevada e por isso não temos partículas alfa no estado excitado.
+\paragraph{} Note-se também que o ângulo de scattering para a segunda partícula alfa $\alpha_{0_2}$ é de 12 graus, ângulo no qual não há detetor, logo, apenas vamos detetar um pico de partículas em 7.6298 MeV.
+
+\paragraph*{} Para o boro temos:
+
+\begin{equation}
+\begin{split}
+  E_{\alpha_0} &= 3.7372 \text{ MeV} \\
+  E_{\ce{^8Be}} &= 3.1086 \text{ MeV} \\
+  \theta_{\alpha_0} &= 165 \\
+  \theta_{\ce{Be8}} &= 11 \\
+  Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
+\end{split}
+\label{eq:energiesBr}
+\end{equation}
+
 \printbibliography

### 2023-12-15 analise.py
hihi

@@ -1,0 +1,59 @@
+import numpy as np
+import xml.etree.ElementTree as ET
+import matplotlib.pyplot as plt
+
+# Function to read the data from the file
+def reader (name, Chn):
+    # Parse the XML data from the file
+    tree = ET.parse(name)
+    root = tree.getroot()
+    # Access spectrum data
+    spectrum_data_str = root.find(f".//Spectrum[@id='RadMeasurement-{Chn}_Spectrum-{Chn}']/ChannelData").text
+    # Convert the spectrum data string to a vector of data
+    spectrum_data_array = np.array([float(value) for value in spectrum_data_str.split()])
+
+    return spectrum_data_array
+
+# Reading the data files
+
+# TT_20
+TT20_Chn0 = reader('data/TT_20/UNFILTERED/CH0@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 0)
+TT20_Chn1 = reader('data/TT_20/UNFILTERED/CH1@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 1)
+TT20_Chn2 = reader('data/TT_20/UNFILTERED/CH2@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 2)
+
+# TT_21
+TT21_Chn0 = reader('data/TT_21/UNFILTERED/CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
+TT21_Chn1 = reader('data/TT_21/UNFILTERED/CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
+TT21_Chn2 = reader('data/TT_21/UNFILTERED/CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
+
+# TT_22
+TT22_Chn0 = reader('data/TT_22/UNFILTERED/CH0@N6781_21198_Espectrum_TT_22_20231207_144117.n42', 0)
+TT22_Chn1 = reader('data/TT_22/UNFILTERED/CH1@N6781_21198_Espectrum_TT_22_20231207_144117.n42', 1)
+TT22_Chn2 = reader('data/TT_22/UNFILTERED/CH2@N6781_21198_Espectrum_TT_22_20231207_144117.n42', 2)
+
+# TT_23
[…]

### 2023-12-15 analise.py
SHIT

@@ -1,6 +1,7 @@
 import numpy as np
 import xml.etree.ElementTree as ET
 import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
 
 # Function to read the data from the file
 def reader (name, Chn):
@@ -55,5 +56,143 @@ TT26_Chn2 = reader('data/TT_26/UNFILTERED/CH2@N6781_21198_Espectrum_TT_26_202312
 TT27_Chn0 = reader('data/TT_27/UNFILTERED/CH0@N6781_21198_Espectrum_TT_27_20231207_160722.n42', 0)
 TT27_Chn1 = reader('data/TT_27/UNFILTERED/CH1@N6781_21198_Espectrum_TT_27_20231207_160722.n42', 1)
 TT27_Chn2 = reader('data/TT_27/UNFILTERED/CH2@N6781_21198_Espectrum_TT_27_20231207_160722.n42', 2)
+
+
+# Plotting the data
+
+# TT_20
+plt.figure()
+plt.plot(TT20_Chn0, label='Chnannel 0')
+plt.plot(TT20_Chn1, label='Channel 1')
+plt.plot(TT20_Chn2, label='Channel 2')
+plt.legend()
+plt.title('TT_20')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.savefig('TT_20.png')    
+plt.show()
+
+# TT_21
+plt.figure()
+plt.plot(TT21_Chn0, label='Chnannel 0')
+plt.plot(TT21_Chn1, label='Channel 1')
+plt.plot(TT21_Chn2, label='Channel 2')
+plt.legend()
+plt.title('TT_21')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.savefig('TT_21.png')
+plt.show()
+
+# TT_22
+plt.figure()
+plt.plot(TT22_Chn0, label='Chnannel 0')
+plt.plot(TT22_Chn1, label='Channel 1')
+plt.plot(TT22_Chn2, label='Channel 2')
+plt.legend()
+plt.title('TT_22')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.savefig('TT_22.png')
+plt.show()
+
+# TT_23
+plt.figure()
+plt.plot(TT23_Chn0, label='Chnannel 0')
[…]

### 2023-12-15 analise.py
SHIT

@@ -189,7 +189,6 @@ y = TT21_Chn0[500:700]
 
 
 
-
 energy = np.array(5156.59, 5485.56, 5804.82)
 
 

### 2023-12-15 analise.py
SHIT

@@ -188,7 +188,7 @@ x = np.linspace(500, 700, 1000)
 y = TT21_Chn0[500:700]
 
 
-
+# valores de energia para os 3 picos
 energy = np.array(5156.59, 5485.56, 5804.82)
 
 

### 2023-12-15 analise.py
SHIT

@@ -184,8 +184,10 @@ def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
     return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
 
 # Peaks
-x = np.linspace(500, 700, 1000)
+x = np.linspace(500, 700, 200)
 y = TT21_Chn0[500:700]
+plt.plot(x, y, label='data') 
+
 
 
 # valores de energia para os 3 picos

### 2023-12-15 analise.py
SHIT

@@ -187,6 +187,18 @@ def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
 x = np.linspace(500, 700, 200)
 y = TT21_Chn0[500:700]
 plt.plot(x, y, label='data') 
+# curve fit
+popt, pcov = curve_fit(gaussian, x, y, p0=[100, 550, 1, 100, 580, 1, 100, 610, 1, 0])
+plt.plot(x, gaussian(x, *popt), label='fit')
+plt.legend()
+plt.title('TT_21')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.yscale('log')
+plt.xlim(500, 700)
+#plt.savefig('TT_21_Chn0_fit.png')
+plt.show()
+
 
 
 

### 2023-12-15 calib.py
calib.py

@@ -1,7 +1,7 @@
 import numpy as np
 import xml.etree.ElementTree as ET
 import matplotlib.pyplot as plt
-import scipy as scipy
+from scipy.optimize import curve_fit
 
 # Function to read the data from the file
 def reader (name, Chn):
@@ -59,7 +59,19 @@ x = np.array(range(500, 701))
 print(x)
 y = TT21_Chn0[500:700]
 print(y)
-scipy.optimize.curve_fit(gaussian, x,y)
+popt, popcov = curve_fit(gaussian, x,y)
+
+plt.figure()
+plt.plot(TT21_Chn0, label='Chnannel 0')
+plt.plot(x, gaussian(x, *popt), label='Chnannel 0 fit')
+plt.legend()
+plt.xlabel('Channel')
+plt.yscale('log')
+plt.ylabel('Counts')
+plt.xlim(500, 700)
+plt.ylim(1, 1e3)
+plt.savefig('TT_21_Chn0.png')
+plt.show()
 
 # valores de energia para os 3 picos
 energy = np.array([5156.59, 5485.56, 5804.82])

### 2023-12-15 calib.py
fit channel0 finnished

@@ -38,38 +38,26 @@ plt.show()
 
 # For the calibration we will use the data from the TT_21 file
 
-plt.figure()
-plt.plot(TT21_Chn0, label='Chnannel 0')
-plt.legend()
-plt.title('TT_21')
-plt.xlabel('Channel')
-plt.yscale('log')
-plt.ylabel('Counts')
-plt.xlim(500, 700)
-plt.ylim(1, 1e3)
-plt.savefig('TT_21_Chn0.png')
-plt.show()
-
 # Gaussian fit for the peaks (soma 3 de gaussianas)
 def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
     return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
 
 # Peaks
-x = np.array(range(500, 701))
-print(x)
-y = TT21_Chn0[500:700]
-print(y)
-popt, popcov = curve_fit(gaussian, x,y)
+x = np.array(range(545, 650))
+y = TT21_Chn0[545:650]
+# curve fit
+popt, pcov = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
 
 plt.figure()
 plt.plot(TT21_Chn0, label='Chnannel 0')
 plt.plot(x, gaussian(x, *popt), label='Chnannel 0 fit')
+plt.title('Fit channel 0')
 plt.legend()
 plt.xlabel('Channel')
 plt.yscale('log')
 plt.ylabel('Counts')
 plt.xlim(500, 700)
-plt.ylim(1, 1e3)
+plt.ylim(1, 1e2)
 plt.savefig('TT_21_Chn0.png')
 plt.show()
 

### 2023-12-16 analise.py
Organize Paths

@@ -2,6 +2,24 @@ import numpy as np
 import xml.etree.ElementTree as ET
 import matplotlib.pyplot as plt
 from scipy.optimize import curve_fit
+from pathlib import Path
+
+# Define paths
+# data
+DATA_PATH = Path('data')
+# data folders
+TT_20_PATH = DATA_PATH / 'TT_20' / 'UNFILTERED'
+TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED'
+TT_22_PATH = DATA_PATH / 'TT_22' / 'UNFILTERED'
+TT_23_PATH = DATA_PATH / 'TT_23' / 'UNFILTERED'
+TT_24_PATH = DATA_PATH / 'TT_24' / 'UNFILTERED'
+TT_25_PATH = DATA_PATH / 'TT_25' / 'UNFILTERED'
+TT_26_PATH = DATA_PATH / 'TT_26' / 'UNFILTERED'
+TT_27_PATH = DATA_PATH / 'TT_27' / 'UNFILTERED'
+# image
+IMAGE_PATH = Path('images')
+IMAGE_PATH.mkdir(exist_ok=True, parents=True)
+
 
 # Function to read the data from the file
 def reader (name, Chn):
@@ -18,44 +36,44 @@ def reader (name, Chn):
 # Reading the data files
 
 # TT_20
-TT20_Chn0 = reader('data/TT_20/UNFILTERED/CH0@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 0)
-TT20_Chn1 = reader('data/TT_20/UNFILTERED/CH1@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 1)
-TT20_Chn2 = reader('data/TT_20/UNFILTERED/CH2@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 2)
+TT20_Chn0 = reader(TT_20_PATH / 'CH0@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 0)
+TT20_Chn1 = reader(TT_20_PATH /'CH1@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 1)
+TT20_Chn2 = reader(TT_20_PATH /'CH2@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 2)
 
 # TT_21
[…]

### 2023-12-17 calib.py
NUNO GIT PULL

@@ -42,7 +42,7 @@ plt.show()
 def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
     return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
 
-# Peaks
+# Channel 0
 x = np.array(range(545, 650))
 y = TT21_Chn0[545:650]
 # curve fit
@@ -60,6 +60,45 @@ plt.xlim(500, 700)
 plt.ylim(1, 1e2)
 plt.savefig('TT_21_Chn0.png')
 plt.show()
+
+# Channel 1
+x = np.array(range(545, 700))
+y = TT21_Chn1[545:700]
+# curve fit
+popt, pcov = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
+
+plt.figure()
+plt.plot(TT21_Chn1, label='Chnannel 1')
+plt.plot(x, gaussian(x, *popt), label='Chnannel 1 fit')
+plt.title('Fit channel 1')
+plt.legend()
+plt.xlabel('Channel')
+plt.yscale('log')
+plt.ylabel('Counts')
+plt.xlim(500, 700)
+plt.ylim(1, 1e2)
+plt.savefig('TT_21_Chn1.png')
+plt.show()
+
+# Channel 2
+x = np.array(range(545, 700))
+y = TT21_Chn2[545:700]
+# curve fit
+popt, pcov = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
+
+plt.figure()
+plt.plot(TT21_Chn2, label='Chnannel 2')
+plt.plot(x, gaussian(x, *popt), label='Chnannel 2 fit')
+plt.title('Fit channel 2')
+plt.legend()
+plt.xlabel('Channel')
+plt.yscale('log')
+plt.ylabel('Counts')
+plt.xlim(500, 700)
+plt.ylim(1, 1e2)
+plt.savefig('TT_21_Chn2.png')
+plt.show()
+
 
 # valores de energia para os 3 picos
 energy = np.array([5156.59, 5485.56, 5804.82])

### 2023-12-17 calib.py
NUNO GIT PULL

@@ -46,11 +46,11 @@ def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
 x = np.array(range(545, 650))
 y = TT21_Chn0[545:650]
 # curve fit
-popt, pcov = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
+popt0, pcov0 = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
 
 plt.figure()
 plt.plot(TT21_Chn0, label='Chnannel 0')
-plt.plot(x, gaussian(x, *popt), label='Chnannel 0 fit')
+plt.plot(x, gaussian(x, *popt0), label='Chnannel 0 fit')
 plt.title('Fit channel 0')
 plt.legend()
 plt.xlabel('Channel')
@@ -65,11 +65,11 @@ plt.show()
 x = np.array(range(545, 700))
 y = TT21_Chn1[545:700]
 # curve fit
-popt, pcov = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
+popt1, pcov1 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
 plt.figure()
 plt.plot(TT21_Chn1, label='Chnannel 1')
-plt.plot(x, gaussian(x, *popt), label='Chnannel 1 fit')
+plt.plot(x, gaussian(x, *popt1), label='Chnannel 1 fit')
 plt.title('Fit channel 1')
 plt.legend()
 plt.xlabel('Channel')
@@ -84,11 +84,11 @@ plt.show()
 x = np.array(range(545, 700))
 y = TT21_Chn2[545:700]
 # curve fit
-popt, pcov = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
+popt2, pcov2 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
 plt.figure()
 plt.plot(TT21_Chn2, label='Chnannel 2')
-plt.plot(x, gaussian(x, *popt), label='Chnannel 2 fit')
+plt.plot(x, gaussian(x, *popt2), label='Chnannel 2 fit')
[…]

### 2023-12-17 calib.py
NUNO GIT PULL

@@ -133,6 +133,10 @@ plt.ylabel('Energy (keV)')
 plt.savefig('TT_21_Chn0_calib.png')
 plt.show()
 
+m0 = params0[0]
+c0 = params0[1]
+print(f'Calibração_0: E = {m0:.2f} * Chn + {c0:.2f}')
+
 # Calibração (channel 1)
 params1, params_cov1 = curve_fit(linear, x1, energy)
 errors1 = np.sqrt(np.diag(params_cov1))
@@ -147,6 +151,10 @@ plt.ylabel('Energy (keV)')
 plt.savefig('TT_21_Chn1_calib.png')
 plt.show()
 
+m1 = params1[0]
+c1 = params1[1]
+print(f'Calibração_1: E = {m1:.2f} * Chn + {c1:.2f}')
+
 # Calibração (channel 2)
 params2, params_cov2 = curve_fit(linear, x2, energy)
 errors2 = np.sqrt(np.diag(params_cov2))
@@ -160,3 +168,7 @@ plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
 plt.savefig('TT_21_Chn2_calib.png')
 plt.show()
+
+m2 = params2[0]
+c2 = params2[1]
+print(f'Calibração_2: E = {m2:.2f} * Chn + {c2:.2f}')

### 2023-12-17 calib.py
NUNO GIT PULL

@@ -122,6 +122,8 @@ a3 = np.array([popt2[0], popt2[3], popt2[6]])
 # Calibração (channel 0)
 params0, params_cov0 = curve_fit(linear, x0, energy)
 errors0 = np.sqrt(np.diag(params_cov0))
+m0 = params0[0]
+c0 = params0[1]
 
 plt.figure()
 plt.plot(x0, energy, 'o', label='data')
@@ -133,13 +135,13 @@ plt.ylabel('Energy (keV)')
 plt.savefig('TT_21_Chn0_calib.png')
 plt.show()
 
-m0 = params0[0]
-c0 = params0[1]
 print(f'Calibração_0: E = {m0:.2f} * Chn + {c0:.2f}')
 
 # Calibração (channel 1)
 params1, params_cov1 = curve_fit(linear, x1, energy)
 errors1 = np.sqrt(np.diag(params_cov1))
+m1 = params1[0]
+c1 = params1[1]
 
 plt.figure()
 plt.plot(x1, energy, 'o', label='data')
@@ -148,16 +150,17 @@ plt.title('Calibration channel 1')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
+plt.text(560, 5500, f'E = {m1:.2f} * Chn + {c1:.2f}')
 plt.savefig('TT_21_Chn1_calib.png')
 plt.show()
 
-m1 = params1[0]
-c1 = params1[1]
 print(f'Calibração_1: E = {m1:.2f} * Chn + {c1:.2f}')
 
 # Calibração (channel 2)
 params2, params_cov2 = curve_fit(linear, x2, energy)
 errors2 = np.sqrt(np.diag(params_cov2))
+m2 = params2[0]
+c2 = params2[1]
 
 plt.figure()
 plt.plot(x2, energy, 'o', label='data')
@@ -169,6 +172,4 @@ plt.ylabel('Energy (keV)')
 plt.savefig('TT_21_Chn2_calib.png')
 plt.show()
 
-m2 = params2[0]
-c2 = params2[1]
 print(f'Calibração_2: E = {m2:.2f} * Chn + {c2:.2f}')

### 2023-12-17 calib.py
NUNO GIT PULL

@@ -132,10 +132,11 @@ plt.title('Calibration channel 0')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
+plt.text(560, 5600, f'E = ({m0:.2f} ± {errors0[0]:.2f}) * Chn + {c0:.2f} ± {errors0[1]:.2f}', fontsize=10)
 plt.savefig('TT_21_Chn0_calib.png')
 plt.show()
 
-print(f'Calibração_0: E = {m0:.2f} * Chn + {c0:.2f}')
+print(f'Calibração_0: E = ({m0:.2f} ± {errors0[0]:.2f}) * Chn + {c0:.2f} ± {errors0[1]:.2f}')
 
 # Calibração (channel 1)
 params1, params_cov1 = curve_fit(linear, x1, energy)
@@ -150,11 +151,11 @@ plt.title('Calibration channel 1')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
-plt.text(560, 5500, f'E = {m1:.2f} * Chn + {c1:.2f}')
+plt.text(570, 5600, f'E = ({m1:.2f} ± {errors1[0]:.2f}) * Chn + {c1:.2f} ± {errors1[1]:.2f}', fontsize=10)
 plt.savefig('TT_21_Chn1_calib.png')
 plt.show()
 
-print(f'Calibração_1: E = {m1:.2f} * Chn + {c1:.2f}')
+print(f'Calibração_1: E = ({m1:.2f} ± {errors1[0]:.2f}) * Chn + {c1:.2f} ± {errors1[1]:.2f}')
 
 # Calibração (channel 2)
 params2, params_cov2 = curve_fit(linear, x2, energy)
@@ -169,7 +170,9 @@ plt.title('Calibration channel 2')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
+plt.text(565, 5600, f'E = ({m2:.2f} ± {errors2[0]:.2f}* Chn + {c2:.2f}) ± {errors2[1]:.2f}', fontsize=10)
 plt.savefig('TT_21_Chn2_calib.png')
 plt.show()
 
-print(f'Calibração_2: E = {m2:.2f} * Chn + {c2:.2f}')
+print(f'Calibração_2: E = ({m2:.2f} ± {errors2[0]:.2f}) * Chn + {c2:.2f} ± {errors2[1]:.2f}')
+

### 2023-12-18 text/paper/paper.tex
início do paper

--- 2023-11-30 text/notas/notas.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[portuguese]{babel}
-\usepackage[letterpaper,top=10mm,bottom=15mm,left=10mm,right=10mm,marginparwidth=1.75cm]{geometry}
+\usepackage[english]{babel}
+\usepackage[letterpaper,top=6mm,bottom=15mm,left=6mm,right=6mm,marginparwidth=1.55cm]{geometry}
 \usepackage{multicol}
@@
 
-\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Notas Teóricas \vspace{-5mm}}
-\date{30 Novembro 2023}
+\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}
+% \\ \vspace{1mm} LFEUI \vspace{1mm} 
+\\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-5mm}}
+\date{December 2023 - January 2024}
 
@@
 \addbibresource{../references/prop.bib} % Specify your .bib file
+
+\setlist[enumerate]{itemsep=3pt}
 
@@
 
-\vspace{-5mm}
+\vspace{-8mm}
+
+\hrulefill
+
+% \begin{center}
+%   \textbf{\Large Abstract}
+% \end{center}
+
+\vspace{-0.1cm}
+\begin{abstract}
+  \par 
+\end{abstract}
+\vspace{-0.1cm}
 
@@
 \begin{multicols}{2}
-
-\section{Acelerador de Partículas}
-
-O acelerador utilizado possui no seu interior um gás.
-
-\section{Lítio}
-
-Vamos trabalhar com o lítio, um elemento altamente reativo, maioritariamente através de reações de oxidação com o vapor de água. 
-A solução para trabalhar com este elemento é usar fontes inertes - uma fonte inerte de lítio consiste em lítio no interior de duas camadas de vidro.
-O vidro é composto por silicatos de boro.
-
-\section{Objetivo}
-
-O objetivo é obter o espetro de emissão de duas fontes diferentes, ambas contendo lítio.
-
-\section{Reação}
-
-A reação que pretendemos estudar envolve o choque de um átomo de lítio 7 com um protão, que resulta na formação de um átomo de berílio 8 que decai para duas partículas alfa (núcleo de hélio 4) - o decaimento é tão rápido que não é detetado.
-
-\begin{equation}
-    \ce{^7Li + ^1H -> ^8Be -> 2\alpha = 2^4He}
-\end{equation}
-
-Esta reação é idêntica para as duas amostras que vamos estudar.
-
-\subsection*{Valor $Q$ da reação}
-
-O valor $Q$ da reação é a energia libertada na reação - é a energia que as partículas alfa possuem quando são emitidas.
-Calcula-se a partir da diferença de massa entre os reagentes e os produtos da reação:
-
-\begin{equation}
-  \begin{split}
-      Q &= \Delta m c^2 \\
-      &= c^2 \left( m_{\text{reagents}} - m_{\text{products}} \right) \\
-      &= c^2 \left( m_{\ce{^7Li}} + m_{\ce{^1H}} - m_{\ce{^8Be}} \right)
-  \end{split}
-\end{equation}
-
-\section{Processo Experimental}
-
-Queremos obter o espetro energético para perceber qual a energia com que as partículas são emitidas
-
-\section{Montagem Experimental}
-
-Temos um acelerador Tandem de 2.5 MeV, que acelera protões - temos 2 tubos de aceleração em paralelo o que faz com que a tensão à qual os partículas protões são aceleradas possa chegar aos 5 MV.
-Os dois tubos estão alinhados o que faz com que o protão seja 'duplamente' acelerado.
-Acelerar um protão a 5 MeV significa que a variação de energia cinética do eletrão desde o íncio do tubo (em repouso) até ao seu fim é de 5 MeV - este é o valor da diferença de potencial elétrico entre as duas extremidades/terminais do tubo de aceleração.
-Este acelerador possui um elevador de tensão - associação de condensadores e resistências que permitem aumentar a tensão.
-
-\paragraph*{Vácuo} O feixe propaga-se ao longo dos dois tubos de aceleração que se encontra sobre um vácuo de $10^{-7}$ mbar.
-O vácuo é necessário para que as partículas não colidam com moléculas de ar e percam energia.
-Este vácuo é bastante bom sendo possível que existam colisão no caminho apesar de em média o número destas colisões é nulo.
-A este nível de pressão ($10^{-7}$ mbar), existem cerca de $10^9$ moléculas de ar por cm$^3$.
-Na atmosfera normal (pressão de 1 bar) existem $10^{19}$ moléculas de ar por cm$^3$ - no vácuo intergalático temos 5 moléculas por cm$^3$ - apesar de o vácuo deste vácuo não ser comparável ao do espaço, é um vácuo bastante bom.
-Para estimar estas quantidades de partículas por cm$^3$ consideramos que a diferença de pressão entre a atmosfera normal e o vácuo e consideramos que o volume de um mol de gás é $22.4$ L/mol na atmosfera normal - isto implica que há uma mol de moléculas por litro - $6.02 \times 10^{23}$ moléculas por litro.
-
-\paragraph*{Eletroímane} No nosso eletroímane, partículas com diferentes racios carga/massa têm ângulos de deflexão diferentes - o ângulo de deflexão é inversamente proporcional ao racio carga/massa.
-Este elemento é responsável por fazer com que todas as partículas que não os protões que se encontrem no feixe sejam desviadas para fora do feixe e por isso não atingem o sensor.
-É por isto que existe um grande benefício em ter um detetor não linear - fitração do feixe de partículas.
-
-\paragraph*{Detetores} Vamos usar detetores de silício - a energia de uma partícula incidente no detetor é calculada a partir do número de pares electrão-buraco que são criados no detetor devido à sua passagem.
-
-\paragraph*{Eletrónica} Existe um sistema de eletrónica digital responsável por fazer a leitura dos detetores e por enviar os sinais para o computador onde podemos observar os espetros obtidos.
-
-\section{Amostras}
-
-A atividade envolve o estudo dos espetros provenientes de duas amostras diferentes, ambas contendo lítio - uma amostra que é apenas um vidro composto por silicatos e óxidos e outra que é uma amostra com lítio implantado em silício.
-
-\paragraph*{Implantação} A implantação é o processo de formacação de amostras implantadas - para obter uma amostra de lítio 7 implantado em silício usou-se um feixe de lítio 7 com potência na ordem dos kilovolts - quanto maior a potência do feixe, maior é a profundidade a que o lítio é implantado no silício.
-Existe um fórmula que nos permite saber a profundidade a que o lítio é implantado no silício em função da potência do feixe de lítio 7 e da densidade do silício.
-
-\paragraph*{Diferenças} O objetivo de estudar estas duas amostras é avaliar a diferença de espetros de emissão de partículas alfa provenientes de lítio 7 em diferentes profundidades.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|w{0.2\linewidth}|w{0.7\linewidth}|}
-\hline
-Amostra & Descrição \\ \hline
-Vidro Normal & Amostra formada por vidro normal com silício, lítio, boro e talvez outros resíduos - o lítio aparece na forma de óxido de lítio de modo a que seja pouco reativo.
-Esta amaostra é homogénea na sua constituição ao longo de toda a sua espessura uma vez que o vidro é formado diretamente com os elementos.
-\\ \hline
-Implantada & Amostra obtida através da implatação de uma camada de lítio numa amostra de sílicio (pura) - o que obtemos é uma amostra que nos primeiros nanómetrod e espessura possui lítio e depois apenas possui silício.
-\\ \hline
-\end{tabular}
-\end{table}
-
-\section{Resultados Esperados}
-
-\paragraph*{Perfis} Um perfil é uma parte contínua do espetro na zona energética anterior aos picos de partículas alfa.
-Esperamos ver isto na amostra de vidro devido às partículas alfa que provém de diferentes profundidades da amostra de vidro.
-Quanto maior for a profundidade, como estamos a falar de partículas carregadas e não de fotões, maior é a energia dissipada para o material durante o percurso. 
-Para partículas emitidas com a mesma energia, quanto maior for a profundidade, menor é a energia com que a partícula é detetada - cria-se um perfil no espetro devido a isto. 
-Como na amostra implantada o lítio está apenas presente numa camada superior muito fina, não esperamos obter um perfil no espetro.
-
-\paragraph*{RBS - Rutherford Backscattering} é um fenómeno no qual as partículas alfa são retrodispaersas para outras direções e por isso possuem energias diferentes. 
-A maior parte das partículas alfa são detetadas com o valor de energia esperada mas podemos ter outras zonas do espetro onde tenhamos partículas alfa com energias diferentes e dispersas devido a este fenómeno.
-
-\paragraph*{Elastic Backscattering} Isto acontece quando um dos protões do feixe incidente é refletido elasticamente na superfície da amostra - o protão é refletido com a mesma energia com que incidiu na amostra e continua a trajetória na mesma direção mas com sentido oposto.
-
-\section{Utilidade}
-
-\begin{enumerate}
-  \item Fusão Nuclear - lítio 6 para produção de neutrões -  usam-se materiais muito enriquecidos em lítio 6;
-  \item Baterias - lítio 7;
-  \item Indústria Farmaceûtica - lítio é utilizado em vários medicamentos;
-\end{enumerate}
 

### 2023-12-18 text/proposta/prop.tex
Ok
gtrg
início do paper

--- 2023-12-02 text/proposta/prop.tex
@@
 
-\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Investigation Proposal \\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-0mm}}
+\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{0mm}
+LFEUI - Investigation Proposal \\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-5mm}}
 \date{November 2023}
@@
 
-The reaction results in the production of a $\ce{^8Be}$ nucleus which then rapidly decays into two alpha particles. These are the particles we detect.
+Before the emission of the reaction products, the reaction leads to the formation of a $\ce{^8Be}$ compound nucleus (intermediate state), which then rapidly decays into two alpha particles.
+These are the particles we detect.
 
-According to the incident proton's energy and the scattering angle of the detectors ($\theta = 165 $ degrees) we were able to use the NRA Calculator \cite{NRAEnergyCalc} to calculate the energy of the alpha particles resulting from the reaction as well as it's $Q$ value:
+According to the incident proton's energy ($E_p = 2.7 $ MeV) and the scattering angle of the detectors ($\theta = 165 $ degrees) we were able to use the NRA Calculator \cite{NRAEnergyCalc} to calculate the energy of the alpha particles resulting from the reaction as well as it's $Q$ value:
 
@@
 The electromagnets deflect particles with different angles according to their charge-to-mass ratio and, with that, allows us to remove any unwanted particles from the beam and only let the protons pass through.
-The detectors used are silicon detectors - these detect charged particles through the ionization they produce in the silicon.
+The detectors used are silicon based detectors - these detect the energy of incident charged particles through the ionisation they produce in the silicon.
 
-In addition, they are connected to a electronic system that amplifies the signal and sends it to a computer where we
-can see the results.
+In addition, they are connected to a electronic system that amplifies and formats the electric signal arising from individual interaction events and sends it to a multichannel analyser where we can see the collected energy spectra.
 
@@
 
-We are going to work with two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and
-lithium and the other one is a implant sample, which is a thin layer of lithium implanted in a silicon substrate.
+We are going to work with two different Lithium-7 samples: one of them is a glass sample composed of silicon, boron and lithium and the other ones will be silicon substrates implanted with $\ce{^7Li}$.
 
@@
   \item Accelerator Startup;
-  \item Calibration of the detectors with a sample of known composition;
+  \item Channel-energy calibration of the detection system with an alpha source of known composition;
   \item Obtaining the energy spectrums of both samples with the appropriate beam energy;
@@
 
-\section{Results expected}
+\section{Expected results}
 
@@
 In both cases, the samples are composed by other types of atoms such as silicon and boron, which will also emit charged
-particles. By obtaining the expected energy spectrums we can infer the presence of Li in the samples. If possible we will also try
-to quantify the amount of Li present by analyzing the amount of energy deposited in the spectrums.
+particles. By obtaining the expected energy spectra we can infer the presence of Li in the samples. If possible we will also try
+to quantify the amount of Li present by analysing the yield of the $\ce{^7Li}$(p,$\ce{^4He})\ce{^4He}$ nuclear reaction.
 
-\paragraph*{Energy Values} The alpha particles are supposed to have the energies calculated in equation \ref{eq:energies} ($E_{\alpha_1}$ and $E_{\alpha_2}$). 
-In the glass sample, we expect to see an energy profile before each peak and in the implant sample we only expect to see the peaks.
+\paragraph*{Energy Values} %The alpha particles are supposed to have the energies calculated in equation \ref{eq:energies} ($E_{\alpha_1}$ and $E_{\alpha_2}$). 
+From the glass sample we expect to see a continuous nuclear reaction yield and from the implanted samples we only expect to see well defined nuclear reaction yields
 
@@
 
-Other phenomena such as Rutherford Backscattering and Elastic Scattering are also expected to occur.
+Other interaction phenomena such as Rutherford Backscattering and Elastic Scattering will also be visible in the spectra.
 These are responsible for energy peaks in both sample's spectra at lower than the expected energy values.
@@
 
-By identifying the presence of light isotopes at low energies ($<5$MeV), we achieve a practical method to evaluate the presence and the amount of these light isotopes in technological materials.
-Until now, the only way to do this was through atomic spectroscopy, which, besides not being able to detect the presence of light isotopes, is a much more complex and expensive method.
+By identifying the presence of light isotopes at low energies ($<5$ MeV), we achieve a practical method to evaluate the presence and the amount of these light isotopes in technological materials.
+Until now, the only way to do this was through atomic spectroscopy techniques, which, besides not being able to detect the presence of lighter elements, they also present lower sensitivities for elemental quantification.
 

### 2023-12-18 analise.py
small changes

@@ -145,7 +145,6 @@ plt.show()
 
 # TT_25
 plt.figure()
-plt.subplot(1, 3, 1)
 plt.plot(TT25_Chn0, label='Chnannel 0')
 plt.plot(TT25_Chn1, label='Channel 1')
 plt.plot(TT25_Chn2, label='Channel 2')

### 2023-12-18 calib.py

@@ -43,8 +43,8 @@ def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
     return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
 
 # Channel 0
-x = np.array(range(545, 650))
-y = TT21_Chn0[545:650]
+x = np.array(range(500, 700))
+y = TT21_Chn0[500:700]
 # curve fit
 popt0, pcov0 = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
 
@@ -62,8 +62,8 @@ plt.savefig('TT_21_Chn0.png')
 plt.show()
 
 # Channel 1
-x = np.array(range(545, 700))
-y = TT21_Chn1[545:700]
+x = np.array(range(500, 700))
+y = TT21_Chn1[500:700]
 # curve fit
 popt1, pcov1 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
@@ -81,8 +81,8 @@ plt.savefig('TT_21_Chn1.png')
 plt.show()
 
 # Channel 2
-x = np.array(range(545, 700))
-y = TT21_Chn2[545:700]
+x = np.array(range(500, 700))
+y = TT21_Chn2[500:700]
 # curve fit
 popt2, pcov2 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
@@ -170,7 +170,7 @@ plt.title('Calibration channel 2')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
-plt.text(565, 5600, f'E = ({m2:.2f} ± {errors2[0]:.2f}* Chn + {c2:.2f}) ± {errors2[1]:.2f}', fontsize=10)
+plt.text(565, 5600, f'E = ({m2:.2f} ± {errors2[0]:.2f})* Chn + {c2:.2f} ± {errors2[1]:.2f}', fontsize=10)
 plt.savefig('TT_21_Chn2_calib.png')
 plt.show()
 

### 2023-12-18 analise.py
Solve Merge

@@ -3,6 +3,7 @@ import xml.etree.ElementTree as ET
 import matplotlib.pyplot as plt
 from scipy.optimize import curve_fit
 from pathlib import Path
+from reader import reader
 
 # Define paths
 # data
@@ -19,19 +20,6 @@ TT_27_PATH = DATA_PATH / 'TT_27' / 'UNFILTERED'
 # image
 IMAGE_PATH = Path('images')
 IMAGE_PATH.mkdir(exist_ok=True, parents=True)
-
-
-# Function to read the data from the file
-def reader (name, Chn):
-    # Parse the XML data from the file
-    tree = ET.parse(name)
-    root = tree.getroot()
-    # Access spectrum data
-    spectrum_data_str = root.find(f".//Spectrum[@id='RadMeasurement-{Chn}_Spectrum-{Chn}']/ChannelData").text
-    # Convert the spectrum data string to a vector of data
-    spectrum_data_array = np.array([float(value) for value in spectrum_data_str.split()])
-
-    return spectrum_data_array
 
 # Reading the data files
 

### 2023-12-18 calib.py

@@ -2,23 +2,24 @@ import numpy as np
 import xml.etree.ElementTree as ET
 import matplotlib.pyplot as plt
 from scipy.optimize import curve_fit
+from pathlib import Path
+from reader import reader
+import pandas as pd
 
-# Function to read the data from the file
-def reader (name, Chn):
-    # Parse the XML data from the file
-    tree = ET.parse(name)
-    root = tree.getroot()
-    # Access spectrum data
-    spectrum_data_str = root.find(f".//Spectrum[@id='RadMeasurement-{Chn}_Spectrum-{Chn}']/ChannelData").text
-    # Convert the spectrum data string to a vector of data
-    spectrum_data_array = np.array([float(value) for value in spectrum_data_str.split()])
-
-    return spectrum_data_array
+# Define paths
+# For the calibration we will use the data from the TT_21 file
+# data
+DATA_PATH = Path('data')
+# data folders
+TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED'
+# image
+IMAGE_PATH = Path('images')
+IMAGE_PATH.mkdir(exist_ok=True, parents=True)
 
 # TT_21
-TT21_Chn0 = reader('data/TT_21/UNFILTERED/CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
-TT21_Chn1 = reader('data/TT_21/UNFILTERED/CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
-TT21_Chn2 = reader('data/TT_21/UNFILTERED/CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
+TT21_Chn0 = reader(TT_21_PATH/'CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
+TT21_Chn1 = reader(TT_21_PATH/'CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
+TT21_Chn2 = reader(TT_21_PATH/'CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
 
 # TT_21
 plt.figure()
[…]

### 2023-12-18 calib.py
Arrange Paths in calib.py

@@ -48,8 +48,8 @@ def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
     return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
 
 # Channel 0
-x = np.array(range(545, 650))
-y = TT21_Chn0[545:650]
+x = np.array(range(500, 700))
+y = TT21_Chn0[500:700]
 # curve fit
 popt0, pcov0 = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
 
@@ -67,8 +67,8 @@ plt.savefig(IMAGE_PATH/'TT_21_Chn0.png')
 plt.show()
 
 # Channel 1
-x = np.array(range(545, 700))
-y = TT21_Chn1[545:700]
+x = np.array(range(500, 700))
+y = TT21_Chn1[500:700]
 # curve fit
 popt1, pcov1 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
@@ -86,8 +86,8 @@ plt.savefig(IMAGE_PATH/'TT_21_Chn1.png')
 plt.show()
 
 # Channel 2
-x = np.array(range(545, 700))
-y = TT21_Chn2[545:700]
+x = np.array(range(500, 700))
+y = TT21_Chn2[500:700]
 # curve fit
 popt2, pcov2 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
 
@@ -176,7 +176,7 @@ plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy (keV)')
 plt.text(565, 5600, f'E = ({m2:.2f} ± {errors2[0]:.2f}* Chn + {c2:.2f}) ± {errors2[1]:.2f}', fontsize=10)
-plt.savefig(IMAGE_PATH/'TT_21_Chn2_calib.png')
+plt.savefig('TT_21_Chn2_calib.png')
 plt.show()
 
 print(f'Calibração_2: E = ({m2:.2f} ± {errors2[0]:.2f}) * Chn + {c2:.2f} ± {errors2[1]:.2f}')

### 2023-12-18 text/LogBooks/logbook2.tex
Resources and Images

--- 2023-12-07 text/LogBooks/logbook2.tex
@@
 \title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm}
-LFEUI - Logbook 1\vspace{-5mm}}
-\date{30 November 2023}
+LFEUI - Logbook 2\vspace{-5mm}}
+\date{7 December 2023}
 
@@
 
+\vspace{5mm}
+
+We arrived at Centro Tecnológico e Nuclear at 9:30 AM in the 9$^{\text{th}}$ of December to meet with investigators Rodrigo Mateus and Norberto Catarino.
+
+\section{Samples}
+
+We started by discussing the samples that we will be using during the day.
+In the previous session, we had already concluded that there would be two different types of samples: glass samples with lithium in their composition and also implanted samples.
+The latter consists of a thin layer of lithium implanted in an aluminium substrate.
+As for the glass samples, they are not available and we will simply use three different samples, each one with a different substance/material whose composition varies from sample to sample.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|w{2cm}|w{4cm}|w{3cm}|}
+\hline
+Samples & Description & Composition \\ \hline
+% 1 & Lithium battery & $\ce{Li}$ \\ \hline
+1 & Lithium fluoride & $\ce{LiF}$ \\ \hline
+2 & Lithium aluminate & $\ce{LiAlO2}$ \\ \hline
+3 & Implanted sample & $\ce{Li}$ implanted in $\ce{Al}$ \\ \hline
+\end{tabular}
+\caption{Samples used in the experiment}
+\label{tab:samples}
+\end{table}
+
+We displayed the samples in a sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
+From top to bottom, we have samples 1 to 3 in the sample holder.
+
+\section{Experimental Procedure}
+
+First of all, we assembled the source holder like it is described in the previous section.
+So that we can do a calibration from the number of counts per channel to energy, we put an additional americium-241 source on the holder.
+Using a ruler, we measured the distance between the samples.
+We put the sample holder on a support and afterwards both of them inside a chamber.
+A vacuum machine was connected to the accelerator to reduce the pressure inside the accelerator.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|w{2cm}|w{4cm}|}
+\hline
+Samples & Distance [cm]  \\ \hline
+Americium & 1.5  \\ \hline
+LiF & 3.3 \\ \hline
+LiAlO2 & 4.8  \\ \hline
+Implanted Li-7 in Al & 6.0  \\ \hline
+\end{tabular}
+\caption{Distance between the samples in the sample holder}
+\label{tab:samples}
+\end{table}
+
+Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the manuals provided.
+Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses.
+Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
+
+In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples.
+Then we turned on the camera next to the samples to observe if it was well adjusted.
+The beam wasn't rightly centred so we adjusted the angles and positions of the lenses until it was in the correct position.
+Finally, we used a collimator with a 2mm radius to collimate the beam.
+
+We used a computer program to obtain the spectra of the different sources.
+To start, we obtained the spectrum of americium-241 in order to do a calibration posteriorly.
+Then we obtained the spectra of the different sources.
+The multichannel analyzer (MCA) utilized to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
+
+After obtaining the lithium spectra, we obtained the spectrum of borum as well. Borum can be detected in the middle of the sources since the encapsulation of the sources is made of boric acid.
+
+eletromagnets adjustment: there are 3 magnets:
+- low energy switch magnet current - 1.7 A;
+- high energy switch 90º magnet current - 13.2 A;
+- second high energy switch magnet current - 14.4 A;
+1315 keV
+
+proton source adjustment - 15 A
+
+Temos um feiexe e conseguimos ver.
+ajustamos o feixe com os campos magenticos e com o steering e com o electrostatic steerer no fim.
+Depois colimamos o nosso feixe através de metodos oftalmologicos.
+(colimador de raio 2mm)
+
+Eletronica de aquisicao de dados a 70V, evoluindo gradualmente.O angulo entre os detetores é 165º (valor q usamos no calculo da energia)
+Offset porta alvos - 15mm
+Escrever a distancia entre as amostras no alvo
+
+Virar po porta alvos de costas para ter o americio241 de frente de modo a calibrar
+
+Desviar a amostra 18 mm de modo a ter a 1 amostra a contar de cima.
+Mudar o modo de aquisição de tempo para carga (tempo para a fonte radioativa para calibrar e carga para as fontes com litio para o detetar)
+O encapsulamento das amostras é feito com acido borico pelo que a detecao de boro é no meio das amostras (pastilhas).
+
+We left CTN at around 16:45 PM.
+
+\printbibliography
+\nocite{*}
+
 \end{document}

### 2023-12-18 analise.py
Olá Curvo

@@ -185,31 +185,9 @@ plt.xlim(500, 700)
 plt.savefig(IMAGE_PATH/'TT_21_Chn0.png')
 plt.show()
 
-# Gaussian fit for the peaks (soma 3 de gaussianas)
-def gaussian(x, a1, x01, sigma1, a2, x02, sigma2, a3, x03, sigma3, c):
-    return a1*np.exp(-(x-x01)**2/(2*sigma1**2)) + a2*np.exp(-(x-x02)**2/(2*sigma2**2)) + a3*np.exp(-(x-x03)**2/(2*sigma3**2)) + c
-
-# Peaks
-x = np.linspace(500, 700, 200)
-y = TT21_Chn0[500:700]
-plt.plot(x, y, label='data') 
-# curve fit
-popt, pcov = curve_fit(gaussian, x, y, p0=[100, 550, 1, 100, 580, 1, 100, 610, 1, 0])
-plt.plot(x, gaussian(x, *popt), label='fit')
-plt.legend()
-plt.title('TT_21')
-plt.xlabel('Channel')
-plt.ylabel('Counts')
-plt.yscale('log')
-plt.xlim(500, 700)
-#plt.savefig(IMAGE_PATH/'TT_21_Chn0_fit.png')
-plt.show()
 
 
 
-
-# valores de energia para os 3 picos
-energy = np.array(5156.59, 5485.56, 5804.82)
 
 
 

### 2024-01-05 calib.py
Print vallues calib.py

@@ -5,6 +5,7 @@ from scipy.optimize import curve_fit
 from pathlib import Path
 from reader import reader
 import pandas as pd
+from tabulate import tabulate
 
 # -----------------------------------------------------------------------------
 # Define paths
@@ -62,6 +63,7 @@ x = np.array(range(500, 700))
 y = TT21_Chn0[500:700]
 # curve fit
 popt0, pcov0 = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5, 0])
+ercov0 = np.sqrt(np.diag(pcov0))
 
 plt.figure()
 plt.plot(TT21_Chn0, label='Channel 0')
@@ -76,12 +78,22 @@ plt.ylim(1, 1e2)
 plt.savefig(IMAGE_PATH/'TT_21_Chn0.png')
 # plt.show()
 
+
+info = [
+    ['a1', 'x01', 'sigma1', 'a2', 'x02', 'sigma2', 'a3', 'x03', 'sigma3', 'c'],
+    [f'{popt0[0]:.2f} ± {ercov0[0]:.2f}', f'{popt0[1]:.2f} ± {ercov0[1]:.2f}', f'{popt0[2]:.2f} ± {ercov0[2]:.2f}', f'{popt0[3]:.2f} ± {ercov0[3]:.2f}', f'{popt0[4]:.2f} ± {ercov0[4]:.2f}', f'{popt0[5]:.2f} ± {ercov0[5]:.2f}', f'{popt0[6]:.2f} ± {ercov0[6]:.2f}', f'{popt0[7]:.2f} ± {ercov0[7]:.2f}', f'{popt0[8]:.2f} ± {ercov0[8]:.2f}', f'{popt0[9]:.2f} ± {ercov0[9]:.2f}']
+]
+
+print(tabulate([['Channel 0']]))
+print(tabulate(info, headers='firstrow', tablefmt='fancy_grid'))
+
 # -----------------------------------------------------------------------------
 # Channel 1
 x = np.array(range(500, 700))
 y = TT21_Chn1[500:700]
 # curve fit
 popt1, pcov1 = curve_fit(gaussian, x, y, p0=[90, 565, 5, 70, 605, 5, 70, 640, 5, 0])
+ercov1 = np.sqrt(np.diag(pcov1))
 
[…]

### 2024-01-08 text/paper/paper.tex
relatório

@@
 \usepackage{makeidx}
+\usepackage{mathtools}
 % \usepackage{tocbibind}
@@
 
+\section{Main Goal}
+    \label{sec:maingoal}
+
+Our main goal is to develop a scientific research project alongside three senior investigators from CTN (Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
+Our project will be focused on the detection of Li in technological materials through nuclear reactions.
+The main motivation for this project is to study and develop a detection technique for elements such as Li with high sensitivity.
+
+\section{Scientific Background}
+    \label{sec:scientificbackground}
+
+The development of this scientific procedure is based on a physics phenomenon: nuclear reactions.
+We have an accelerator in order to allow protons with specific energy to collide with the sample whose composition we are interested in studying.
+The protons will undergo a nuclear reaction with the ones in the samples and byproducts of these reaction will be released.
+These byproducts are charged particles and we can detect their energy spectrum.
+Through the analysis of it, we want to infer what were the elements with whom they reacted on the sample and what is their composition percentage.
+
+\subsection{Nuclear Reaction}
+    \label{sec:nuclearreactions}
+
+The main nuclear reaction in which we are interested is the one between $\ce{^7Li}$ and an incoming proton on the beam:
+
+\begin{equation}
+  \ce{^7Li + ^1H -> ^8Be -> 2\alpha}
+\label{eq:reactionLi}
+\end{equation}
+
+Before the emission of the reaction products, the reaction leads to the formation of a $\ce{^8Be}$ compound nucleus (intermediate state), which then rapidly decays into two alpha particles.
+These are the particles we detect.
+An alpha particle is a helium nucleus, composed of two protons and two neutrons.
+
+\paragraph{} Some of the samples we are going to work with also include boron and according to the beam energy we use it is possible that the protons lead to a nuclear reaction with the boron atoms in the samples.
+Therefore, the $\ce{^{11}B}$ reaction with a proton is as follows:
+
+\begin{equation}
+  \ce{ ^{11}B + ^1H -> ^{12}C -> ^8Be + \alpha}
+\label{eq:reactionB}
+\end{equation}
+
+The reaction first leads to the formation of a $\ce{^{12}C}$ compound nucleus which decays into one alpha particle and a $\ce{^8Be}$ nucleus.
+The $\alpha$ particle can be detected with more or less energy according to the energy level on which the $\ce{^8Be}$ nucleus is emitted.
+If the latter is on the fundamental state, the $\alpha$ particle will have more energy and if it is on a higher state, the $\alpha$ particle will have less energy.
+
+\subsection{Energy Peaks and Energy Profile}
+    \label{sec:energy_profile_peak}
+
+In what comes to the phenomena we expect to see in the collected energy spectrum's, the first one are the energy values themselves.
+Both the alpha particles and the other byproducts are emitted with a specific energy just after the nuclear reaction, and this energy can be computed according to the reaction itself and also to the beam's energy.
+This calculus will be done in section \ref{sec:theoreticalvalues}.
+\paragraph{}
+Firstly, we expect to see intensity peaks on the spectrum on these specific energies.
+However, we must take into account the fact that before arriving at the detector, the emitted particles have to go through the other atoms on the sample.
+As an incident proton comes in, it might react with another proton from the material that sits on its surface but it can also react with one inside it.
+The byproducts produced inside the sample initially have the energy we compute but do lose part of it as they collide with other particles before reaching the detector.
+Basically, we must consider the sample's thickness.
+The bigger the depth at which the reaction takes place, the smaller the energy with whom they are detected because more of it gets dissipated.
+Therefore, this will result in an energy profile - the energy peak will be gradually preceded by less intense peaks due to particles that lost energy on their way to the detector.
+It is less likely for the particles that are emitted inside the sample to reach the detector than for those who were emitted on its surface since on each collision there is a chance that it's direction might change.
+This is why the intensity on the energy profile as a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction
+\paragraph{}
+We expect this phenomena to be particularly significant in samples where the atom that takes part in the reaction can be found not only at the surface of the sample but also in its depth.
+Therefore, in implanted samples where the substrate is only a layer, the energy profile will be less evident.
+
+\subsection{Elastic Back-scattering}
+
+One other feature that might occur is elastic back-scattering.
+This happens when the incoming proton from the accelerator is elastically reflected by the sample and then detected.
+After the reflection, the energy is the same and this means that there will be an intensity peak located at the selected beam energy.
+
+\subsection{Rutherford Back scattering}
+
+Rutherford back-scattering is a phenomena that occurs when the incoming proton is reflected by the sample but, in this case, the energy is not the same.
+This means that the proton loses energy on the reflection and therefore the intensity peak will be located at a lower energy than the selected beam energy.
+
+\section{Experimental Procedure}
+    \label{sec:experimentalprocedure}
+
+\subsection{Work Plan}
+    \label{sec:workplan}
+
+Obtaining the energy spectrum of the particles resulting from the nuclear reactions between the protons and the samples will take the following steps:
+
+\begin{enumerate}
+  \item Accelerator Startup - beam energy increase;
+  \item Sample Preparation;
+  \item Channel-energy calibration of the detection system with a well-known triple $\alpha$ source;
+  \item Lenses and electromagnets set up and adjustment;
+  \item Obtaining the energy spectrum of all samples;
+\end{enumerate}
+
+\subsection{Experimental Set-Up}
+    \label{sec:setup}
+
+The experimental set-up consists of a Tandem accelerator, a beam line with electromagnets and a chamber with both the samples and the detectors.
+
+Theoretically, the accelerator produces a beam of protons with a kinetic energy of up to 5 MeV, which is deviated by the steering electromagnet before hitting the sample.
+The electromagnet deflect particles with different angles according to their charge-to-mass ratio and, with that, allows us to remove any unwanted particles from the beam and only let the protons pass through.
+
+The detectors used, which are silicon-based, detect charged particles through the ionisation they produce in the silicon.
+
+In addition, the detectors are connected to a electronic system that amplifies the signal and sends it to a computer where the spectra can be seen.
+
+The beam line is under a vacuum of around $10^{-7}$ mbar in order to avoid the protons colliding with other particles and losing energy.
+They reach the sample with the specified kinetic energy and cause a nuclear reaction whose products are the charged particles detected.
+We aim to analyse the energy spectrum of these particles.
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.15]{images/scheme.jpeg}
+    \captionof{figure}{Experimental Set-up}
+\end{center}
+
+\subsection{Beam Energy Selection}
+    \label{sec:beamenergyselection}
+
+In order to choose the energy of the beam we need to take into account the cross section of the reaction between the protons and the Li-7.
+We collected data from the IBANDL database for this reaction in a range of [0.5,7] MeV and plotted it in the following graph:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.17]{images/Li_crosssection_energy.jpeg}
+    \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy}
+\end{center}
+
+We are only interested in the energy range of [0.5,5] MeV because the energy of the beam is limited by the accelerators range.
+As we can see in the above figure, the cross section of the reaction is maximum in the [2.5, 3] MeV range.
+Therefore, since we want to detect the Li-7 in the samples, we should choose the energy of the beam to be in this range.
+\paragraph{}
+However, practically, the accelerator cannot reach a beam within the [2.5, 3] MeV energy range.
+Increasing beam energies take exponentially more time to be set up.
+Also, we are interested in the possible boron identification in some of the used samples.
+Boron does not react with such high energies but a lower energy beam is effective in doing that.
+Having this into consideration, we choose a beam energy of $1.3$ MeV for our experimental procedure.
+
+\subsection{NRA Calculations - Theoretical Values}
+    \label{sec:theoreticalvalues}
+
+After concluding that the best beam energy to work within the previously described conditions is $1.3$ MeV, we now need to obtain theoretical values for their energy.
+In order to do this, we will use the NRA calculator \cite{NRAEnergyCalc}.
+This is an online calculator developed by investigators at CTN that allows researchers to compute the $Q$ value of a reaction as well as the energy values and scattering angles of the particles these reactions emit.
+Considering an incident beam whose protons have $E_p = 1.3$ MeV and a scattering angle of $\theta = 165 ^{\circ}$ of our detector we computed the following values:
+
+\paragraph{Lithium} Firstly, for the reaction \ref{eq:reactionLi}:
+
+\begin{equation}
+\begin{split}
+  E_{\alpha_{0_1}} &= 7.6298 \text{ MeV} \\
+  E_{\alpha_{0_2}} &= 11.0314 \text{ MeV} \\
+  \theta_{\alpha_{0_1}} &= 165 ^{\circ} \\
+  \theta_{\alpha_{0_2}} &= 12 ^{\circ} \\
+  Q (\ce{^7Li}) &= 17.346 \text{ MeV} \\
+\end{split}
+\label{eq:energiesli}
+\end{equation}
+
+$E_{\alpha_{0_1}}$ and $E_{\alpha_{0_2}}$ are the energies of both alpha particles considering that they are in the fundamental state.
+In this case, the only other option would be for the $\ce{^8Be}$ to be formed in an excited state and therefore give origin to alpha particles also in an excited state.
+There is not enough energy for this to happen since the difference between the ground state and the the first state in an helium nuclei is very big, around $INSERT$ MeV.
+\paragraph{}
+Also, we can see that the scattering angle of the second alpha particle is around $12 ^{\circ}$ meaning that it is different from the detector's angle and will not be detected.
+To sum up, if there is lithium in any sample that reacts with the incoming protons according to \ref{eq:reactionLi}, we expect to detect alpha particles that were emitted with $7.6298 \text{ MeV}$.
+
+\paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:
+
+\begin{equation}
+\begin{split}
+  E_{\alpha_0} &= 3.7372 \text{ MeV} \\
+  E_{\ce{^8Be}} &= 3.1086 \text{ MeV} \\
+  \theta_{\alpha_0} &= 165 ^{\circ} \\
+  \theta_{\ce{^8Be}} &= 11 ^{\circ}\\
+  Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
+\end{split}
+\label{eq:energiesBr}
+\end{equation}
+
+We need to comment these energies.
+
+\subsection{Procedure}
+    \label{sec:procedure}
+
+We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
+So that we can do a calibration from the number of counts per channel to energy, we put an additional source on the holder just above the lithium fluoride one.
+Using a ruler, we measured the distance between the samples so that afterwards we could control the lift in order to change the sample on the beam's target.
+
+We placed the sample holder on a support and afterwards both of them inside the chamber. A vacuum machine was connected to the accelerator to reduce the pressure inside both the chamber and the beam.
+
+Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the provided instruction manual.
+Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses.
+Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
+
+In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples.
+Finally, we used a collimator with a 2mm radius to collimate the beam.
+
+Finally, we used a computer program to obtain the spectra of the different sources.
+The multichannel analyzer (MCA) utilized to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
+
+\section*{Analysis}
+
+\section{Calibration}
+    \label{sec:calibration}
+
+For the calibration, we used a triple $\alpha$ source. 
+In order to do this procedure, for each of spectrum relevant peaks, we need to find the channel with the maximum number of counts.
+Then, we find the theoretical values for these peaks and, using both the experimental values for the peaks in channel units and the theoretical ones in MeV, we want to obtain a mathematical relation that allows us to convert channel values into energy values.
+To accomplish this, we will make a linear regression with the energy-channel data.
+
+The spectrum obtained for the triple source was the following:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21.png}
+    \captionof{figure}{Spectrum of the Calibration source}
+\end{center}
+
+We separated the spectra of channels 0, 1 and 2 and used a triple gaussian sum function to fit on the three energy peaks.
+According to theory, the first peak corresponds to the decay of plutonium into uranium, the second of americium into neptunium and the third peak of curium into plutonium.
+The decay equations are:
+
+\begin{equation}
+    \ce{ ^{239}Pu -> ^{235}U + \alpha} 
+\end{equation}
+\begin{equation}
+    \ce{ ^{241}Am -> ^{237}Np + \alpha}
+\end{equation} 
+\begin{equation}
+    \ce{ ^{244}Cm -> ^{240}Pu + \alpha}
+\end{equation}
+
+The equation used for the triple gaussian fit is:
+
+\begin{equation}
+    \begin{split}
+    f(x) = a_0\exp{\left(-\frac{(x-\mu_0)^2}{\sigma_0}\right)} &+ a_1\exp{\left(-\frac{(x-\mu_1)^2}{\sigma_1}\right)} + \\
+    &+ a_2\exp{\left(-\frac{(x-\mu_2)^2}{\sigma_2}\right)} + c
+    \end{split}
+    \label{eq:califit}
+\end{equation}
+
+Using this function, we obtained the following fits for channels 0, 1 and 2:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn0.png}
+    \captionof{figure}{Channel 0 spectrum fit with [\ref{eq:califit}]}
+\end{center}
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn1.png}
+    \captionof{figure}{Channel 1 spectrum fit with [\ref{eq:califit}]}
+\end{center}
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn2.png}
+    \captionof{figure}{Channel 2 spectrum fit with [\ref{eq:califit}]}
+\end{center}
+
+The fit parameters for means $\mu_0$, $\mu_1$ and $\mu_2$ in equation [\ref{eq:califit}] are the ones that represent the energy value of each peak in channels.
+The computed values for these parameters, as well as the theoretical values of the energy in MeV for the peaks, are organised in the following tables:
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
+\hline
+Peak & Channel 0 [Chn] & Theoretical Energy [keV] \\ \hline
+1\textsuperscript{st} & $ \mu_0 = 557.00 \pm 0.11 $ & 5156.59 \\ \hline
+2\textsuperscript{nd} & $ \mu_1 = 593.14 \pm 0.10 $ & 5485.56 \\ \hline
+3\textsuperscript{rd} & $ \mu_2 = 627.44 \pm 0.10 $ & 5804.82 \\ \hline
+\end{tabular}
+\caption{Calibration Values for Channel 0}
+\label{tab:calibration0}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
+\hline
+Peak & Channel 1 [Chn] & Theoretical Energy [keV] \\ \hline
+1\textsuperscript{st} & $ \mu_0 = 568.52 \pm 0.17 $ & 5156.59 \\ \hline
+2\textsuperscript{nd} & $ \mu_1 = 605.39 \pm 0.17 $ & 5485.56 \\ \hline
+3\textsuperscript{rd} & $ \mu_2 = 640.77 \pm 0.17 $ & 5804.82 \\ \hline
+\end{tabular}
+\caption{Calibration Values for Channel 1}
+\label{tab:calibration1}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
+\hline
+Peak & Channel 2 [Chn] & Theoretical Energy [keV] \\ \hline
+1\textsuperscript{st} & $ \mu_0 = 563.09 \pm 0.28 $ & 5156.59 \\ \hline
+2\textsuperscript{nd} & $ \mu_1 = 600.15 \pm 0.26 $ & 5485.56 \\ \hline
+3\textsuperscript{rd} & $ \mu_2 = 634.56 \pm 0.36 $ & 5804.82 \\ \hline
+\end{tabular}
+\caption{Calibration Values for Channel 2}
+\label{tab:calibration2}
+\end{table}
+
+Using the values in these three tables we now need to compute a calibration equation for each of the channels.
+We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn0_calib.png}
+    \captionof{figure}{Calibration Channel 0}
+\end{center}
+
+The calibration equation for channel 0 is:
+
+\begin{equation}
+    \begin{split}
+        E[\text{MeV}] &= m_0 \times E[\text{Chn}] + b_0 \\
+        m_0 &= 9.20 \pm 0.06 \\
+        b_0 &= 30.22 \pm 34.64
+    \end{split}
+    \label{eq:calib0}
+\end{equation}
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn1_calib.png}
+    \captionof{figure}{Calibration Channel 1}
+\end{center}
+
+The calibration equation for channel 1 is:
+
+\begin{equation}
+    \begin{split}
+        E[\text{MeV}] &= m_1 \times E[\text{Chn}] + b_1 \\
+        m_1 &= 8.97 \pm 0.03 \\
+        b_1 &= 55.40 \pm 17.53
+    \end{split}
+    \label{eq:calib1}
+\end{equation}
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{images/TT_21_Chn2_calib.png}
+    \captionof{figure}{Calibration Channel 2}
+\end{center}
+
+The calibration equation for channel 2 is:
+
+\begin{equation}
+    \begin{split}
+        E[\text{MeV}] &= m_2 \times E[\text{Chn}] + b_2 \\
+        m_2 &= 9.07 \pm 0.12 \\
+        b_2 &= 48.17 \pm 69.69
+    \end{split}
+    \label{eq:calib2}
+\end{equation}
+
+\section{ Lithium Fluoride Sample}
+
+$\ce{LiF}$
+
+\section{ Lithium Aluminate Sample}
+
+$\ce{LiAlO_2}$
+
+\subsection{Energy Profile}
+
+\section{Implanted Sample}
+
+\section{Conclusion}
+
+\section{Perguntas}
+
+\begin{enumerate}
+    \item Para a calibração, colocamos os valores e as incertezas de todos os 10 parâmetros para cada channel ou apenas colocamos as médias ?
+\end{enumerate}
+
+\section{}
+
 \printbibliography

### 2024-01-08 analise.py
Edited Images

@@ -108,37 +108,37 @@ TT27_Chn2 = reader(TT_27_PATH / 'CH2@N6781_21198_Espectrum_TT_27_20231207_160722
 # plt.savefig(IMAGE_PATH/'TT_22.png')
 # plt.show()
 
-# # TT_23 -> 1024 bins, 2mC (LiF)
-# plt.figure()
-# plt.plot(TT23_Chn0, label='Chnannel 0')
-# plt.plot(TT23_Chn1, label='Channel 1')
+# TT_23 -> 1024 bins, 2mC (LiF)
+plt.figure()
+plt.plot(TT23_Chn0, label='Channel 0')
+plt.plot(TT23_Chn1, label='Channel 1')
 # plt.plot(TT23_Chn2, label='Channel 2')
-# plt.legend()
-# plt.title('TT_23')
-# plt.yscale('log')
-# plt.xlabel('Channel')
-# plt.ylabel('Counts')
-# plt.savefig(IMAGE_PATH/'TT_23.png')
-# plt.show()
+plt.legend()
+plt.title('Sample with LiF')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.savefig(IMAGE_PATH/'SampleLiF.png')
+plt.show()
 
-# # TT_24 -> 1024 bins, 2mC (LiAlO2)
-# plt.figure()
-# plt.plot(TT24_Chn0, label='Chnannel 0')
-# plt.plot(TT24_Chn1, label='Channel 1')
+# TT_24 -> 1024 bins, 2mC (LiAlO2)
+plt.figure()
+plt.plot(TT24_Chn0, label='Channel 0')
+plt.plot(TT24_Chn1, label='Channel 1')
 # plt.plot(TT24_Chn2, label='Channel 2')
-# plt.legend()
-# plt.title('TT_24')
-# plt.yscale('log')
-# plt.xlabel('Channel')
-# plt.ylabel('Counts')
-# plt.savefig(IMAGE_PATH/'TT_24.png')
-# plt.show()
+plt.legend()
+plt.title('Sample with LiAlO2')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.ylabel('Counts')
+plt.savefig(IMAGE_PATH/'SampleLiAlO2.png')
+plt.show()
 
 # # TT_25 -> 1024 bins, 2mC (Li Implantado em Al)
 # plt.figure()
 # plt.plot(TT25_Chn0, label='Chnannel 0')
[…]

### 2024-01-09 analise.py
CURVO

@@ -193,3 +193,4 @@ plt.show()
 
 
 
+

### 2024-01-19 text/paper/paper.tex
paper

@@
 
-\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}
+\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{images/IST_Logo.png}
 % \\ \vspace{1mm} LFEUI \vspace{1mm} 
@@
 
-In what comes to the phenomena we expect to see in the collected energy spectrum's, the first one are the energy values themselves.
-Both the alpha particles and the other byproducts are emitted with a specific energy just after the nuclear reaction, and this energy can be computed according to the reaction itself and also to the beam's energy.
-This calculus will be done in section \ref{sec:theoreticalvalues}.
+Regarding the phenomena expected in the collected energy spectrum, the primary aspect involves the energy values.
+Both the alpha particles and other byproducts are emitted with distinct energies immediately following the nuclear reaction.
+These energies can be calculated based on the reaction and the beam's energy.
+The computation process will be detailed in section \ref{sec:theoreticalvalues}.
 \paragraph{}
@@
 It is less likely for the particles that are emitted inside the sample to reach the detector than for those who were emitted on its surface since on each collision there is a chance that it's direction might change.
-This is why the intensity on the energy profile as a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction
+This is why the intensity on the energy profile as a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction.
 \paragraph{}
@@
 \subsection{Elastic Back-scattering}
+    \label{sec:elastic_back}
 
-One other feature that might occur is elastic back-scattering.
-This happens when the incoming proton from the accelerator is elastically reflected by the sample and then detected.
-After the reflection, the energy is the same and this means that there will be an intensity peak located at the selected beam energy.
+Another possible occurrence is elastic back-scattering, where the incoming proton from the accelerator is elastically reflected back by the sample and subsequently detected.
+During this reflection, the energy remains unchanged, resulting in an intensity peak positioned close to the beam energy selected for our experiment.
 
-\subsection{Rutherford Back scattering}
-
-Rutherford back-scattering is a phenomena that occurs when the incoming proton is reflected by the sample but, in this case, the energy is not the same.
-This means that the proton loses energy on the reflection and therefore the intensity peak will be located at a lower energy than the selected beam energy.
+o backscattering é todo elástico e um caso particular deste é o de rutherford.
+O caso particular de rutherford considera os atomos como esferas perfeitas para calcular a secção eficaz deste fenomeno.
+Em geral, as secções eficazes do backscattering obtém-se experimentalmente e não usamos as de rutherford.
+O que interessa é que o backscattering elastico como fenomeno e não as suas diferentes secções eficazes.
+Pileup - o detetor recebe duas partículas de backscattering em simultaneo e assum que é so uma e por isso soma uma contagem com o dobro da energia - A particula emitida no backscattering pode ter diferentes valores de energia de acordo com a colisao elastica entre os dois atomos (a mecanica diz nos como calular a energia com que cada átomo, o incidente e o da amosta à superfície, fica depois da colisão) e por isso a soma da energia de duas também pode estar num dado intervalo.
+O pico inicial é do ruido eletronico e por isso desconsideramos.
+Usar detector em vez de channel.
+A secção eficaz da reação nuclear com o litio varia com a energia do protão incidente.
+O protão incidente perde energia à medida que percorre a amostra devido ao stopping power da amostra em si e por isso a secção eficaz de reacções que ocorrem mais no interior da amostra é diferente das que ocorrem à sua superfície.
 
@@
     \centering
-    \includegraphics[scale = 0.6]{images/TT_21.png}
+    \includegraphics[scale = 0.35]{images/Calib.png}
     \captionof{figure}{Spectrum of the Calibration source}
@@
 
-We separated the spectra of channels 0, 1 and 2 and used a triple gaussian sum function to fit on the three energy peaks.
+We separated the spectra of channeldetectors 0, 1 and 2 (2 corresponding to the bottom one) and used a triple gaussian sum function to fit on the three3 energy peaks.
 According to theory, the first peak corresponds to the decay of plutonium into uranium, the second of americium into neptunium and the third peak of curium into plutonium.
@@
 
-Using this function, we obtained the following fits for channels 0, 1 and 2:
+We decided to do the analysis only with "detector 0". This anlaysis could be further corroborated using the "detector 1" data.
+Using the triple gaussian, we obtained the following fits for "detector 0":
 
@@
 
-\begin{center}
-    \label{TT_21}
-    \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn1.png}
-    \captionof{figure}{Channel 1 spectrum fit with [\ref{eq:califit}]}
-\end{center}
+%\begin{center}
+%    \label{TT_21}
+%    \centering
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn1.png}
+%    \captionof{figure}{Channel 1 spectrum fit with [\ref{eq:califit}]}
+%\end{center}
+%
+%\begin{center}
+%    \label{TT_21}
+%    \centering
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn2.png}
+%    \captionof{figure}{Channel 2 spectrum fit with [\ref{eq:califit}]}
+%\end{center}
 
-\begin{center}
-    \label{TT_21}
-    \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn2.png}
-    \captionof{figure}{Channel 2 spectrum fit with [\ref{eq:califit}]}
-\end{center}
-
-The fit parameters for means $\mu_0$, $\mu_1$ and $\mu_2$ in equation [\ref{eq:califit}] are the ones that represent the energy value of each peak in channels.
+The fit parameters for means $\mu_0$, $\mu_1$ and $\mu_2$ in equation [\ref{eq:califit}] represent the energy value of each peak in channels.
 The computed values for these parameters, as well as the theoretical values of the energy in MeV for the peaks, are organised in the following tables:
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
-\hline
-Peak & Channel 1 [Chn] & Theoretical Energy [keV] \\ \hline
-1\textsuperscript{st} & $ \mu_0 = 568.52 \pm 0.17 $ & 5156.59 \\ \hline
-2\textsuperscript{nd} & $ \mu_1 = 605.39 \pm 0.17 $ & 5485.56 \\ \hline
-3\textsuperscript{rd} & $ \mu_2 = 640.77 \pm 0.17 $ & 5804.82 \\ \hline
-\end{tabular}
-\caption{Calibration Values for Channel 1}
-\label{tab:calibration1}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
-\hline
-Peak & Channel 2 [Chn] & Theoretical Energy [keV] \\ \hline
-1\textsuperscript{st} & $ \mu_0 = 563.09 \pm 0.28 $ & 5156.59 \\ \hline
-2\textsuperscript{nd} & $ \mu_1 = 600.15 \pm 0.26 $ & 5485.56 \\ \hline
-3\textsuperscript{rd} & $ \mu_2 = 634.56 \pm 0.36 $ & 5804.82 \\ \hline
-\end{tabular}
-\caption{Calibration Values for Channel 2}
-\label{tab:calibration2}
-\end{table}
+%\begin{table}[H]
+%\centering
+%\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
+%\hline
+%Peak & Channel 1 [Chn] & Theoretical Energy [keV] \\ \hline
+%1\textsuperscript{st} & $ \mu_0 = 568.52 \pm 0.17 $ & 5156.59 \\ \hline
+%2\textsuperscript{nd} & $ \mu_1 = 605.39 \pm 0.17 $ & 5485.56 \\ \hline
+%3\textsuperscript{rd} & $ \mu_2 = 640.77 \pm 0.17 $ & 5804.82 \\ \hline
+%\end{tabular}
+%\caption{Calibration Values for Channel 1}
+%\label{tab:calibration1}
+%\end{table}
+%
+%\begin{table}[H]
+%\centering
+%\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
+%\hline
+%Peak & Channel 2 [Chn] & Theoretical Energy [keV] \\ \hline
+%1\textsuperscript{st} & $ \mu_0 = 563.09 \pm 0.28 $ & 5156.59 \\ \hline
+%2\textsuperscript{nd} & $ \mu_1 = 600.15 \pm 0.26 $ & 5485.56 \\ \hline
+%3\textsuperscript{rd} & $ \mu_2 = 634.56 \pm 0.36 $ & 5804.82 \\ \hline
+%\end{tabular}
+%\caption{Calibration Values for Channel 2}
+%\label{tab:calibration2}
+%\end{table}
 
@@
     \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn0_calib.png}
-    \captionof{figure}{Calibration Channel 0}
+    \includegraphics[scale = 0.35]{images/Chn0_calib_with_Backscattering.png}
+    \captionof{figure}{Calibration "detector 0"}
 \end{center}
 
-The calibration equation for channel 0 is:
+The calibration equation for "detector 0" is:
 
@@
     \begin{split}
-        E[\text{MeV}] &= m_0 \times E[\text{Chn}] + b_0 \\
-        m_0 &= 9.20 \pm 0.06 \\
-        b_0 &= 30.22 \pm 34.64
+        E[\text{keV}] &= m_0 \times E[\text{Chn}] + b_0 \\
+        m_0 &= 9.01 \pm 0.02 \\
+        b_0 &= 142.83 \pm 8.99
     \end{split}
@@
 
+%\begin{center}
+%    \label{TT_21}
+%    \centering
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn1_calib.png}
+%    \captionof{figure}{Calibration Channel 1}
+%\end{center}
+%
+%The calibration equation for channel 1 is:
+%
+%\begin{equation}
+%    \begin{split}
+%        E[\text{keV}] &= m_1 \times E[\text{Chn}] + b_1 \\
+%        m_1 &= 8.97 \pm 0.03 \\
+%        b_1 &= 55.40 \pm 17.53
+%    \end{split}
+%    \label{eq:calib1}
+%\end{equation}
+%
+%\begin{center}
+%    \label{TT_21}
+%    \centering
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn2_calib.png}
+%    \captionof{figure}{Calibration Channel 2}
+%\end{center}
+%
+%The calibration equation for channel 2 is:
+%
+%\begin{equation}
+%    \begin{split}
+%        E[\text{keV}] &= m_2 \times E[\text{Chn}] + b_2 \\
+%        m_2 &= 9.07 \pm 0.12 \\
+%        b_2 &= 48.17 \pm 69.69
+%    \end{split}
+%    \label{eq:calib2}
+%\end{equation}
+%
+\section{Alfpha Peaks}
+For the analysis, we began by calculating the theoretical energy values of the two alfphas.
+
 \begin{center}
-    \label{TT_21}
+    \label{alfa_peaks}
     \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn1_calib.png}
-    \captionof{figure}{Calibration Channel 1}
+    \includegraphics[scale = 0.35]{images/AlphaPeaks.png}
+    \captionof{figure}{Theoretical energy values of the two alfphas}
 \end{center}
 
-The calibration equation for channel 1 is:
+The energies obtained were 7629.80 kEv and 11031.44 kEv. (FALTA EXPLICAR PERCENTAGENS)
 
-\begin{equation}
-    \begin{split}
-        E[\text{MeV}] &= m_1 \times E[\text{Chn}] + b_1 \\
-        m_1 &= 8.97 \pm 0.03 \\
-        b_1 &= 55.40 \pm 17.53
-    \end{split}
-    \label{eq:calib1}
-\end{equation}
+\section{Lithium Fluoride Sample}
+
+For the LiF sample, we overlapped the expected $\alpha$ energy getting the following spectrum:
 
@@
     \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn2_calib.png}
-    \captionof{figure}{Calibration Channel 2}
+    \includegraphics[scale = 0.35]{images/OverlapLiF_NC.png}
+    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
 \end{center}
 
-The calibration equation for channel 2 is:
+The experimental $\alpha_0$ is extremely close to the theoretical value, therefore we can confirm the presence of lithium in this sample. Analysing this spectrum further:
 
-\begin{equation}
-    \begin{split}
-        E[\text{MeV}] &= m_2 \times E[\text{Chn}] + b_2 \\
-        m_2 &= 9.07 \pm 0.12 \\
-        b_2 &= 48.17 \pm 69.69
-    \end{split}
-    \label{eq:calib2}
-\end{equation}
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.4]{images/FullAnalysisLiF.png}
+    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
+\end{center}
 
-\section{ Lithium Fluoride Sample}
+There are three noticeable structures in this spectrum. 
+The first one is a peak around channel 0, followed by a less intense one around 100 Chn and a small plateau between 150 and 250 Chn. 
+Lastly, there is an evident energy profile that starts at around 300 Chn and goes up to 900 Chn with increasing intensity.
 
-$\ce{LiF}$
+\paragraph*{}
+
+After identifying these structures we used the previous calibration [\ref{sec:calibration}] to convert the channel values into energy values so that we could more effectively analyse the spectrum and its peaks.
 
@@
 
-$\ce{LiAlO_2}$
+We obtained the following spectrum for the lithium aluminate sample:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.35]{images/OverlapLiAlO2.png}
+    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
+\end{center}
 
@@
 \section{Implanted Sample}
+
+The obtained spectrum for the implanted sample was the following:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.35]{images/OverlapImplanted.png}
+    \captionof{figure}{Spectrum of the implanted sample}
+\end{center}
+
+\section{Quantitative Analysis}
+According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:
+\begin{equation}
+    N\tau_i=\frac{H_i}{\sigma (E_i)\Omega Q}
+\end{equation}
+\begin{equation}
+    N\Delta x=\frac{A}{\sigma (E_i)\Omega Q}
+\end{equation}
+
+Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \cdot$\#channels.\\
+$H_i$ is the number of counts in a FAT MUDAR spectrum peak (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.
+
+With this we get:\\
+APROXIMACOES:\\
+implanted: $1\times10^{17}$ atoms/cm$^2$\\
+$\ce{LiAlO2}$: $3\times10^{18}$ atoms/cm$^2$\\
+etc...\\
 

### 2024-01-19 full_analysis.py

@@ -17,7 +17,7 @@ import pandas as pd
 SHOW = False
 
 # Save Pictures 
-SAVE = False
+SAVE = True
 
 # Put Titles
 TITLES = False
@@ -94,9 +94,9 @@ def Detector_Calibration(detector : int = 0, bins : int = 1024) -> np.ndarray:
 # it might be Lithium
 # -----------------------------------------------------------------------------
 plt.figure(figsize=(10, 10))
-plt.plot(LIF_Chn0, label='LiF', alpha=0.5)
-plt.plot(LiAlO2_Chn0, label='LiAlO2', alpha=0.5)
-plt.plot(Implanted_Chn0, label='Implanted', alpha=0.5)
+#plt.plot(LIF_Chn0, label='LiF', alpha=0.5)
+#plt.plot(LiAlO2_Chn0, label='LiAlO2', alpha=0.5)
+#plt.plot(Implanted_Chn0, label='Implanted', alpha=0.5)
 plt.plot(Cal_Chn0, label='Calibration')
 plt.legend()
 if TITLES: plt.title('Overlap')
@@ -262,6 +262,7 @@ for i in range(len(alphas)):
 # Guassian for Channel 0
 x = np.array(range(800, 900))
 y = Implanted_Chn0[800:900]
+
 # curve fit
 popt0, pcov0 = curve_fit(gaussian, x, y, p0=[12, 850, 10])
 errors0 = np.sqrt(np.diag(pcov0))
@@ -630,7 +631,7 @@ c0 = params0[1]
 plt.figure(figsize=(10, 10))
 plt.plot(Xpoints, Ypoints, 'o', label='data')
 plt.plot(Xpoints, linear(Xpoints, *params0), label='fit')
-if TITLES: plt.title('Calibration Channel 0')
+if TITLES: plt.title('Calibration "detector 0"')
 plt.legend()
 plt.xlabel('Channel')
 plt.ylabel('Energy [keV]')

### 2024-01-19 full_analysis.py
merda

@@ -17,10 +17,10 @@ import pandas as pd
 SHOW = False
 
 # Save Pictures 
-SAVE = False
+SAVE = True
 
 # Put Titles
-TITLES = False
+TITLES = True
 
 # Define paths
 # data

### 2024-01-19 full_analysis.py
feat(full_analysis.py): area of implanted lithium

@@ -17,7 +17,7 @@ import pandas as pd
 SHOW = False
 
 # Save Pictures 
-SAVE = True
+SAVE = False
 
 # Put Titles
 TITLES = True
@@ -94,9 +94,9 @@ def Detector_Calibration(detector : int = 0, bins : int = 1024) -> np.ndarray:
 # it might be Lithium
 # -----------------------------------------------------------------------------
 plt.figure(figsize=(10, 10))
-#plt.plot(LIF_Chn0, label='LiF', alpha=0.5)
-#plt.plot(LiAlO2_Chn0, label='LiAlO2', alpha=0.5)
-#plt.plot(Implanted_Chn0, label='Implanted', alpha=0.5)
+plt.plot(LIF_Chn0, label='LiF', alpha=0.5)
+plt.plot(LiAlO2_Chn0, label='LiAlO2', alpha=0.5)
+plt.plot(Implanted_Chn0, label='Implanted', alpha=0.5)
 plt.plot(Cal_Chn0, label='Calibration')
 plt.legend()
 if TITLES: plt.title('Overlap')
@@ -782,3 +782,26 @@ plt.ylim(0.1, max(Implanted_Chn0))
 if SAVE: plt.savefig(IMAGE_PATH / 'OverlapImplanted_NC.png', dpi=600)
 plt.grid()
 if SHOW: plt.show()
+
+# ----------------------------------------------------------------------------------------------------------------------
+# Area for Implanted Sample
+# ----------------------------------------------------------------------------------------------------------------------
+# Define the x range
+X = 800
+Y = 950
+total_area_Ch0 = np.sum(Implanted_Chn0[X:Y])
+total_area_Ch1 = np.sum(Implanted_Chn1[X:Y])
+total_area_Ch2 = np.sum(Implanted_Chn2[900:1000])
+# Print on terminal the results
+print(tabulate([['Area of Sample with Implanted Li']], tablefmt='fancy_grid'))
+info = [
+    ['Detector', 'Area in Counts'],
+    [0, f'{total_area_Ch0:.2f}'],
+    [1, f'{total_area_Ch1:.2f}'],
+    [2, f'{total_area_Ch1:.2f}']
+]
+print(tabulate(info, tablefmt='fancy_grid'), end='\n\n')
+
+# Print table for LaTeX
+print('Code for LaTeX\n\n')
+print(tabulate(info, headers='firstrow', tablefmt='latex'), '\n\n')

### 2024-01-19 analise.py
feat(full_analysis.py) : height for LIF and LiAlO2

@@ -1,336 +1,0 @@
-import numpy as np
-import xml.etree.ElementTree as ET
-import matplotlib.pyplot as plt
-from scipy.optimize import curve_fit
-from pathlib import Path
-from reader import reader
-from tabulate import tabulate
-from tools import calibration_Ch0, calibration_Ch1, calibration_Ch2, error_Ch0, error_Ch1, error_Ch2
-from tools import gaussian
-
-# Define paths
-# data
-DATA_PATH = Path('data')
-# data folders
-TT_20_PATH = DATA_PATH / 'TT_20' / 'UNFILTERED'
-TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED'
-TT_22_PATH = DATA_PATH / 'TT_22' / 'UNFILTERED'
-TT_23_PATH = DATA_PATH / 'TT_23' / 'UNFILTERED'
-TT_24_PATH = DATA_PATH / 'TT_24' / 'UNFILTERED'
-TT_25_PATH = DATA_PATH / 'TT_25' / 'UNFILTERED'
-TT_26_PATH = DATA_PATH / 'TT_26' / 'UNFILTERED'
-TT_27_PATH = DATA_PATH / 'TT_27' / 'UNFILTERED'
-# image
-IMAGE_PATH = Path('images')
-IMAGE_PATH.mkdir(exist_ok=True, parents=True)
-
-# Reading the data files
-
-# TT_20
-TT20_Chn0 = reader(TT_20_PATH / 'CH0@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 0)
-TT20_Chn1 = reader(TT_20_PATH /'CH1@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 1)
-TT20_Chn2 = reader(TT_20_PATH /'CH2@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 2)
-
-# TT_21
-TT21_Chn0 = reader(TT_21_PATH / 'CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
-TT21_Chn1 = reader(TT_21_PATH / 'CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
-TT21_Chn2 = reader(TT_21_PATH / 'CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
-
-# TT_22
[…]

### 2024-01-19 calib.py

@@ -1,252 +1,0 @@
-import numpy as np
-import xml.etree.ElementTree as ET
-import matplotlib.pyplot as plt
-from scipy.optimize import curve_fit
-from pathlib import Path
-from reader import reader
-import pandas as pd
-from tabulate import tabulate
-
-# -----------------------------------------------------------------------------
-# Define paths
-# For the calibration we will use the data from the TT_21 file
-DATA_PATH = Path('data') # Path to the data folder
-TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED' # Path to the TT_21 folder
-# Path to the image folder
-IMAGE_PATH = Path('images') 
-IMAGE_PATH.mkdir(exist_ok=True, parents=True)
-# Calibration
-# File where energy values are stored
-ENERGY_FILE = DATA_PATH / 'calibration_energy.csv'
-# Define CSV file to keep the calibration data
-CALIB_FILE = DATA_PATH / 'calibration.csv'
-
-
-# -----------------------------------------------------------------------------
-# TT_21
-TT21_Chn0 = reader(TT_21_PATH/'CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
-TT21_Chn1 = reader(TT_21_PATH/'CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
-TT21_Chn2 = reader(TT_21_PATH/'CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
-
-# TT_21
-plt.figure()
-plt.plot(TT21_Chn0, label='Chnannel 0')
-plt.plot(TT21_Chn1, label='Channel 1')
-plt.plot(TT21_Chn2, label='Channel 2')
-plt.legend()
-plt.title('TT_21')
-plt.yscale('log')
-plt.xlabel('Channel')
-plt.xlim(0, 1024)
-plt.ylabel('Counts')
-plt.savefig(IMAGE_PATH/'TT_21.png')
-# plt.show()
-
-
-# -----------------------------------------------------------------------------
[…]

### 2024-01-19 full_analysis.py

@@ -805,3 +805,27 @@ print(tabulate(info, tablefmt='fancy_grid'), end='\n\n')
 # Print table for LaTeX
 print('Code for LaTeX\n\n')
 print(tabulate(info, headers='firstrow', tablefmt='latex'), '\n\n')
+
+
+# ----------------------------------------------------------------------------------------------------------------------
+# Height for LIF and LiAlO2
+# ----------------------------------------------------------------------------------------------------------------------
+# Find max height 
+LiAlO2_Chn0_max = np.max(LiAlO2_Chn0[760:860])
+LIF_Chn0_max = np.max(LIF_Chn0[760:860])
+
+# Print on terminal the results
+print(tabulate([['Height']], tablefmt='fancy_grid'))
+info = [
+    ['Sample', 'Height in Counts'],
+    ['LiAlO2', f'{LiAlO2_Chn0_max:.2f}'],
+    ['LiF', f'{LIF_Chn0_max:.2f}'],
+]
+print(tabulate(info, tablefmt='fancy_grid'), end='\n\n')
+
+# Print table for LaTeX
+print('Code for LaTeX\n\n')
+print(tabulate(info, headers='firstrow', tablefmt='latex'), '\n\n')
+
+
+

### 2024-01-19 reader.py

@@ -1,15 +1,0 @@
-import xml.etree.ElementTree as ET
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Function to read the data from the file
-def reader (name, Chn):
-    # Parse the XML data from the file
-    tree = ET.parse(name)
-    root = tree.getroot()
-    # Access spectrum data
-    spectrum_data_str = root.find(f".//Spectrum[@id='RadMeasurement-{Chn}_Spectrum-{Chn}']/ChannelData").text
-    # Convert the spectrum data string to a vector of data
-    spectrum_data_array = np.array([float(value) for value in spectrum_data_str.split()])
-
-    return spectrum_data_array

### 2024-01-19 tools.py

@@ -1,70 +1,0 @@
-import numpy as np
-import pandas as pd
-from pathlib import Path
-
-# Path for the caliration csv 
-CALIB_FILE = Path('data/calibration.csv')
-
-
-# Load the calibration data
-calibration_data = pd.read_csv(CALIB_FILE)
-
-# Define Gaussian function
-def gaussian(x, a, x0, sigma, c=0):
-    return a*np.exp(-(x-x0)**2/(2*sigma**2)) + c
-
-# -----------------------------------------------------------------------------
-# Calibration function for channel 0
-m0 = calibration_data['m'].to_numpy()[0]
-b0 = calibration_data['c'].to_numpy()[0]
-error_m0 = calibration_data['error_m'].to_numpy()[0]
-error_b0 = calibration_data['error_c'].to_numpy()[0]
-
-def calibration_Ch0(x):
-    return m0 * x + b0
-
-def inverse_calibration_Ch0(x):
-    return (x - b0) / m0
-
-def inverse_error_Ch0(x):
-    return abs(-(1 / m0)) * error_b0 + abs((b0 - x) / m0 ** 2) * error_m0
-
-def error_Ch0(bins, error_bins):
-    '''Error propagation for the calibration function of channel 0.
-    The error is calculated using the errorpropagation website,
-    absolute error.'''
-    return abs(m0) * error_bins + abs(bins) * error_m0 + error_b0
-
-# Calibration function for channel 1
-m1 = calibration_data['m'].to_numpy()[1]
-b1 = calibration_data['c'].to_numpy()[1]
-error_m1 = calibration_data['error_m'].to_numpy()[1]
-error_b1 = calibration_data['error_c'].to_numpy()[1]
-
-def calibration_Ch1(x):
-    return m1 * x + b1
-
-def inverse_calibration_Ch1(x):
-    return (x - b1) / m1
-
-def error_Ch1(bins, error_bins):
-    return abs(m1) * error_bins + abs(bins) * error_m1 + error_b1
-
-# Calibration function for channel 2
-m2 = calibration_data['m'].to_numpy()[2]
-b2 = calibration_data['c'].to_numpy()[2]
-error_m2 = calibration_data['error_m'].to_numpy()[2]
-error_b2 = calibration_data['error_c'].to_numpy()[2]
-
-def calibration_Ch2(x):
-    return m2 * x + b2
-
-def inverse_calibration_Ch2(x):
[…]

### 2024-01-19 analise.py
All

@@ -1,0 +1,336 @@
+import numpy as np
+import xml.etree.ElementTree as ET
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from pathlib import Path
+from reader import reader
+from tabulate import tabulate
+from tools import calibration_Ch0, calibration_Ch1, calibration_Ch2, error_Ch0, error_Ch1, error_Ch2
+from tools import gaussian
+
+# Define paths
+# data
+DATA_PATH = Path('data')
+# data folders
+TT_20_PATH = DATA_PATH / 'TT_20' / 'UNFILTERED'
+TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED'
+TT_22_PATH = DATA_PATH / 'TT_22' / 'UNFILTERED'
+TT_23_PATH = DATA_PATH / 'TT_23' / 'UNFILTERED'
+TT_24_PATH = DATA_PATH / 'TT_24' / 'UNFILTERED'
+TT_25_PATH = DATA_PATH / 'TT_25' / 'UNFILTERED'
+TT_26_PATH = DATA_PATH / 'TT_26' / 'UNFILTERED'
+TT_27_PATH = DATA_PATH / 'TT_27' / 'UNFILTERED'
+# image
+IMAGE_PATH = Path('images')
+IMAGE_PATH.mkdir(exist_ok=True, parents=True)
+
+# Reading the data files
+
+# TT_20
+TT20_Chn0 = reader(TT_20_PATH / 'CH0@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 0)
+TT20_Chn1 = reader(TT_20_PATH /'CH1@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 1)
+TT20_Chn2 = reader(TT_20_PATH /'CH2@N6781_21198_Espectrum_TT_20_20231207_141538.n42', 2)
+
+# TT_21
+TT21_Chn0 = reader(TT_21_PATH / 'CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
+TT21_Chn1 = reader(TT_21_PATH / 'CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
+TT21_Chn2 = reader(TT_21_PATH / 'CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
+
+# TT_22
[…]

### 2024-01-19 calib.py

@@ -1,0 +1,252 @@
+import numpy as np
+import xml.etree.ElementTree as ET
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from pathlib import Path
+from reader import reader
+import pandas as pd
+from tabulate import tabulate
+
+# -----------------------------------------------------------------------------
+# Define paths
+# For the calibration we will use the data from the TT_21 file
+DATA_PATH = Path('data') # Path to the data folder
+TT_21_PATH = DATA_PATH / 'TT_21' / 'UNFILTERED' # Path to the TT_21 folder
+# Path to the image folder
+IMAGE_PATH = Path('images') 
+IMAGE_PATH.mkdir(exist_ok=True, parents=True)
+# Calibration
+# File where energy values are stored
+ENERGY_FILE = DATA_PATH / 'calibration_energy.csv'
+# Define CSV file to keep the calibration data
+CALIB_FILE = DATA_PATH / 'calibration.csv'
+
+
+# -----------------------------------------------------------------------------
+# TT_21
+TT21_Chn0 = reader(TT_21_PATH/'CH0@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 0)
+TT21_Chn1 = reader(TT_21_PATH/'CH1@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 1)
+TT21_Chn2 = reader(TT_21_PATH/'CH2@N6781_21198_Espectrum_TT_21_20231207_143401.n42', 2)
+
+# TT_21
+plt.figure()
+plt.plot(TT21_Chn0, label='Chnannel 0')
+plt.plot(TT21_Chn1, label='Channel 1')
+plt.plot(TT21_Chn2, label='Channel 2')
+plt.legend()
+plt.title('TT_21')
+plt.yscale('log')
+plt.xlabel('Channel')
+plt.xlim(0, 1024)
+plt.ylabel('Counts')
+plt.savefig(IMAGE_PATH/'TT_21.png')
+# plt.show()
+
+
+# -----------------------------------------------------------------------------
[…]

### 2024-01-19 reader.py

@@ -1,0 +1,15 @@
+import xml.etree.ElementTree as ET
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Function to read the data from the file
+def reader (name, Chn):
+    # Parse the XML data from the file
+    tree = ET.parse(name)
+    root = tree.getroot()
+    # Access spectrum data
+    spectrum_data_str = root.find(f".//Spectrum[@id='RadMeasurement-{Chn}_Spectrum-{Chn}']/ChannelData").text
+    # Convert the spectrum data string to a vector of data
+    spectrum_data_array = np.array([float(value) for value in spectrum_data_str.split()])
+
+    return spectrum_data_array

### 2024-01-19 tools.py

@@ -1,0 +1,70 @@
+import numpy as np
+import pandas as pd
+from pathlib import Path
+
+# Path for the caliration csv 
+CALIB_FILE = Path('data/calibration.csv')
+
+
+# Load the calibration data
+calibration_data = pd.read_csv(CALIB_FILE)
+
+# Define Gaussian function
+def gaussian(x, a, x0, sigma, c=0):
+    return a*np.exp(-(x-x0)**2/(2*sigma**2)) + c
+
+# -----------------------------------------------------------------------------
+# Calibration function for channel 0
+m0 = calibration_data['m'].to_numpy()[0]
+b0 = calibration_data['c'].to_numpy()[0]
+error_m0 = calibration_data['error_m'].to_numpy()[0]
+error_b0 = calibration_data['error_c'].to_numpy()[0]
+
+def calibration_Ch0(x):
+    return m0 * x + b0
+
+def inverse_calibration_Ch0(x):
+    return (x - b0) / m0
+
+def inverse_error_Ch0(x):
+    return abs(-(1 / m0)) * error_b0 + abs((b0 - x) / m0 ** 2) * error_m0
+
+def error_Ch0(bins, error_bins):
+    '''Error propagation for the calibration function of channel 0.
+    The error is calculated using the errorpropagation website,
+    absolute error.'''
+    return abs(m0) * error_bins + abs(bins) * error_m0 + error_b0
+
+# Calibration function for channel 1
+m1 = calibration_data['m'].to_numpy()[1]
+b1 = calibration_data['c'].to_numpy()[1]
+error_m1 = calibration_data['error_m'].to_numpy()[1]
+error_b1 = calibration_data['error_c'].to_numpy()[1]
+
+def calibration_Ch1(x):
+    return m1 * x + b1
+
+def inverse_calibration_Ch1(x):
+    return (x - b1) / m1
+
+def error_Ch1(bins, error_bins):
+    return abs(m1) * error_bins + abs(bins) * error_m1 + error_b1
+
+# Calibration function for channel 2
+m2 = calibration_data['m'].to_numpy()[2]
+b2 = calibration_data['c'].to_numpy()[2]
+error_m2 = calibration_data['error_m'].to_numpy()[2]
+error_b2 = calibration_data['error_c'].to_numpy()[2]
+
+def calibration_Ch2(x):
+    return m2 * x + b2
+
+def inverse_calibration_Ch2(x):
[…]

### 2024-01-19 text/paper/paper.tex
tenso

--- 2024-01-08 text/paper/paper.tex
@@
     \begin{split}
-        E[\text{MeV}] &= m_0 \times E[\text{Chn}] + b_0 \\
+        E[\text{keV}] &= m_0 \times E[\text{Chn}] + b_0 \\
         m_0 &= 9.20 \pm 0.06 \\
@@
     \begin{split}
-        E[\text{MeV}] &= m_1 \times E[\text{Chn}] + b_1 \\
+        E[\text{keV}] &= m_1 \times E[\text{Chn}] + b_1 \\
         m_1 &= 8.97 \pm 0.03 \\
@@
     \begin{split}
-        E[\text{MeV}] &= m_2 \times E[\text{Chn}] + b_2 \\
+        E[\text{keV}] &= m_2 \times E[\text{Chn}] + b_2 \\
         m_2 &= 9.07 \pm 0.12 \\
@@
 
-\section{ Lithium Fluoride Sample}
+\section{Alpha Peaks}
+For the analysis, we began by calculating the theoretical energy values of the two alphas.
 
-$\ce{LiF}$
+\begin{center}
+    \label{alfa_peaks}
+    \centering
+    \includegraphics[scale = 0.35]{images/AlphaPeaks.png}
+    \captionof{figure}{Theoretical energy values of the two alfas}
+\end{center}
+
+The energies obtained were 7629.80 kEv and 11031.44 kEv. (FALTA EXPLICAR PERCENTAGENS)
+
+\section{Lithium Fluoride Sample}
+
+The obtained spectrum for the lithium fluoride sample overlapped with the theoretical alpha values is the following:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{../images/SampleLiF.png}
+    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
+\end{center}
+
+There are three noticeable structures in this spectrum. 
+The first one is a peak around channel 0, followed a less intense one around 100 Chn and a small plateau between 150 and 250 Chn. 
+Lastly there is an evident energy profile that starts at around 300 Chn and goes up to 900 Chn with increasing intensity.
+
+\paragraph*{}
+
+After identifying these structures we used the previous calibration [\ref{sec:calibration}] to convert the channel values into energy values so that we could more effectively analyse the spectrum and its peaks.
 
@@
 
-$\ce{LiAlO_2}$
+We obtained the following spectrum for the lithium aluminate sample:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{../images/SampleLiAlO2.png}
+    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
+\end{center}
 
@@
 \section{Implanted Sample}
+
+The obtained spectrum for the implanted sample was the following:
+
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.6]{../images/ImplantedSample.png}
+    \captionof{figure}{Spectrum of the implanted sample}
+\end{center}
 

### 2024-01-19 text/paper/paper.tex

--- 2024-01-19 text/paper/paper.tex
@@
 
-\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{images/IST_Logo.png}
+\title{\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}
 % \\ \vspace{1mm} LFEUI \vspace{1mm} 
@@
     \centering
-    \includegraphics[scale = 0.15]{images/scheme.jpeg}
+    \includegraphics[scale = 0.15]{../images/scheme.jpeg}
     \captionof{figure}{Experimental Set-up}
@@
     \centering
-    \includegraphics[scale = 0.17]{images/Li_crosssection_energy.jpeg}
+    \includegraphics[scale = 0.17]{../images/Li_crosssection_energy.jpeg}
     \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy}
@@
     \centering
-    \includegraphics[scale = 0.35]{images/Calib.png}
+    \includegraphics[scale = 0.35]{../../images/Calib.png}
     \captionof{figure}{Spectrum of the Calibration source}
@@
     \centering
-    \includegraphics[scale = 0.6]{images/TT_21_Chn0.png}
+    \includegraphics[scale = 0.6]{../../images/TT_21_Chn0.png}
     \captionof{figure}{Channel 0 spectrum fit with [\ref{eq:califit}]}
@@
     \centering
-    \includegraphics[scale = 0.35]{images/Chn0_calib_with_Backscattering.png}
+    \includegraphics[scale = 0.35]{../../images/Chn0_calib_with_Backscattering.png}
     \captionof{figure}{Calibration "detector 0"}
@@
     \centering
-    \includegraphics[scale = 0.35]{images/AlphaPeaks.png}
+    \includegraphics[scale = 0.35]{../../images/AlphaPeaks.png}
     \captionof{figure}{Theoretical energy values of the two alfphas}
@@
 The energies obtained were 7629.80 kEv and 11031.44 kEv. (FALTA EXPLICAR PERCENTAGENS)
+For the identification of lithium in our samples, we need to compare the theoretical energy values with the experimental ones.
+Since we know the the emited alpha in our samples is more energetic than the theoretical values, it is expected that the experimental peaks are slightly higher than the theoretical ones.
+If the seamples contains a peak in the proximity of the alpha peaks it shows that the sample contains lithium.
 
@@
 
-For the LiF sample, we overlapped the expected $\alpha$ energy getting the following spectrum:
+For the LiF sample, we used the calibration [\ref*{sec:calibration}] to convert the channel values into energy values.
+we overlapped the expected $\alpha$ energy with the spectrum, and obtained the following:
+
+\begin{center}
+    \centering
+    \includegraphics[scale = 0.35]{../../images/OverlapLiF_NC.png}
+    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
+\end{center}
+
+The experimental $\alpha_0$ is extremely close to the theoretical value, therefore we can confirm the presence of lithium in this sample. 
+Moreover, it is according to the expected hyphotesis that the observed lithium energy is slightly higher than the theoretical value.
+Analysing this spectrum further:
 
@@
     \centering
-    \includegraphics[scale = 0.35]{images/OverlapLiF_NC.png}
-    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
+    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiF.png}
+    \captionof{figure}{Full Analysis of $\ce{LiF}$ Spectrum}
 \end{center}
 
-The experimental $\alpha_0$ is extremely close to the theoretical value, therefore we can confirm the presence of lithium in this sample. Analysing this spectrum further:
+There are three noticeable structures in this spectrum. 
+The first one is a peak around 0 keV, which is just electronic noise. The following structure is backscattering, identified by the barrier of 1146.5 kEv. 
+The backscattering is followed by event pile-up, which is the energy profile referenced in section \ref{sec:energy_profile_peak}.
+
+\section{ Lithium Aluminate Sample}
+
+We used the calibration again to obtain the Lithium Alimunate energy spectrum overlapped with the expected $\alpha$ energy:
 
@@
     \centering
-    \includegraphics[scale = 0.4]{images/FullAnalysisLiF.png}
-    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
+    \includegraphics[scale = 0.35]{../../images/OverlapLiAlO2.png}
+    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
 \end{center}
 
-There are three noticeable structures in this spectrum. 
-The first one is a peak around channel 0, followed by a less intense one around 100 Chn and a small plateau between 150 and 250 Chn. 
-Lastly, there is an evident energy profile that starts at around 300 Chn and goes up to 900 Chn with increasing intensity.
-
-\paragraph*{}
-
-After identifying these structures we used the previous calibration [\ref{sec:calibration}] to convert the channel values into energy values so that we could more effectively analyse the spectrum and its peaks.
-
-\section{ Lithium Aluminate Sample}
-
-We obtained the following spectrum for the lithium aluminate sample:
+The experimental $\alpha_0$ is very similar to the theoretical value, hence we can confirm the presence of lithium in this sample. 
+If we analyse this spectrum further, we can see the following:
 
@@
     \centering
-    \includegraphics[scale = 0.35]{images/OverlapLiAlO2.png}
-    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
+    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiAlO2.png}
+    \captionof{figure}{Full Analysis of $\ce{LiAlO_2}$ Spectrum}
 \end{center}
+
+It is evident that the structure of the spectrum follows a similar pattern to the one in the previous sample.
+First, there is electrical noise, followed by backscattering and event pile-up. 
+The backscattering is identified by the barrier of 1150.0 kEv.
 
@@
 
-The obtained spectrum for the implanted sample was the following:
+Finally, we used the calibration to obtain the implanted sample energy spectrum overlapped with the expected $\alpha$ energy:
+\begin{center}
+    \label{TT_21}
+    \centering
+    \includegraphics[scale = 0.35]{../../images/OverlapImplanted.png}
+    \captionof{figure}{Spectrum of the implanted sample}
+\end{center}
+
+There is a peak in the proximity of the $\alpha_0$ peak, which is the expected lithium peak.
+One more time we can confirm the presence of lithium in this sample.
+
+If we analyse the spectrum further:
 
@@
     \centering
-    \includegraphics[scale = 0.35]{images/OverlapImplanted.png}
-    \captionof{figure}{Spectrum of the implanted sample}
+    \includegraphics[scale = 0.4]{../../images/FullAnalysisImplanted.png}
+    \captionof{figure}{Full Analysis of the Implanted Sample Spectrum}
 \end{center}

### 2024-01-19 text/paper/paper.tex
Friday, 19 January 2024 at 17:45:23
MEEEERDA

@@
 The first one is a peak around 0 keV, which is just electronic noise. The following structure is backscattering, identified by the barrier of 1146.5 kEv. 
-The backscattering is followed by event pile-up, which is the energy profile referenced in section \ref{sec:energy_profile_peak}.
+The backscattering is followed by event pile-up, which happens whenever two particles are detected at the same time and the detector only detects one with the energy doubled.
+The last structure is the lithium peak. The depth of the sample creates the visible energy profile and this procedure is explained in section \ref{sec:energy_profile_peak}.
 

### 2024-01-19 text/paper/paper.tex
CURVO
VAI

@@
 The backscattering is identified by the barrier of 1150.0 kEv.
+Finally, there is the lithium peak, which has an energy profile as well due to the depth of the sample.
 
@@
 \end{center}
+
+Just like before, there is electrical noise, followed by backscattering and event pile-up.
+However, this sample is implanted only on the surface, therefore there is no energy profile as expected, since the lithium isn't being detected at different depths.
 

### 2024-01-19 contas.py
contas.py

@@ -1,0 +1,23 @@
+import numpy as np
+largura = 9.6e-3
+distancia = 10e-2
+Omega = 2*np.pi*(1-np.cos(np.arctan(largura/(2*distancia))))
+print(f"Solid Angle: {round(Omega,4)} Sr")
+
+# Hight of the LiAlO2 and LiF 
+H = np.array([54, 144])
+QH=2e-6 #Coulombs
+# Area of the Implanted Sample  (in 50 channels) in "detector 0"
+A = 116
+QA = 20e-6 #Coulombs
+e = 1.602e-19 # Coulombs
+
+sigma = 1.39e-27 # cm^2/sr
+NTau = H/(Omega*sigma*QH/e)
+print(f"NTau LiAlO2: {'{:.2e}'.format(NTau[0])} atoms per cm^2")
+print(f"NTau LiF: {'{:.2e}'.format(NTau[1])} atoms per cm^2")
+NX = A/(Omega*sigma*QA/e)
+print(f"NX: {'{:.2e}'.format(NX)} atoms per cm^2")
+print("For comparison:")
+print(f"NX LiAlO2: {'{:.2e}'.format(NTau[0]*50)} atoms per cm^2")
+print(f"NX LiF: {'{:.2e}'.format(NTau[1]*50)} atoms per cm^2")

### 2024-01-19 text/paper/paper.tex
À
paper.tex

@@
 \paragraph{Lithium} Firstly, for the reaction \ref{eq:reactionLi}:
-
 \begin{equation}
@@
 \paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:
-
 \begin{equation}
@@
 We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
-So that we can do a calibration from the number of counts per channel to energy, we put an additional source on the holder just above the lithium fluoride one.
-Using a ruler, we measured the distance between the samples so that afterwards we could control the lift in order to change the sample on the beam's target.
+In order to do a calibration from the number of counts per channel to energy, we added an additional source on the holder just above the lithium fluoride one.
+Using a pachymeter, we measured the distance between the samples so that afterwards we could control the hight in order to change the sample on the beam's target.
 
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/Calib.png}
+    \includegraphics[scale = 0.3]{../../images/Calib.png}
     \captionof{figure}{Spectrum of the Calibration source}
@@
 
-The equation used for the triple gaussian fit is:
+%The equation used for the triple gaussian fit is:
+%\begin{equation}
+%    \begin{split}
+%    f(x) = a_0\exp{\left(-\frac{(x-\mu_0)^2}{\sigma_0}\right)} &+ a_1\exp{\left(-\frac{(x-\mu_1)^2}{\sigma_1}\right)} + \\
+%    &+ a_2\exp{\left(-\frac{(x-\mu_2)^2}{\sigma_2}\right)} + c
+%    \end{split}
+%    \label{eq:califit}
+%\end{equation}
 
-\begin{equation}
-    \begin{split}
-    f(x) = a_0\exp{\left(-\frac{(x-\mu_0)^2}{\sigma_0}\right)} &+ a_1\exp{\left(-\frac{(x-\mu_1)^2}{\sigma_1}\right)} + \\
-    &+ a_2\exp{\left(-\frac{(x-\mu_2)^2}{\sigma_2}\right)} + c
-    \end{split}
-    \label{eq:califit}
-\end{equation}
-
-We decided to do the analysis only with "detector 0". This anlaysis could be further corroborated using the "detector 1" data.
-Using the triple gaussian, we obtained the following fits for "detector 0":
+We decided to do the analysis only with "detector 0". This analysis could be further corroborated using the "detector 1" data.
+Using the triple gaussian, we obtained the following fit for "detector 0":
 
@@
     \centering
-    \includegraphics[scale = 0.6]{../../images/TT_21_Chn0.png}
-    \captionof{figure}{Channel 0 spectrum fit with [\ref{eq:califit}]}
+    \includegraphics[scale = 0.5]{images/TT_21_Chn0.png}
+    \captionof{figure}{Detector 0 spectrum fit with triple gaussian}
 \end{center}
@@
 
-The fit parameters for means $\mu_0$, $\mu_1$ and $\mu_2$ in equation [\ref{eq:califit}] represent the energy value of each peak in channels.
-The computed values for these parameters, as well as the theoretical values of the energy in MeV for the peaks, are organised in the following tables:
+The fit parameters for the means, $\mu_0$, $\mu_1$ and $\mu_2$ obtained from the triple gaussian fit represent the energy value of each peak in channels.
+The computed values for these parameters, as well as the theoretical values of the energy, in MeV, for the peaks, are organised in the following tables:
 
@@
 \hline
-Peak & Channel 0 [Chn] & Theoretical Energy [keV] \\ \hline
+Peak & Detector 0 [Chn] & Theoretical Energy [keV] \\ \hline
 1\textsuperscript{st} & $ \mu_0 = 557.00 \pm 0.11 $ & 5156.59 \\ \hline
@@
 \end{tabular}
-\caption{Calibration Values for Channel 0}
+\caption{Calibration Values for detector 0}
 \label{tab:calibration0}
@@
 Using the values in these three tables we now need to compute a calibration equation for each of the channels.
-We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:
+We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:\vspace{-2mm}
 
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/Chn0_calib_with_Backscattering.png}
+    \includegraphics[scale = 0.3]{../../images/Chn0_calib_with_Backscattering.png}
     \captionof{figure}{Calibration "detector 0"}
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/AlphaPeaks.png}
+    \includegraphics[scale = 0.3]{../../images/AlphaPeaks.png}
     \captionof{figure}{Theoretical energy values of the two alfphas}
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/OverlapLiF_NC.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapLiF_NC.png}
     \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
@@
     \centering
-    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiF.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiF.png}
     \captionof{figure}{Full Analysis of $\ce{LiF}$ Spectrum}
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/OverlapLiAlO2.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapLiAlO2.png}
     \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
@@
     \centering
-    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiAlO2.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiAlO2.png}
     \captionof{figure}{Full Analysis of $\ce{LiAlO_2}$ Spectrum}
@@
 
+(NÃO DEVIA ESTAR NA TEORIA?)
+
 \section{Implanted Sample}
@@
     \centering
-    \includegraphics[scale = 0.35]{../../images/OverlapImplanted.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapImplanted.png}
     \captionof{figure}{Spectrum of the implanted sample}
@@
     \centering
-    \includegraphics[scale = 0.4]{../../images/FullAnalysisImplanted.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisImplanted.png}
     \captionof{figure}{Full Analysis of the Implanted Sample Spectrum}
@@
 \section{Quantitative Analysis}
-According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:
+According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:\\
+\begin{minipage}[t]{0.5\linewidth-0.5em} % some separation
+    \vspace{0pt} % anchor for [t]
+    \vspace{\dimexpr\ht\strutbox-\topskip}% remove excess glue
     \begin{equation}
@@
     \end{equation}
+\end{minipage}\hfill
+\begin{minipage}[t]{0.5\linewidth-0.5em} % some separation
+    \vspace{0pt} % anchor for [t]
     \begin{equation}
@@
     \end{equation}
+\end{minipage}
 
-Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \cdot$\#channels.\\
-$H_i$ is the number of counts in a FAT MUDAR spectrum peak (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.
+\vspace{4mm}
+
+Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$\#channels.
+
+$H_i$ is the hight, in number of counts, in a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.
 
 With this we get:\\
-APROXIMACOES:\\
-implanted: $1\times10^{17}$ atoms/cm$^2$\\
-$\ce{LiAlO2}$: $3\times10^{18}$ atoms/cm$^2$\\
-etc...\\
+Implanted Sample: $N\Delta x = 9.25\times 10^{16}$ atoms per cm$^2$\\
+$\ce{LiAlO2}$: $N\tau_i=4.31\times10^{17}$ atoms per cm$^2$\\
+$\ce{LiF}$: $N\tau_i=1.15\times10^{18}$ atoms per cm$^2$
+
+For comparison:
+$\ce{LiAlO2}$: $N\Delta x=2.15\times10^{19}$ atoms per cm$^2$\\
+$\ce{LiF}$: $N\Delta x=5.71\times10^{19}$ atoms per cm$^2$\\
+From this, it is noticeable that the implanted sample has around 200x less Lithium than the $\ce{LiAlO2}$ and 600x less Lithium than the $\ce{LiF}$, as expected.
 
 \section{Conclusion}
+To conclude, we ended up with a successful experiment. We were able to detect the presence of lithium in all the 
+analysed samples trough nuclear reactions. Furthermore, we managed to go beyond our proposal and extimate the amount of 
+lithium in each sample.
+The results obtained were in accordance with the values obtained from the Danish Institute that provided the samples. MEIO INVENTADO
 
@@
 
-\section{}
-
 \printbibliography

### 2024-01-19 calib.py

@@ -66,9 +66,9 @@ popt0, pcov0 = curve_fit(gaussian, x, y, p0=[90, 556, 5, 70, 593, 5, 70, 627, 5,
 ercov0 = np.sqrt(np.diag(pcov0))
 
 plt.figure()
-plt.plot(TT21_Chn0, label='Channel 0')
-plt.plot(x, gaussian(x, *popt0), label='Chnannel 0 fit')
-plt.title('Fit channel 0')
+plt.plot(TT21_Chn0, label='Detector 0')
+plt.plot(x, gaussian(x, *popt0), label='Detector 0 fit')
+plt.title('Fit detector 0')
 plt.legend()
 plt.xlabel('Channel')
 plt.yscale('log')
@@ -84,7 +84,7 @@ info = [
     [f'{popt0[0]:.2f} ± {ercov0[0]:.2f}', f'{popt0[1]:.2f} ± {ercov0[1]:.2f}', f'{popt0[2]:.2f} ± {ercov0[2]:.2f}', f'{popt0[3]:.2f} ± {ercov0[3]:.2f}', f'{popt0[4]:.2f} ± {ercov0[4]:.2f}', f'{popt0[5]:.2f} ± {ercov0[5]:.2f}', f'{popt0[6]:.2f} ± {ercov0[6]:.2f}', f'{popt0[7]:.2f} ± {ercov0[7]:.2f}', f'{popt0[8]:.2f} ± {ercov0[8]:.2f}', f'{popt0[9]:.2f} ± {ercov0[9]:.2f}']
 ]
 
-print(tabulate([['Channel 0']]))
+print(tabulate([['Detector 0']]))
 print(tabulate(info, headers='firstrow', tablefmt='fancy_grid'))
 
 # -----------------------------------------------------------------------------

### 2024-01-20 text/paper/paper.tex
Saturday, 20 January 2024 at 01:09:58

@@
   \par 
+  The following paper describes a scientific research on the ability to detect the presence of lithium ($\ce{^7Li}$) through the byproducts of nuclear reactions.
+  The research was conducted in three different samples: lithium fluoride ($\ce{LiF}$), lithium aluminate ($\ce{LiAlO2}$) and an implanted sample. 
+  Energy spectra were obtained for each sample with a tandem accelerator and $1.3$ MeV protons.
+  Considering the theoretical values for the nuclear reaction of $\ce{^7Li}$, we were able to identify the presence of lithium in all samples.
+  Also, some other phenomena were observed, such as back-scattering, pile-up and energy profiling.
+  Finally, using the information on each spectrum, we computed the quantity of lithium in each sample.
 \end{abstract}
@@
 
-\paragraph{} Some of the samples we are going to work with also include boron and according to the beam energy we use it is possible that the protons lead to a nuclear reaction with the boron atoms in the samples.
+\paragraph{}
+
+Some of the samples we are going to work with also include boron and according to the beam energy we use it is possible that the protons lead to a nuclear reaction with the boron atoms in the samples.
 Therefore, the $\ce{^{11}B}$ reaction with a proton is as follows:
@@
 
-\subsection{Elastic Back-scattering}
+\subsection{Back-scattering}
     \label{sec:elastic_back}
 
-Another possible occurrence is elastic back-scattering, where the incoming proton from the accelerator is elastically reflected back by the sample and subsequently detected.
-During this reflection, the energy remains unchanged, resulting in an intensity peak positioned close to the beam energy selected for our experiment.
+Another possible occurrence is back-scattering, where the incoming proton from the accelerator is elastically reflected back by the sample and subsequently detected.
+During this reflection, the energy of the proton and the target is conserved and distributed according to the kinematics of the collision.
+This will result in a range of energy values for the back-scattered proton.
 
-o backscattering é todo elástico e um caso particular deste é o de rutherford.
-O caso particular de rutherford considera os atomos como esferas perfeitas para calcular a secção eficaz deste fenomeno.
-Em geral, as secções eficazes do backscattering obtém-se experimentalmente e não usamos as de rutherford.
-O que interessa é que o backscattering elastico como fenomeno e não as suas diferentes secções eficazes.
-Pileup - o detetor recebe duas partículas de backscattering em simultaneo e assum que é so uma e por isso soma uma contagem com o dobro da energia - A particula emitida no backscattering pode ter diferentes valores de energia de acordo com a colisao elastica entre os dois atomos (a mecanica diz nos como calular a energia com que cada átomo, o incidente e o da amosta à superfície, fica depois da colisão) e por isso a soma da energia de duas também pode estar num dado intervalo.
-O pico inicial é do ruido eletronico e por isso desconsideramos.
-Usar detector em vez de channel.
-A secção eficaz da reação nuclear com o litio varia com a energia do protão incidente.
-O protão incidente perde energia à medida que percorre a amostra devido ao stopping power da amostra em si e por isso a secção eficaz de reacções que ocorrem mais no interior da amostra é diferente das que ocorrem à sua superfície.
+\subsection{Pile-up}
+    \label{sec:pileup}
+
+Pile-up is a phenomenon that is a consequence of the finite time resolution of the detector.
+It occurs when two particles are detected at the same time and the detector only detects one with the energy doubled.
+This is a problem because it can lead to the misinterpretation of the data.
+Therefore, we expect that, for very intense energy ranges in the spectrum, the pile-up might be significant. 
+If we have particles emitted with an $[a,b]$ keV energy range, the pile-up will occur in the $[2a,2b]$ keV.
+
+% o backscattering é todo elástico e um caso particular deste é o de rutherford. 
+% O caso particular de rutherford considera os atomos como esferas perfeitas para calcular a secção eficaz deste fenomeno. 
+% Em geral, as secções eficazes do backscattering obtém-se experimentalmente e não usamos as de rutherford. 
+% O que interessa é que o backscattering elastico como fenomeno e não as suas diferentes secções eficazes. 
+% Pileup - o detetor recebe duas partículas de backscattering em simultaneo e assum que é so uma e por isso soma uma contagem com o dobro da energia.
+% A particula emitida no backscattering pode ter diferentes valores de energia de acordo com a colisao elastica entre os dois atomos (a mecanica diz nos como calular a energia com que cada átomo, o incidente e o da amosta à superfície, fica depois da colisão) e por isso a soma da energia de duas também pode estar num dado intervalo.
+% O pico inicial é do ruido eletronico e por isso desconsideramos. Usar detector em vez de channel. 
+% A secção eficaz da reação nuclear com o litio varia com a energia do protão incidente. 
+% O protão incidente perde energia à medida que percorre a amostra devido ao stopping power da amostra em si e por isso a secção eficaz de reacções que ocorrem mais no interior da amostra é diferente das que ocorrem à sua superfície.
 
@@
 Finally, we used a computer program to obtain the spectra of the different sources.
-The multichannel analyzer (MCA) utilized to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
+The multichannel analyzer (MCA) used to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
 
@@
 
-We separated the spectra of channeldetectors 0, 1 and 2 (2 corresponding to the bottom one) and used a triple gaussian sum function to fit on the three3 energy peaks.
-According to theory, the first peak corresponds to the decay of plutonium into uranium, the second of americium into neptunium and the third peak of curium into plutonium.
+We separated the spectra of channel/detectors 0, 1 and 2 (2 corresponding to the bottom one) and used a triple gaussian sum function to fit on the three energy peaks.
+According to theory, the first peak corresponds to the decay of plutonium into uranium, the second of americium into neptunium and the third of curium into plutonium.
 The decay equations are:
@@
 
-We decided to do the analysis only with "detector 0". This analysis could be further corroborated using the "detector 1" data.
-Using the triple gaussian, we obtained the following fit for "detector 0":
+We decided to do the analysis only with detector/channel 0. This analysis could be further corroborated using the "detector 1" data.
+Using the triple gaussian, we obtained the following fit for channel 0:
 
@@
     \centering
-    \includegraphics[scale = 0.5]{images/TT_21_Chn0.png}
+    \includegraphics[scale = 0.5]{../../images/TT_21_Chn0.png}
     \captionof{figure}{Detector 0 spectrum fit with triple gaussian}
@@
 
-Using the values in these three tables we now need to compute a calibration equation for each of the channels.
+Using the values in the table we now have to compute a calibration equation for channel/detector 0.
 We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:\vspace{-2mm}
@@
 
-The calibration equation for "detector 0" is:
+The calibration equation for channel 0 is:
 
@@
 %
-\section{Alfpha Peaks}
-For the analysis, we began by calculating the theoretical energy values of the two alfphas.
+
+\section{Alpha Peaks}
+    \label{sec:alphapeaks}
+
+For the analysis, we began by calculating the theoretical energy values of the two alphas.
+
+DE ONDE VEM ISTO CARAGO? TEMOS DE TER REFERÊNCIAS
 
@@
     \includegraphics[scale = 0.3]{../../images/AlphaPeaks.png}
-    \captionof{figure}{Theoretical energy values of the two alfphas}
+    \captionof{figure}{Theoretical energy values of the two alphas}
 \end{center}
@@
 For the identification of lithium in our samples, we need to compare the theoretical energy values with the experimental ones.
-Since we know the the emited alpha in our samples is more energetic than the theoretical values, it is expected that the experimental peaks are slightly higher than the theoretical ones.
-If the seamples contains a peak in the proximity of the alpha peaks it shows that the sample contains lithium.
+Since we know the emited alpha in our samples is more energetic than the theoretical values, it is expected that the experimental peaks are slightly higher than the theoretical ones.
+If the samples contains a peak in the proximity of the alpha peaks it shows that the sample contains lithium.
 
 \section{Lithium Fluoride Sample}
+    \label{sec:lif}
 
-For the LiF sample, we used the calibration [\ref*{sec:calibration}] to convert the channel values into energy values.
+For the LiF sample, we used the calibration [\ref{sec:calibration}] to convert the channel values into energy values.
 we overlapped the expected $\alpha$ energy with the spectrum, and obtained the following:
@@
 The experimental $\alpha_0$ is extremely close to the theoretical value, therefore we can confirm the presence of lithium in this sample. 
-Moreover, it is according to the expected hyphotesis that the observed lithium energy is slightly higher than the theoretical value.
+Moreover, it is according to the expected hypothesis that the observed lithium energy is slightly higher than the theoretical value.
 Analysing this spectrum further:
@@
 There are three noticeable structures in this spectrum. 
-The first one is a peak around 0 keV, which is just electronic noise. The following structure is backscattering, identified by the barrier of 1146.5 kEv. 
-The backscattering is followed by event pile-up, which happens whenever two particles are detected at the same time and the detector only detects one with the energy doubled.
-The last structure is the lithium peak. The depth of the sample creates the visible energy profile and this procedure is explained in section \ref{sec:energy_profile_peak}.
+The first one is a peak around 0 keV, which is just electronic noise. The following structure is back scattering [\ref{sec:elastic_back}], identified by the barrier of 1146.5 kEv. 
+The back scattering is followed by event pile-up [\ref{sec:pileup}].
+The last structure is the lithium peak. The depth of the sample creates the visible energy profile and this procedure is explained in section [\ref{sec:energy_profile_peak}].
 
-\section{ Lithium Aluminate Sample}
+\section{Lithium Aluminate Sample}
+    \label{sec:lialo2}
 
-We used the calibration again to obtain the Lithium Alimunate energy spectrum overlapped with the expected $\alpha$ energy:
+We used the calibration again to obtain the Lithium Aluminate energy spectrum overlapped with the expected $\alpha$ energy:
 
@@
 The experimental $\alpha_0$ is very similar to the theoretical value, hence we can confirm the presence of lithium in this sample. 
-If we analyse this spectrum further, we can see the following:
+If we analyze this spectrum further, we can see the following:
 
@@
 It is evident that the structure of the spectrum follows a similar pattern to the one in the previous sample.
-First, there is electrical noise, followed by backscattering and event pile-up. 
-The backscattering is identified by the barrier of 1150.0 kEv.
-Finally, there is the lithium peak, which has an energy profile as well due to the depth of the sample.
-
-\subsection{Energy Profile}
-
-(NÃO DEVIA ESTAR NA TEORIA?)
+First, there is electrical noise, followed by back scattering [\ref{sec:elastic_back}] and event pile-up [\ref{sec:pileup}]. 
+The back scattering is identified by the barrier of 1150.0 keV.
+Finally, there is the lithium peak, which has an energy profile [\ref{sec:energy_profile_peak}] as well due to the depth of the sample.
 
 \section{Implanted Sample}
+    \label{sec:implanted}
 
 Finally, we used the calibration to obtain the implanted sample energy spectrum overlapped with the expected $\alpha$ energy:
+
 \begin{center}
@@
 
-If we analyse the spectrum further:
+If we analyze the spectrum further:
 
@@
 
-Just like before, there is electrical noise, followed by backscattering and event pile-up.
+Just like before, there is electrical noise, followed by back scattering [\ref{sec:elastic_back}] and event pile-up [\ref{sec:pileup}].
 However, this sample is implanted only on the surface, therefore there is no energy profile as expected, since the lithium isn't being detected at different depths.
@@
 \section{Quantitative Analysis}
-According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:\\
+    \label{sec:quantitative}
+
+According to \cite{chu1978backscattering} it is possible to calculate the concentration of the element in the samples using with the following equations:
+
 \begin{minipage}[t]{0.5\linewidth-0.5em} % some separation
@@
 
-Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$\#channels.
+Where $N\tau_i$ are $N\Delta x$ different values for the superficial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$ \# channels.
 
-$H_i$ is the hight, in number of counts, in a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.
+$H_i$ is the height, in counts, on a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector and $Q$ is the max charge acquisition divided by $e$.
 
 With this we get:\\
+
 Implanted Sample: $N\Delta x = 9.25\times 10^{16}$ atoms per cm$^2$\\
@@
 $\ce{LiF}$: $N\Delta x=5.71\times10^{19}$ atoms per cm$^2$\\
+
 From this, it is noticeable that the implanted sample has around 200x less Lithium than the $\ce{LiAlO2}$ and 600x less Lithium than the $\ce{LiF}$, as expected.
@@
 \section{Conclusion}
+    \label{sec:conclusion}
+
 To conclude, we ended up with a successful experiment. We were able to detect the presence of lithium in all the 
-analysed samples trough nuclear reactions. Furthermore, we managed to go beyond our proposal and extimate the amount of 
+analysed samples trough nuclear reactions. Furthermore, we managed to go beyond our proposal and estimate the amount of 
 lithium in each sample.
 The results obtained were in accordance with the values obtained from the Danish Institute that provided the samples. MEIO INVENTADO
-
-\section{Perguntas}
-
-\begin{enumerate}
-    \item Para a calibração, colocamos os valores e as incertezas de todos os 10 parâmetros para cada channel ou apenas colocamos as médias ?
-\end{enumerate}
 

### 2024-01-20 text/paper/paper.tex
paper.tex

@@
 % pdflatex -output-directory=/Users/salvatorpes/Desktop/LFEUI/text/proposta /Users/salvatorpes/Desktop/LFEUI/text/proposta/prop.tex
+% pdflatex -output-directory=/Users/estevao/Desktop/Universidade/LFEUI/text/paper /Users/estevao/Desktop/Universidade/LFEUI/text/paper/paper.tex
 
@@
   The following paper describes a scientific research on the ability to detect the presence of lithium ($\ce{^7Li}$) through the byproducts of nuclear reactions.
-  The research was conducted in three different samples: lithium fluoride ($\ce{LiF}$), lithium aluminate ($\ce{LiAlO2}$) and an implanted sample. 
+  The research was conducted in 3 different samples: lithium fluoride ($\ce{LiF}$), lithium aluminate ($\ce{LiAlO2}$) and an implanted sample. 
   Energy spectra were obtained for each sample with a tandem accelerator and $1.3$ MeV protons.
   Considering the theoretical values for the nuclear reaction of $\ce{^7Li}$, we were able to identify the presence of lithium in all samples.
-  Also, some other phenomena were observed, such as back-scattering, pile-up and energy profiling.
-  Finally, using the information on each spectrum, we computed the quantity of lithium in each sample.
+  Some other phenomena were observed, such as back-scattering, pile-up and energy profiling.
+  Finally, using the information on each spectrum, we estimated the quantity of lithium in each sample.
 \end{abstract}
@@
 
-Our main goal is to develop a scientific research project alongside three senior investigators from CTN (Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
+Our main goal is to develop a scientific research project alongside 3 senior investigators from CTN (Centro Tecnológico e Nuclear), Rodrigo Mateus, Rui Silva and Norberto Catarino.
 Our project will be focused on the detection of Li in technological materials through nuclear reactions.
@@
 
-The development of this scientific procedure is based on a physics phenomenon: nuclear reactions.
-We have an accelerator in order to allow protons with specific energy to collide with the sample whose composition we are interested in studying.
-The protons will undergo a nuclear reaction with the ones in the samples and byproducts of these reaction will be released.
-These byproducts are charged particles and we can detect their energy spectrum.
-Through the analysis of it, we want to infer what were the elements with whom they reacted on the sample and what is their composition percentage.
+The development of this scientific procedure is based on nuclear reactions.
+We have an accelerator that allows protons to reach high energyies in order to collide with the sample whose composition we are interested in studying.
+The protons will undergo a nuclear reaction with the nuclei in the samples and byproducts of these reaction will be released.
+These are charged particles that we can detect and quantify through their energy spectrum.
 
@@
 
-The main nuclear reaction in which we are interested is the one between $\ce{^7Li}$ and an incoming proton on the beam:
+The main nuclear reaction we are interested in is between $\ce{^7Li}$ and an incoming proton from the beam:
 
@@
 
-Before the emission of the reaction products, the reaction leads to the formation of a $\ce{^8Be}$ compound nucleus (intermediate state), which then rapidly decays into two alpha particles.
-These are the particles we detect.
-An alpha particle is a helium nucleus, composed of two protons and two neutrons.
+Before the emission of the reaction products, the reaction leads to an intermediate state, a $\ce{^8Be}$ compound nucleus, which then rapidly decays into two alpha particles that we can detect.% An alpha particle is a helium nucleus, composed of two protons and two neutrons.
+%toda a gente sabe o que é uma particula alpha valhamedeus
 
@@
 It is less likely for the particles that are emitted inside the sample to reach the detector than for those who were emitted on its surface since on each collision there is a chance that it's direction might change.
-This is why the intensity on the energy profile as a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction.
+This is why the intensity on the energy profile has a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction.
 \paragraph{}
@@
 \begin{split}
-  E_{\alpha_{0_1}} &= 7.6298 \text{ MeV} \\
-  E_{\alpha_{0_2}} &= 11.0314 \text{ MeV} \\
+  E_{\alpha_{0_1}} &= 7.630 \text{ MeV} \\
+  E_{\alpha_{0_2}} &= 11.031 \text{ MeV} \\
   \theta_{\alpha_{0_1}} &= 165 ^{\circ} \\
-  \theta_{\alpha_{0_2}} &= 12 ^{\circ} \\
+  \theta_{\alpha_{0_2}} &= 12.4 ^{\circ} \\
   Q (\ce{^7Li}) &= 17.346 \text{ MeV} \\
@@
 \begin{split}
-  E_{\alpha_0} &= 3.7372 \text{ MeV} \\
-  E_{\ce{^8Be}} &= 3.1086 \text{ MeV} \\
+  E_{\alpha_0} &= 3.749 \text{ MeV} \\
+  E_{\ce{^8Be}} &= 3.1116 \text{ MeV} \\
   \theta_{\alpha_0} &= 165 ^{\circ} \\
-  \theta_{\ce{^8Be}} &= 11 ^{\circ}\\
+  \theta_{\ce{^8Be}} &= 11.6 ^{\circ}\\
   Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
@@
 
-We separated the spectra of channel/detectors 0, 1 and 2 (2 corresponding to the bottom one) and used a triple gaussian sum function to fit on the three energy peaks.
+We separated the spectra of 
+detectors 0, 1 and 2 (2 corresponding to the bottom one) and used a triple gaussian sum function to fit on the three energy peaks.
 According to theory, the first peak corresponds to the decay of plutonium into uranium, the second of americium into neptunium and the third of curium into plutonium.
@@
 
-We decided to do the analysis only with detector/channel 0. This analysis could be further corroborated using the "detector 1" data.
-Using the triple gaussian, we obtained the following fit for channel 0:
+We decided to do the analysis only with detector 0. This analysis could be further corroborated using detector 1 data.
+Using the triple gaussian, we obtained the following fit for this detector:
 
@@
 
-Using the values in the table we now have to compute a calibration equation for channel/detector 0.
+Using the values in the table we now have to compute a calibration equation for detector 0.
 We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:\vspace{-2mm}
@@
 
-The calibration equation for channel 0 is:
+The calibration equation for the aforementioned detector is:
 
@@
 
-Where $N\tau_i$ are $N\Delta x$ different values for the superficial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$ \# channels.
+Where $N\tau_i$ are $N\Delta x$ different values for the superficial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$ \#channels.
 
-$H_i$ is the height, in counts, on a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector and $Q$ is the max charge acquisition divided by $e$.
+$H_i$ is the height, in counts, on a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector and $Q$ is the max charge acquisition divided by $e$.\\
 
-With this we get:\\
+With this we get:
 

### 2024-01-20 text/notas/notas.tex
Saturday, 20 January 2024 at 16:31:20

@@
 
+\section{Back Scattering - Dúvidas}
+
+o backscattering é todo elástico e um caso particular deste é o de rutherford.
+O caso particular de rutherford considera os atomos como esferas perfeitas para calcular a secção eficaz deste fenomeno.
+Em geral, as secções eficazes do backscattering obtém-se experimentalmente e não usamos as de rutherford.
+O que interessa é que o backscattering elastico como fenomeno e não as suas diferentes secções eficazes.
+Pileup - o detetor recebe duas partículas de backscattering em simultaneo e assum que é so uma e por isso soma uma contagem com o dobro da energia.
+A particula emitida no backscattering pode ter diferentes valores de energia de acordo com a colisao elastica entre os dois atomos (a mecanica diz nos como calular a energia com que cada átomo, o incidente e o da amosta à superfície, fica depois da colisão) e por isso a soma da energia de duas também pode estar num dado intervalo.
+O pico inicial é do ruido eletronico e por isso desconsideramos. Usar detector em vez de channel.
+A secção eficaz da reação nuclear com o litio varia com a energia do protão incidente.
+O protão incidente perde energia à medida que percorre a amostra devido ao stopping power da amostra em si e por isso a secção eficaz de reacções que ocorrem mais no interior da amostra é diferente das que ocorrem à sua superfície.
+
 \printbibliography

### 2024-01-20 text/paper/paper.tex
Saturday, 20 January 2024 at 16:31:20

@@
 
-Before the emission of the reaction products, the reaction leads to an intermediate state, a $\ce{^8Be}$ compound nucleus, which then rapidly decays into two alpha particles that we can detect.% An alpha particle is a helium nucleus, composed of two protons and two neutrons.
-%toda a gente sabe o que é uma particula alpha valhamedeus
+Before the emission of the reaction products, the reaction leads to an intermediate state, a $\ce{^8Be}$ compound nucleus, which then rapidly decays into two alpha particles that we can detect.
+An alpha particle is a helium nucleus, composed of two protons and two neutrons. % toda a gente sabe o que é uma particula alpha valhamedeus
 
-\paragraph{}
+% \paragraph{}
 
-Some of the samples we are going to work with also include boron and according to the beam energy we use it is possible that the protons lead to a nuclear reaction with the boron atoms in the samples.
+% Some of the samples we are going to work with also include boron and according to the beam energy we use it is possible that the protons lead to a nuclear reaction with the boron atoms in the samples.
 Therefore, the $\ce{^{11}B}$ reaction with a proton is as follows:
 
-\begin{equation}
-  \ce{ ^{11}B + ^1H -> ^{12}C -> ^8Be + \alpha}
-\label{eq:reactionB}
-\end{equation}
+% \begin{equation}
+%   \ce{ ^{11}B + ^1H -> ^{12}C -> ^8Be + \alpha}
+% \label{eq:reactionB}
+% \end{equation}
 
-The reaction first leads to the formation of a $\ce{^{12}C}$ compound nucleus which decays into one alpha particle and a $\ce{^8Be}$ nucleus.
+% The reaction first leads to the formation of a $\ce{^{12}C}$ compound nucleus which decays into one alpha particle and a $\ce{^8Be}$ nucleus.
 The $\alpha$ particle can be detected with more or less energy according to the energy level on which the $\ce{^8Be}$ nucleus is emitted.
@@
 The computation process will be detailed in section \ref{sec:theoreticalvalues}.
+
 \paragraph{}
+
 Firstly, we expect to see intensity peaks on the spectrum on these specific energies.
@@
 The byproducts produced inside the sample initially have the energy we compute but do lose part of it as they collide with other particles before reaching the detector.
-Basically, we must consider the sample's thickness.
-The bigger the depth at which the reaction takes place, the smaller the energy with whom they are detected because more of it gets dissipated.
+Basically, we must consider the sample's thickness. The bigger the depth at which the reaction takes place, the smaller the energy with whom they are detected because more of it gets dissipated.
 Therefore, this will result in an energy profile - the energy peak will be gradually preceded by less intense peaks due to particles that lost energy on their way to the detector.
@@
 This is why the intensity on the energy profile has a steady increase before hitting a maximum value at the energy with whom the particle comes out of the reaction.
+
 \paragraph{}
+
 We expect this phenomena to be particularly significant in samples where the atom that takes part in the reaction can be found not only at the surface of the sample but also in its depth.
@@
 If we have particles emitted with an $[a,b]$ keV energy range, the pile-up will occur in the $[2a,2b]$ keV.
-
-% o backscattering é todo elástico e um caso particular deste é o de rutherford. 
-% O caso particular de rutherford considera os atomos como esferas perfeitas para calcular a secção eficaz deste fenomeno. 
-% Em geral, as secções eficazes do backscattering obtém-se experimentalmente e não usamos as de rutherford. 
-% O que interessa é que o backscattering elastico como fenomeno e não as suas diferentes secções eficazes. 
-% Pileup - o detetor recebe duas partículas de backscattering em simultaneo e assum que é so uma e por isso soma uma contagem com o dobro da energia.
-% A particula emitida no backscattering pode ter diferentes valores de energia de acordo com a colisao elastica entre os dois atomos (a mecanica diz nos como calular a energia com que cada átomo, o incidente e o da amosta à superfície, fica depois da colisão) e por isso a soma da energia de duas também pode estar num dado intervalo.
-% O pico inicial é do ruido eletronico e por isso desconsideramos. Usar detector em vez de channel. 
-% A secção eficaz da reação nuclear com o litio varia com a energia do protão incidente. 
-% O protão incidente perde energia à medida que percorre a amostra devido ao stopping power da amostra em si e por isso a secção eficaz de reacções que ocorrem mais no interior da amostra é diferente das que ocorrem à sua superfície.
 
@@
 
-In order to choose the energy of the beam we need to take into account the cross section of the reaction between the protons and the Li-7.
+In order to choose the energy of the beam we need to take into account the cross-section of the reaction between the protons and the Li-7.
 We collected data from the IBANDL database for this reaction in a range of [0.5,7] MeV and plotted it in the following graph:
@@
     \includegraphics[scale = 0.17]{../images/Li_crosssection_energy.jpeg}
-    \captionof{figure}{Cross section of the reaction between protons and Li-7 as a function of the beam energy}
+    \captionof{figure}{Cross-section of the reaction between protons and Li-7 as a function of the beam energy}
 \end{center}
@@
 We are only interested in the energy range of [0.5,5] MeV because the energy of the beam is limited by the accelerators range.
-As we can see in the above figure, the cross section of the reaction is maximum in the [2.5, 3] MeV range.
+As we can see in the above figure, the cross-section of the reaction is maximum in the [2.5, 3] MeV range.
 Therefore, since we want to detect the Li-7 in the samples, we should choose the energy of the beam to be in this range.
@@
 Increasing beam energies take exponentially more time to be set up.
-Also, we are interested in the possible boron identification in some of the used samples.
+Also, we are interested in the possible boron identification in some samples.
 Boron does not react with such high energies but a lower energy beam is effective in doing that.
@@
 
-\paragraph{Lithium} Firstly, for the reaction \ref{eq:reactionLi}:
+\paragraph{Lithium} For the reaction \ref{eq:reactionLi}:
 \begin{equation}
 \begin{split}
-  E_{\alpha_{0_1}} &= 7.630 \text{ MeV} \\
-  E_{\alpha_{0_2}} &= 11.031 \text{ MeV} \\
+  E_{\alpha_{0_1}} &= 7.62980 \text{ MeV} \\
+  E_{\alpha_{0_2}} &= 11.03144 \text{ MeV} \\
   \theta_{\alpha_{0_1}} &= 165 ^{\circ} \\
@@
 
-\paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:
-\begin{equation}
-\begin{split}
-  E_{\alpha_0} &= 3.749 \text{ MeV} \\
-  E_{\ce{^8Be}} &= 3.1116 \text{ MeV} \\
-  \theta_{\alpha_0} &= 165 ^{\circ} \\
-  \theta_{\ce{^8Be}} &= 11.6 ^{\circ}\\
-  Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
-\end{split}
-\label{eq:energiesBr}
-\end{equation}
+% \paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:
+% \begin{equation}
+% \begin{split}
+%   E_{\alpha_0} &= 3.749 \text{ MeV} \\
+%   E_{\ce{^8Be}} &= 3.1116 \text{ MeV} \\
+%   \theta_{\alpha_0} &= 165 ^{\circ} \\
+%   \theta_{\ce{^8Be}} &= 11.6 ^{\circ}\\
+%   Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
+% \end{split}
+% \label{eq:energiesBr}
+% \end{equation}
 
-We need to comment these energies.
+% We need to comment these energies.
 
@@
 We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
-In order to do a calibration from the number of counts per channel to energy, we added an additional source on the holder just above the lithium fluoride one.
-Using a pachymeter, we measured the distance between the samples so that afterwards we could control the hight in order to change the sample on the beam's target.
-
-We placed the sample holder on a support and afterwards both of them inside the chamber. A vacuum machine was connected to the accelerator to reduce the pressure inside both the chamber and the beam.
-
+In order to do a calibration we added a triple $\alpha$ source on the holder above the lithium fluoride. 
+Using a pachymeter, we measured the distance between the samples so that afterwards we could control the height of the holder in the chamber and change the beam's target.
+We placed the sample holder on a support and inserted both of them inside the chamber. 
+A vacuum machine was connected to the accelerator to reduce the pressure inside both the chamber and the beam.
 Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the provided instruction manual. 
@@
 Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
-
 In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples. 
 Finally, we used a collimator with a 2mm radius to collimate the beam.
-
-Finally, we used a computer program to obtain the spectra of the different sources.
-The multichannel analyzer (MCA) used to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
+In order to collect the data, we used a computer program to obtain the spectra of the different sources. 
+The multichannel analyzer (MCA) used to acquire data was gradually set to 70 V and the angle between the top two detectors was $165 ^{\circ}$.
 
@@
 
-For the analysis, we began by calculating the theoretical energy values of the two alphas.
-
-DE ONDE VEM ISTO CARAGO? TEMOS DE TER REFERÊNCIAS
+For the analysis, we began by plotting the theoretical energy values of the two alphas.
 
@@
 
-The energies obtained were 7629.80 kEv and 11031.44 kEv. (FALTA EXPLICAR PERCENTAGENS)
+The energies obtained were 7629.80 kEv and 11031.44 kEv from section [\ref{sec:theoreticalvalues}].
 For the identification of lithium in our samples, we need to compare the theoretical energy values with the experimental ones.
-Since we know the emited alpha in our samples is more energetic than the theoretical values, it is expected that the experimental peaks are slightly higher than the theoretical ones.
-If the samples contains a peak in the proximity of the alpha peaks it shows that the sample contains lithium.
+Since we know the emitted alpha particles from our samples are more energetic than their theoretical values, it is expected that the experimental peaks are slightly ahead of the theoretical ones in the spectrum.
+If there is a peak in the spectrum that is close to the theoretical value, we can confirm the presence of lithium in the sample.
+For the reasons given in section [\ref{sec:theoreticalvalues}], due to the detector's geometry, we expect to only detect $\alpha_0$.
 
@@
 For the LiF sample, we used the calibration [\ref{sec:calibration}] to convert the channel values into energy values.
-we overlapped the expected $\alpha$ energy with the spectrum, and obtained the following:
+We overlapped the expected $\alpha$ energy with the spectrum, and obtained the following:
 
@@
 Moreover, it is according to the expected hypothesis that the observed lithium energy is slightly higher than the theoretical value.
-Analysing this spectrum further:
+% Analysing this spectrum further:
 
@@
 The experimental $\alpha_0$ is very similar to the theoretical value, hence we can confirm the presence of lithium in this sample.
-If we analyze this spectrum further, we can see the following:
+% If we analyze this spectrum further, we can see the following:
 
@@
 One more time we can confirm the presence of lithium in this sample.
-
-If we analyze the spectrum further:
+% If we analyze the spectrum further:
 
@@
 
-To conclude, we ended up with a successful experiment. We were able to detect the presence of lithium in all the 
-analysed samples trough nuclear reactions. Furthermore, we managed to go beyond our proposal and estimate the amount of 
-lithium in each sample.
-The results obtained were in accordance with the values obtained from the Danish Institute that provided the samples. MEIO INVENTADO
+To sum up, we were able to successfully conduct the proposed research project.
+Nuclear reactions is a very effective technique to detect the presence of elements in samples, although its cost is high.
+We were able to detect the presence of lithium in all the samples. 
+Furthermore, we managed to go beyond our proposal and estimate the amount of lithium in each sample.
+% The results obtained were in accordance with the values obtained from the Danish Institute that provided the samples. MEIO INVENTADO
 

### 2024-01-20 text/paper/paper.tex
Saturday, 20 January 2024 at 16:53:26

@@
 The experimental set-up consists of a Tandem accelerator, a beam line with electromagnets and a chamber with both the samples and the detectors.
-
 Theoretically, the accelerator produces a beam of protons with a kinetic energy of up to 5 MeV, which is deviated by the steering electromagnet before hitting the sample. 
-The electromagnet deflect particles with different angles according to their charge-to-mass ratio and, with that, allows us to remove any unwanted particles from the beam and only let the protons pass through.
-
-The detectors used, which are silicon-based, detect charged particles through the ionisation they produce in the silicon.
-
-In addition, the detectors are connected to a electronic system that amplifies the signal and sends it to a computer where the spectra can be seen.
-
+The electromagnets deflect particles with different angles according to their charge-to-mass ratio and, with that, allows us to remove any unwanted particles from the beam and only let the protons pass through.
+The detectors used are silicon-based and detect charged particles through the ionization they produce in the silicon.
+In addition, the detectors are connected to an electronic system that amplifies the signal and sends it to a computer where the spectra can be seen.
 The beam line is under a vacuum of around $10^{-7}$ mbar in order to avoid the protons colliding with other particles and losing energy. 
 They reach the sample with the specified kinetic energy and cause a nuclear reaction whose products are the charged particles detected. 
-We aim to analyse the energy spectrum of these particles.
+We aim to analyze the energy spectrum of these particles.
 
@@
 In this case, the only other option would be for the $\ce{^8Be}$ to be formed in an excited state and therefore give origin to alpha particles also in an excited state. 
-There is not enough energy for this to happen since the difference between the ground state and the the first state in an helium nuclei is very big, around $INSERT$ MeV.
+There is not enough energy for this to happen since the difference between the ground state and the first state in a helium nucleus is very big.
 \paragraph{}

### 2024-01-20 full_analysis.py
fix: images

@@ -17,10 +17,10 @@ import pandas as pd
 SHOW = False
 
 # Save Pictures 
-SAVE = False
+SAVE = True
 
 # Put Titles
-TITLES = True
+TITLES = False
 
 # Define paths
 # data
@@ -93,7 +93,7 @@ def Detector_Calibration(detector : int = 0, bins : int = 1024) -> np.ndarray:
 # calibration, so if we see peaks above the calibration peaks, we know that
 # it might be Lithium
 # -----------------------------------------------------------------------------
-plt.figure(figsize=(10, 10))
+plt.figure(figsize=(10, 8))
 plt.plot(LIF_Chn0, label='LiF', alpha=0.5)
 plt.plot(LiAlO2_Chn0, label='LiAlO2', alpha=0.5)
 plt.plot(Implanted_Chn0, label='Implanted', alpha=0.5)
@@ -104,8 +104,10 @@ plt.yscale('log')
 plt.xlabel('Channel')
 plt.ylabel('Counts')
 plt.xlim(0, 1024)
-if SAVE: plt.savefig(IMAGE_PATH / 'OverlapCalib.png', dpi=600)
+plt.tight_layout()
 plt.grid()
+plt.tight_layout()
+if SAVE: plt.savefig(IMAGE_PATH / 'OverlapCalib.jpeg', dpi=600)
 if SHOW: plt.show()
 
 
@@ -119,7 +121,7 @@ if SHOW: plt.show()
 # -----------------------------------------------------------------------------
 alphas = pd.read_csv(DATA_PATH / 'alphas.csv') # Load the data
 x = Detector_Calibration(detector=0, bins=1024) # Define the x range
-plt.figure(figsize=(10, 10))
+plt.figure(figsize=(10, 8))
 plt.vlines(alphas['Energy'], 0, alphas['Percentage'], color='r')
 # Put text on the peaks
 for i in range(len(alphas)):
@@ -129,15 +131,16 @@ if TITLES: plt.title('Alpha peaks')
 plt.xlabel('Energy [keV]')
 plt.ylabel('Percentage (%)')
 plt.xlim(0, 13000)
-if SAVE: plt.savefig(IMAGE_PATH / 'AlphaPeaks.png', dpi=600)
 plt.grid()
+plt.tight_layout()
[…]

### 2024-01-20 text/paper/paper.tex
Saturday, 20 January 2024 at 18:04:11
paper.tex

@@
 %    \centering
-%    \includegraphics[scale = 0.6]{images/TT_21_Chn1.png}
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn1.jpeg}
 %    \captionof{figure}{Channel 1 spectrum fit with [\ref{eq:califit}]}
@@
 %    \centering
-%    \includegraphics[scale = 0.6]{images/TT_21_Chn2.png}
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn2.jpeg}
 %    \captionof{figure}{Channel 2 spectrum fit with [\ref{eq:califit}]}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/Chn0_calib_with_Backscattering.png}
+    \includegraphics[scale = 0.3]{../../images/Chn0_calib_with_Backscattering.jpeg}
     \captionof{figure}{Calibration "detector 0"}
@@
 %    \centering
-%    \includegraphics[scale = 0.6]{images/TT_21_Chn1_calib.png}
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn1_calib.jpeg}
 %    \captionof{figure}{Calibration Channel 1}
@@
 %    \centering
-%    \includegraphics[scale = 0.6]{images/TT_21_Chn2_calib.png}
+%    \includegraphics[scale = 0.6]{images/TT_21_Chn2_calib.jpeg}
 %    \captionof{figure}{Calibration Channel 2}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/AlphaPeaks.png}
+    \includegraphics[scale = 0.3]{../../images/AlphaPeaks.jpeg}
     \captionof{figure}{Theoretical energy values of the two alphas}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/OverlapLiF_NC.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapLiF_NC.jpeg}
     \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiF.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiF.jpeg}
     \captionof{figure}{Full Analysis of $\ce{LiF}$ Spectrum}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/OverlapLiAlO2.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapLiAlO2.jpeg}
     \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiAlO2.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiAlO2.jpeg}
     \captionof{figure}{Full Analysis of $\ce{LiAlO_2}$ Spectrum}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/OverlapImplanted.png}
+    \includegraphics[scale = 0.3]{../../images/OverlapImplanted.jpeg}
     \captionof{figure}{Spectrum of the implanted sample}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/FullAnalysisImplanted.png}
+    \includegraphics[scale = 0.3]{../../images/FullAnalysisImplanted.jpeg}
     \captionof{figure}{Full Analysis of the Implanted Sample Spectrum}
@@
 To sum up, we were able to successfully conduct the proposed research project.
-Nuclear reactions is a very effective technique to detect the presence of elements in samples, although its cost is high.
+Nuclear reactions can be a very effective technique to detect the presence of elements in samples, despite its high cost.
 We were able to detect the presence of lithium in all the samples. 

### 2024-01-20 text/paper/paper.tex
paper.tex

@@
 
-\subsection{Procedure}
-    \label{sec:procedure}
-
-We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
-In order to do a calibration we added a triple $\alpha$ source on the holder above the lithium fluoride. 
-Using a pachymeter, we measured the distance between the samples so that afterwards we could control the height of the holder in the chamber and change the beam's target.
-We placed the sample holder on a support and inserted both of them inside the chamber. 
-A vacuum machine was connected to the accelerator to reduce the pressure inside both the chamber and the beam.
-Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the provided instruction manual. 
-Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses. 
-Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
-In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples. 
-Finally, we used a collimator with a 2mm radius to collimate the beam.
-In order to collect the data, we used a computer program to obtain the spectra of the different sources. 
-The multichannel analyzer (MCA) used to acquire data was gradually set to 70 V and the angle between the top two detectors was $165 ^{\circ}$.
+%\subsection{Procedure}
+%    \label{sec:procedure}
+%
+%We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$).
+%In order to do a calibration we added a triple $\alpha$ source on the holder above the lithium fluoride. 
+%Using a pachymeter, we measured the distance between the samples so that afterwards we could control the height of the holder in the chamber and change the beam's target.
+%We placed the sample holder on a support and inserted both of them inside the chamber. 
+%A vacuum machine was connected to the accelerator to reduce the pressure inside both the chamber and the beam.
+%Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the provided instruction manual. 
+%Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses. 
+%Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
+%In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples. 
+%Finally, we used a collimator with a 2mm radius to collimate the beam.
+%In order to collect the data, we used a computer program to obtain the spectra of the different sources. 
+%The multichannel analyzer (MCA) used to acquire data was gradually set to 70 V and the angle between the top two detectors was $165 ^{\circ}$.
 
@@
 The spectrum obtained for the triple source was the following:
-
+%Ea quis voluptate nisi culpa.
+Pariatur ad nostrud magna et.
+Occaecat tempor deserunt enim enim Lorem fugiat magna sint.
+Occaecat duis occaecat dolore cupidatat qui laboris reprehenderit.
+Dolor mollit sit tempor labore ad veniam adipisicing duis magna et duis.
+%Sunt nostrud in dolor tempor minim velit incididunt Lorem adipisicing ullamco ad qui ad.
+Veniam magna anim Lorem in ad id qui Lorem ex cillum.
+Laboris cillum ad anim dolore non aute commodo.
+Nulla commodo reprehenderit mollit adipisicing do cupidatat proident.
+Adipisicing eu id incididunt tempor ad fugiat ullamco sit aliqua nulla ut duis labore nostrud.
+Nostrud deserunt excepteur ex et adipisicing nostrud sint dolor laboris nisi reprehenderit esse eu.
+Ex non fugiat mollit exercitation minim enim id ipsum velit reprehenderit ea proident nisi.
 \begin{center}
@@
     \includegraphics[scale = 0.3]{../../images/Calib.png}
+    \vspace{-5mm}
     \captionof{figure}{Spectrum of the Calibration source}

### 2024-01-20 analise.py
Paper.tex

@@ -114,11 +114,12 @@ plt.plot(TT23_Chn0, label='Channel 0')
 plt.plot(TT23_Chn1, label='Channel 1')
 # plt.plot(TT23_Chn2, label='Channel 2')
 plt.legend()
-plt.title('Sample with LiF')
+# plt.title('Sample with LiF')
 plt.yscale('log')
 plt.xlabel('Channel')
 plt.ylabel('Counts')
-plt.savefig(IMAGE_PATH/'SampleLiF.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'SampleLiF.jpeg')
 plt.show()
 
 # TT_24 -> 1024 bins, 2mC (LiAlO2)
@@ -127,11 +128,12 @@ plt.plot(TT24_Chn0, label='Channel 0')
 plt.plot(TT24_Chn1, label='Channel 1')
 # plt.plot(TT24_Chn2, label='Channel 2')
 plt.legend()
-plt.title('Sample with LiAlO2')
+# plt.title('Sample with LiAlO2')
 plt.yscale('log')
 plt.xlabel('Channel')
 plt.ylabel('Counts')
-plt.savefig(IMAGE_PATH/'SampleLiAlO2.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'SampleLiAlO2.jpeg')
 plt.show()
 
 # # TT_25 -> 1024 bins, 2mC (Li Implantado em Al)
@@ -153,11 +155,12 @@ plt.plot(TT26_Chn0, label='Channel 0')
 plt.plot(TT26_Chn1, label='Channel 1')
 # plt.plot(TT26_Chn2, label='Channel 2')
 plt.legend()
-plt.title('Implanted Sample of Li in Al')
+# plt.title('Implanted Sample of Li in Al')
 plt.yscale('log')
 plt.xlabel('Channel')
 plt.ylabel('Counts')
-plt.savefig(IMAGE_PATH/'ImplantedSample.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'ImplantedSample.jpeg')
 plt.show()
 
 # # TT_27
@@ -215,11 +218,12 @@ plt.plot(x1, gaussian(x1, *popt1), label='Channel 1 fit')
 # Style
 plt.ylim(1, 1e3)
 plt.xlim(830, 950)
-plt.title('Sample with LiF')
+# plt.title('Sample with LiF')
 plt.legend()
[…]

### 2024-01-20 calib.py

@@ -34,12 +34,13 @@ plt.plot(TT21_Chn0, label='Chnannel 0')
 plt.plot(TT21_Chn1, label='Channel 1')
 plt.plot(TT21_Chn2, label='Channel 2')
 plt.legend()
-plt.title('TT_21')
+# plt.title('TT_21')
 plt.yscale('log')
 plt.xlabel('Channel')
 plt.xlim(0, 1024)
 plt.ylabel('Counts')
-plt.savefig(IMAGE_PATH/'TT_21.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'TT_21.jpeg')
 # plt.show()
 
 
@@ -68,14 +69,15 @@ ercov0 = np.sqrt(np.diag(pcov0))
 plt.figure()
 plt.plot(TT21_Chn0, label='Detector 0')
 plt.plot(x, gaussian(x, *popt0), label='Detector 0 fit')
-plt.title('Fit detector 0')
+# plt.title('Fit detector 0')
 plt.legend()
 plt.xlabel('Channel')
 plt.yscale('log')
 plt.ylabel('Counts')
 plt.xlim(500, 700)
 plt.ylim(1, 1e2)
-plt.savefig(IMAGE_PATH/'TT_21_Chn0.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'TT_21_Chn0.jpeg')
 # plt.show()
 
 
@@ -98,14 +100,15 @@ ercov1 = np.sqrt(np.diag(pcov1))
 plt.figure()
 plt.plot(TT21_Chn1, label='Chnannel 1')
 plt.plot(x, gaussian(x, *popt1), label='Chnannel 1 fit')
-plt.title('Fit channel 1')
+# plt.title('Fit channel 1')
 plt.legend()
 plt.xlabel('Channel')
 plt.yscale('log')
 plt.ylabel('Counts')
 plt.xlim(500, 700)
 plt.ylim(1, 1e2)
-plt.savefig(IMAGE_PATH/'TT_21_Chn1.png')
+plt.tight_layout()
+plt.savefig(IMAGE_PATH/'TT_21_Chn1.jpeg')
 # plt.show()
 
 info = [
@@ -127,14 +130,15 @@ ercov2 = np.sqrt(np.diag(pcov2))
 plt.figure()
 plt.plot(TT21_Chn2, label='Chnannel 2')
 plt.plot(x, gaussian(x, *popt2), label='Chnannel 2 fit')
-plt.title('Fit channel 2')
[…]

### 2024-01-20 text/paper/paper.tex
Paper.tex
fix: .bib
ready
Final Paper

--- 2024-01-20 text/paper/paper.tex
@@
 
-\usepackage[style=numeric]{biblatex} % Choose your desired citation style
-\addbibresource{../references/prop.bib} % Specify your .bib file
+\usepackage[backend=biber, style=numeric]{biblatex}
+\addbibresource{paper.bib}
 
@@
 \begin{multicols}{2}
-
+\small
 \section{Main Goal}
@@
     \centering
-    \includegraphics[scale = 0.5]{../../images/TT_21_Chn0.png}
+    \includegraphics[scale = 0.5]{../../images/TT_21_Chn0.jpeg}
     \captionof{figure}{Detector 0 spectrum fit with triple gaussian}
@@
     \centering
-    \includegraphics[scale = 0.3]{../../images/AlphaPeaks.jpeg}
+    \includegraphics[scale = 0.28]{../../images/AlphaPeaks.jpeg}
     \captionof{figure}{Theoretical energy values of the two alphas}
@@
 With this we get:
-
-Implanted Sample: $N\Delta x = 9.25\times 10^{16}$ atoms per cm$^2$\\
-$\ce{LiAlO2}$: $N\tau_i=4.31\times10^{17}$ atoms per cm$^2$\\
-$\ce{LiF}$: $N\tau_i=1.15\times10^{18}$ atoms per cm$^2$
+\begin{itemize}[noitemsep, left=1cm]
+    \item Implanted Sample: $N\Delta x = 9.25\times 10^{16}$ atoms per cm$^2$\\
+    \item $\ce{LiAlO2}$: $N\tau_i=4.31\times10^{17}$ atoms per cm$^2$\\
+    \item $\ce{LiF}$: $N\tau_i=1.15\times10^{18}$ atoms per cm$^2$
+\end{itemize}
 
 For comparison:
-$\ce{LiAlO2}$: $N\Delta x=2.15\times10^{19}$ atoms per cm$^2$\\
-$\ce{LiF}$: $N\Delta x=5.71\times10^{19}$ atoms per cm$^2$\\
+\begin{itemize}[noitemsep, left=1cm]
+\item $\ce{LiAlO2}$: $N\Delta x=2.15\times10^{19}$ atoms per cm$^2$\\
+\item $\ce{LiF}$: $N\Delta x=5.71\times10^{19}$ atoms per cm$^2$\\
+\end{itemize}
 
@@
 
+\nocite{*}
 \printbibliography
-\nocite{*}
 

### 2024-01-24 text/LogBooks/logbook2.tex
Wednesday, 24. January 2024 at 16:05:01

@@
 So that we can do a calibration from the number of counts per channel to energy, we put an additional americium-241 source on the holder.
-Using a ruler, we measured the distance between the samples.
-We put the sample holder on a support and afterwards both of them inside a chamber.
-A vacuum machine was connected to the accelerator to reduce the pressure inside the accelerator.
+Using a pachymeter, we measured the distance between the samples.
+We put the sample holder on a support and afterwards both of them inside the chamber.
+The vacuum machine was connected to the accelerator to reduce the pressure inside the accelerator (measured from the top of the container).
 
@@
 \centering
-\begin{tabular}{|w{2cm}|w{4cm}|}
+\begin{tabular}{|w{4cm}|w{3cm}|}
 \hline
@@
 Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the manuals provided.
-Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses.
+This was a very time consuming step that took all morning.
+We had to start the high voltage supply, start the vacuum in the first chamber of the accelerator and then extend it to the accelerator as a whole and gradually increase all the accelerator parameters to the standards used in previous logbooks.
+
+In the afternoon, we fisrtly turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses.
 Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.
@@
 In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples.
+
+The electromagnets were set with the following currents:
+- low energy switch magnet current - 1.5 A;
+- high energy switch 90º magnet current - 13.1 A;
+- second high energy switch magnet current - 13.9 A;
+
 Then we turned on the camera next to the samples to observe if it was well adjusted.
-The beam wasn't rightly centred so we adjusted the angles and positions of the lenses until it was in the correct position.
+The beam wasn't rightly centred so we adjusted the angles and positions of the lenses until it was in the correct position using electrostatic steering.
 Finally, we used a collimator with a 2mm radius to collimate the beam.
 
+The angle between the top two detectors was 165º, a value used in the proposal to get an estimate for the energy used for the beam.
+The final energy reached by the tandem accelerator was 650 keV, or 1300 keV total (as the tandem accelerator has 2 tubes).
+
 We used a computer program to obtain the spectra of the different sources.
-To start, we obtained the spectrum of americium-241 in order to do a calibration posteriorly.
-Then we obtained the spectra of the different sources.
-The multichannel analyzer (MCA) utilized to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.
+To start, we obtained the spectrum of americium-241 in order to do a calibration subsequently.
+Then we changed the acquisition mode from time to charge to obtain the spectra of the different sources, as the time mode is only used for calibrations.
+The multichannel analyser (MCA) utilised to acquire data was gradually set to 70 V.
 
-After obtaining the lithium spectra, we obtained the spectrum of borum as well. Borum can be detected in the middle of the sources since the encapsulation of the sources is made of boric acid.
+After obtaining the lithium spectra, we obtained a spectrum with borum by aiming the detector at the middle of the encapsulated sources (approximately 5.4 cm from the top), since their encapsulation is made of boric acid.
 
-eletromagnets adjustment: there are 3 magnets:
-- low energy switch magnet current - 1.7 A;
-- high energy switch 90º magnet current - 13.2 A;
-- second high energy switch magnet current - 14.4 A;
-1315 keV
+We obtained 8 files with our spectra, all with 1024 channels: from TT\_20 to TT\_27.
+In those files we were only interested in the unfiltered spectra of channels 0, 1 and 2 - channel 2 corresponded to the detector under the sample holder.
 
-proton source adjustment - 15 A
+TT\_20 was a very short calibration that shall not be used.
 
-Temos um feiexe e conseguimos ver.
-ajustamos o feixe com os campos magenticos e com o steering e com o electrostatic steerer no fim.
-Depois colimamos o nosso feixe através de metodos oftalmologicos.
-(colimador de raio 2mm)
+TT\_21 was the used calibration and it took 5 minutes to acquire.
 
-Eletronica de aquisicao de dados a 70V, evoluindo gradualmente.O angulo entre os detetores é 165º (valor q usamos no calculo da energia)
-Offset porta alvos - 15mm
-Escrever a distancia entre as amostras no alvo
+TT\_22 was our first acquisition of the top sample (LiF). Max charge was set to 2 micro Coulombs.
 
-Virar po porta alvos de costas para ter o americio241 de frente de modo a calibrar
+TT\_23 was the acquisition of the same sample but with with a max charge of 2 micro Coulombs.
 
-Desviar a amostra 18 mm de modo a ter a 1 amostra a contar de cima.
-Mudar o modo de aquisição de tempo para carga (tempo para a fonte radioativa para calibrar e carga para as fontes com litio para o detetar)
-O encapsulamento das amostras é feito com acido borico pelo que a detecao de boro é no meio das amostras (pastilhas).
+TT\_24 was the acquisition of LiAlO2 with a max charge of 2 micro Coulombs.
 
-We left CTN at around 16:45 PM.
+TT\_25 had the spectrum of the implanted sample with a 2 micro Coulombs max charge.
+
+TT\_26 was the same sample but with a max charge of 20 micro Coulombs as the implanted lithium was had very small peaks.
+
+TT\_27 we were aiming at the middle of the sources to obtain a spectrum with borum. Max charge of 20 micro Coulombs.
+
+At the end of the acquisition we had to gradually reduce the current and voltage of the accelerator in order to turn it off. Basically doing the inverse procedure we did in the morning.
+
+We left CTN at around 4:45 PM.
+
+\vspace{3cm}
 
